### 2022-01-28 Descriptives.Rmd
Created descriptives file, and added initial data loading

---
title: "Descriptives"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title")

sum(struc_include$imgincl_t1w_include)
sum(struc_include$imgincl_t2w_include)
```

Data cleaning
```{r}

```

### 2022-02-01 Descriptives.Rmd
Added descriptives and plots

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+packages <-  c("tidyverse",
+               "reshape2",
+               "nlme", "lme4",
+               "data.table", "psych",
+               "parallel","lubridate",
+               "ggpubr", "broom", 
+               "apaTables", "MetBrewer")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+lapply(packages, library, character.only = TRUE)
 ```
@@
 struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
-  filter(!collection_title=="collection_title")
+  filter(!collection_title=="collection_title") %>% 
+  select(1:12) %>% 
+  mutate(interview_age = as.numeric(interview_age),
+        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
+        imgincl_t2w_include = as.numeric(imgincl_t2w_include))
 
-sum(struc_include$imgincl_t1w_include)
-sum(struc_include$imgincl_t2w_include)
+struc_include_baseline <- struc_include %>% 
+  filter(eventname == "baseline_year_1_arm_1")
+
+struc_include_followup <- struc_include %>% 
+  filter(eventname == "2_year_follow_up_y_arm_1")
+
 ```
 
-Data cleaning
+Basic descriptives
 ```{r}
+t1_baseline = sum(struc_include_baseline$imgincl_t1w_include)
+t2_baseline = sum(struc_include_baseline$imgincl_t2w_include)
+
+t1_followup = sum(struc_include_followup$imgincl_t1w_include)
+t2_followup = sum(struc_include_followup$imgincl_t2w_include)
+
+t1_baseline_age_mean = mean(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t1w_include==1], na.rm=T)
+t2_baseline_age_mean = mean(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t2w_include==1], na.rm=T)
+t1_baseline_age_sd = sd(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t1w_include==1], na.rm=T)
+
+t1_followup_age_mean = mean(struc_include_followup$interview_age[struc_include_followup$imgincl_t1w_include==1], na.rm=T)
+t2_followup_age_mean = mean(struc_include_followup$interview_age[struc_include_followup$imgincl_t2w_include==1], na.rm=T)
+t1_followup_age_sd = sd(struc_include_followup$interview_age[struc_include_followup$imgincl_t1w_include==1], na.rm=T)
 
 ```
+
+Descriptive plots
+```{r}
+ggplot(struc_include_baseline %>% filter(imgincl_t1w_include==1), aes(x=interview_age)) + 
+  geom_histogram(binwidth=1, fill=met.brewer("Hokusai2", 1), color=met.brewer("Hokusai3", 1)) +
+  theme_classic() +
+  labs(x="Age (Months)", title = "Baseline Age Distribution")
+ggsave("BaselineAgeDist.png")
+
+ggplot(struc_include_followup %>% filter(imgincl_t1w_include==1), aes(x=interview_age)) + 
+  geom_histogram(binwidth=1, fill=met.brewer("Hokusai2", 1), color=met.brewer("Hokusai3", 1)) +
+  theme_classic() +
+  labs(x="Age (Months)", title = "2yr Follow-up Age Distribution")
+ggsave("2YrFollowupAgeDist.png")
+
+```

### 2022-02-17 DataSetup.Rmd
Created Rmd for importing data

---
title: "DataSetup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load Structural Data
```{r}
#Load list of included subjects
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(matches("vol|area"))

#T1 intensity, might not need
smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id)

#Rename data to fit naming conventions
#????

```

Load Other Data
```{r}
#Puberty
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title")

youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
  filter(!collection_title=="collection_title")

#Cognition
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)

#Height/weight metrics
physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)

#Sexual attraction and behavior
gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")

#Grade level, parent report
grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
```

```{r}

```

### 2022-02-17 DataSetup.Rmd
Started renaming abcd variables to match existing model

@@
 
-```{r}
+Template for renaming sMRI variables
+```{r, eval=F}
+renamed_smri <- smri_1 %>% 
+  rename(new_name = old_name)
+
+FS_InterCranial_Vol = 	smri_vol_scs_intracranialv #technically inter vs intra
+FS_BrainSeg_Vol
+FS_BrainSeg_Vol_No_Vent
+FS_BrainSeg_Vol_No_Vent_Surf
+FS_LCort_GM_Vol
+FS_RCort_GM_Vol
+FS_TotCort_GM_Vol
+FS_SubCort_GM_Vol
+FS_Total_GM_Vol
+FS_SupraTentorial_Vol
+FS_SupraTentorial_Vol_No_Vent
+FS_SupraTentorial_No_Vent_Voxel_Count
+FS_Mask_Vol
+FS_BrainSegVol_eTIV_Ratio
+FS_MaskVol_eTIV_Ratio
+FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh
+FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh
+FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh
+FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh
+FS_L_ThalamusProper_Vol = smri_vol_scs_tplh
+FS_L_Caudate_Vol = smri_vol_scs_caudatelh
+FS_L_Putamen_Vol = 	smri_vol_scs_putamenlh
+FS_L_Pallidum_Vol = 	smri_vol_scs_pallidumlh
+FS_3rdVent_Vol = 	smri_vol_scs_3rdventricle
+FS_4thVent_Vol = smri_vol_scs_4thventricle
+FS_BrainStem_Vol = smri_vol_scs_bstem
+FS_L_Hippo_Vol = smri_vol_scs_hpuslh
+FS_L_Amygdala_Vol = 	smri_vol_scs_amygdalalh
+FS_CSF_Vol = 	smri_vol_scs_csf
+FS_L_AccumbensArea_Vol = 	smri_vol_scs_aal
+FS_L_VentDC_Vol = smri_vol_scs_vedclh
+FS_L_Vessel_Vol
+FS_L_ChoroidPlexus_Vol
+FS_R_LatVent_Vol = 	smri_vol_scs_ltventriclerh
+FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh
+FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh
+# no cerebellum cort vol?
+FS_R_ThalamusProper_Vol = smri_vol_scs_tprh
+FS_R_Caudate_Vol = smri_vol_scs_caudaterh
+FS_R_Putamen_Vol = 	smri_vol_scs_putamenrh
+FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh
+FS_R_Hippo_Vol = 	smri_vol_scs_hpusrh
+FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh
+FS_R_AccumbensArea_Vol = 	smri_vol_scs_aar
+FS_R_VentDC_Vol = smri_vol_scs_vedcrh
+FS_R_Vessel_Vol
+FS_R_ChoroidPlexus_Vol
+FS_OpticChiasm_Vol
+FS_CC_Posterior_Vol = smri_vol_scs_ccps
+FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps
+FS_CC_Central_Vol = smri_vol_scs_ccct
+FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat
+FS_CC_Anterior_Vol = smri_vol_scs_ccat
+FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh
+FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh
+FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh
+FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh
+FS_L_Entorhinal_Area = smri_area_cdk_ehinallh
+FS_L_Fusiform_Area = smri_area_cdk_fusiformlh
+FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh
+FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh
+FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh
+FS_L_Lateraloccipital_Area = smri_area_cdk_locclh
+FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh
+FS_L_Lingual_Area = smri_area_cdk_linguallh
+FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh
+FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh
+FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh
+FS_L_Paracentral_Area = smri_area_cdk_paracnlh
+FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh
+FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh
+FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh
+FS_L_Pericalcarine_Area = smri_area_cdk_pericclh
+FS_L_Postcentral_Area = smri_area_cdk_postcnlh
+FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh
+FS_L_Precentral_Area = smri_area_cdk_precnlh
+FS_L_Precuneus_Area = smri_area_cdk_pclh
+FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh
+FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh
+FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh
+FS_L_Superiorparietal_Area = smri_area_cdk_supllh
+FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh
+FS_L_Supramarginal_Area = smri_area_cdk_smlh
+FS_L_Frontalpole_Area = smri_area_cdk_frpolelh
+FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh
+FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh
+FS_L_Insula_Area = smri_area_cdk_insulalh
+FS_R_Bankssts_Area
+FS_R_Caudalanteriorcingulate_Area
+FS_R_Caudalmiddlefrontal_Area
+FS_R_Cuneus_Area
+FS_R_Entorhinal_Area
+FS_R_Fusiform_Area
+FS_R_Inferiorparietal_Area
+FS_R_Inferiortemporal_Area
+FS_R_Isthmuscingulate_Area
+FS_R_Lateraloccipital_Area
+FS_R_Lateralorbitofrontal_Area
+FS_R_Lingual_Area
+FS_R_Medialorbitofrontal_Area
+FS_R_Middletemporal_Area
+FS_R_Parahippocampal_Area
+FS_R_Paracentral_Area
+FS_R_Parsopercularis_Area
+FS_R_Parsorbitalis_Area
+FS_R_Parstriangularis_Area
+FS_R_Pericalcarine_Area
+FS_R_Postcentral_Area
+FS_R_Posteriorcingulate_Area
+FS_R_Precentral_Area
+FS_R_Precuneus_Area
+FS_R_Rostralanteriorcingulate_Area
+FS_R_Rostralmiddlefrontal_Area
+FS_R_Superiorfrontal_Area
+FS_R_Superiorparietal_Area
+FS_R_Superiortemporal_Area
+FS_R_Supramarginal_Area
+FS_R_Frontalpole_Area
+FS_R_Temporalpole_Area
+FS_R_Transversetemporal_Area
+FS_R_Insula_Area
+FS_L_Bankssts_GrayVol
+FS_L_Caudalanteriorcingulate_GrayVol
+FS_L_Caudalmiddlefrontal_GrayVol
+FS_L_Cuneus_GrayVol
+FS_L_Entorhinal_GrayVol
+FS_L_Fusiform_GrayVol
+FS_L_Inferiorparietal_GrayVol
+FS_L_Inferiortemporal_GrayVol
+FS_L_Isthmuscingulate_GrayVol
+FS_L_Lateraloccipital_GrayVol
+FS_L_Lateralorbitofrontal_GrayVol
+FS_L_Lingual_GrayVol
+FS_L_Medialorbitofrontal_GrayVol
+FS_L_Middletemporal_GrayVol
+FS_L_Parahippocampal_GrayVol
+FS_L_Paracentral_GrayVol
+FS_L_Parsopercularis_GrayVol
+FS_L_Parsorbitalis_GrayVol
+FS_L_Parstriangularis_GrayVol
+FS_L_Pericalcarine_GrayVol
+FS_L_Postcentral_GrayVol
+FS_L_Posteriorcingulate_GrayVol
+FS_L_Precentral_GrayVol
+FS_L_Precuneus_GrayVol
+FS_L_Rostralanteriorcingulate_GrayVol
+FS_L_Rostralmiddlefrontal_GrayVol
+FS_L_Superiorfrontal_GrayVol
+FS_L_Superiorparietal_GrayVol
+FS_L_Superiortemporal_GrayVol
+FS_L_Supramarginal_GrayVol
+FS_L_Frontalpole_GrayVol
+FS_L_Temporalpole_GrayVol
+FS_L_Transversetemporal_GrayVol
+FS_L_Insula_GrayVol
+FS_R_Bankssts_GrayVol
+FS_R_Caudalanteriorcingulate_GrayVol
+FS_R_Caudalmiddlefrontal_GrayVol
+FS_R_Cuneus_GrayVol
+FS_R_Entorhinal_GrayVol
+FS_R_Fusiform_GrayVol
+FS_R_Inferiorparietal_GrayVol
+FS_R_Inferiortemporal_GrayVol
+FS_R_Isthmuscingulate_GrayVol
+FS_R_Lateraloccipital_GrayVol
+FS_R_Lateralorbitofrontal_GrayVol
+FS_R_Lingual_GrayVol
+FS_R_Medialorbitofrontal_GrayVol
+FS_R_Middletemporal_GrayVol
+FS_R_Parahippocampal_GrayVol
+FS_R_Paracentral_GrayVol
+FS_R_Parsopercularis_GrayVol
+FS_R_Parsorbitalis_GrayVol
+FS_R_Parstriangularis_GrayVol
+FS_R_Pericalcarine_GrayVol
+FS_R_Postcentral_GrayVol
+FS_R_Posteriorcingulate_GrayVol
+FS_R_Precentral_GrayVol
+FS_R_Precuneus_GrayVol
+FS_R_Rostralanteriorcingulate_GrayVol
+FS_R_Rostralmiddlefrontal_GrayVol
+FS_R_Superiorfrontal_GrayVol
+FS_R_Superiorparietal_GrayVol
+FS_R_Superiortemporal_GrayVol
+FS_R_Supramarginal_GrayVol
+FS_R_Frontalpole_GrayVol
+FS_R_Temporalpole_GrayVol
+FS_R_Transversetemporal_GrayVol
+FS_R_Insula_GrayVol
 

### 2022-02-17 DataSetup.Rmd
Added additional new names for smri variables

@@
 FS_BrainSeg_Vol_No_Vent_Surf
-FS_LCort_GM_Vol
-FS_RCort_GM_Vol
-FS_TotCort_GM_Vol
-FS_SubCort_GM_Vol
-FS_Total_GM_Vol
-FS_SupraTentorial_Vol
-FS_SupraTentorial_Vol_No_Vent
+FS_LCort_GM_Vol = smri_vol_cdk_totallh
+FS_RCort_GM_Vol = smri_vol_cdk_totalrh
+FS_TotCort_GM_Vol = smri_vol_cdk_total #total whole brain cortical volume
+FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv
+FS_Total_GM_Vol #maybe just calculate?
+FS_SupraTentorial_Vol = smri_vol_scs_suprateialv #assuming this includes all ventricles
+FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
 FS_SupraTentorial_No_Vent_Voxel_Count
@@
 FS_L_Insula_Area = smri_area_cdk_insulalh
-FS_R_Bankssts_Area
-FS_R_Caudalanteriorcingulate_Area
-FS_R_Caudalmiddlefrontal_Area
-FS_R_Cuneus_Area
-FS_R_Entorhinal_Area
-FS_R_Fusiform_Area
-FS_R_Inferiorparietal_Area
-FS_R_Inferiortemporal_Area
-FS_R_Isthmuscingulate_Area
-FS_R_Lateraloccipital_Area
-FS_R_Lateralorbitofrontal_Area
-FS_R_Lingual_Area
-FS_R_Medialorbitofrontal_Area
-FS_R_Middletemporal_Area
-FS_R_Parahippocampal_Area
-FS_R_Paracentral_Area
-FS_R_Parsopercularis_Area
-FS_R_Parsorbitalis_Area
-FS_R_Parstriangularis_Area
-FS_R_Pericalcarine_Area
-FS_R_Postcentral_Area
-FS_R_Posteriorcingulate_Area
-FS_R_Precentral_Area
-FS_R_Precuneus_Area
-FS_R_Rostralanteriorcingulate_Area
-FS_R_Rostralmiddlefrontal_Area
-FS_R_Superiorfrontal_Area
-FS_R_Superiorparietal_Area
-FS_R_Superiortemporal_Area
-FS_R_Supramarginal_Area
-FS_R_Frontalpole_Area
-FS_R_Temporalpole_Area
-FS_R_Transversetemporal_Area
-FS_R_Insula_Area
-FS_L_Bankssts_GrayVol
-FS_L_Caudalanteriorcingulate_GrayVol
-FS_L_Caudalmiddlefrontal_GrayVol
-FS_L_Cuneus_GrayVol
-FS_L_Entorhinal_GrayVol
-FS_L_Fusiform_GrayVol
-FS_L_Inferiorparietal_GrayVol
-FS_L_Inferiortemporal_GrayVol
-FS_L_Isthmuscingulate_GrayVol
-FS_L_Lateraloccipital_GrayVol
-FS_L_Lateralorbitofrontal_GrayVol
-FS_L_Lingual_GrayVol
-FS_L_Medialorbitofrontal_GrayVol
-FS_L_Middletemporal_GrayVol
-FS_L_Parahippocampal_GrayVol
-FS_L_Paracentral_GrayVol
-FS_L_Parsopercularis_GrayVol
-FS_L_Parsorbitalis_GrayVol
-FS_L_Parstriangularis_GrayVol
-FS_L_Pericalcarine_GrayVol
-FS_L_Postcentral_GrayVol
-FS_L_Posteriorcingulate_GrayVol
-FS_L_Precentral_GrayVol
-FS_L_Precuneus_GrayVol
-FS_L_Rostralanteriorcingulate_GrayVol
-FS_L_Rostralmiddlefrontal_GrayVol
-FS_L_Superiorfrontal_GrayVol
-FS_L_Superiorparietal_GrayVol
-FS_L_Superiortemporal_GrayVol
-FS_L_Supramarginal_GrayVol
-FS_L_Frontalpole_GrayVol
-FS_L_Temporalpole_GrayVol
-FS_L_Transversetemporal_GrayVol
-FS_L_Insula_GrayVol
-FS_R_Bankssts_GrayVol
-FS_R_Caudalanteriorcingulate_GrayVol
-FS_R_Caudalmiddlefrontal_GrayVol
-FS_R_Cuneus_GrayVol
-FS_R_Entorhinal_GrayVol
-FS_R_Fusiform_GrayVol
-FS_R_Inferiorparietal_GrayVol
-FS_R_Inferiortemporal_GrayVol
-FS_R_Isthmuscingulate_GrayVol
-FS_R_Lateraloccipital_GrayVol
-FS_R_Lateralorbitofrontal_GrayVol
-FS_R_Lingual_GrayVol
-FS_R_Medialorbitofrontal_GrayVol
-FS_R_Middletemporal_GrayVol
-FS_R_Parahippocampal_GrayVol
-FS_R_Paracentral_GrayVol
-FS_R_Parsopercularis_GrayVol
-FS_R_Parsorbitalis_GrayVol
-FS_R_Parstriangularis_GrayVol
-FS_R_Pericalcarine_GrayVol
-FS_R_Postcentral_GrayVol
-FS_R_Posteriorcingulate_GrayVol
-FS_R_Precentral_GrayVol
-FS_R_Precuneus_GrayVol
-FS_R_Rostralanteriorcingulate_GrayVol
-FS_R_Rostralmiddlefrontal_GrayVol
-FS_R_Superiorfrontal_GrayVol
-FS_R_Superiorparietal_GrayVol
-FS_R_Superiortemporal_GrayVol
-FS_R_Supramarginal_GrayVol
-FS_R_Frontalpole_GrayVol
-FS_R_Temporalpole_GrayVol
-FS_R_Transversetemporal_GrayVol
-FS_R_Insula_GrayVol
+FS_R_Bankssts_Area = smri_area_cdk_banksstsrh
+FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh
+FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh
+FS_R_Cuneus_Area = smri_area_cdk_cuneusrh
+FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh
+FS_R_Fusiform_Area = smri_area_cdk_fusiformrh
+FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh
+FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh
+FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh
+FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh
+FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh
+FS_R_Lingual_Area = smri_area_cdk_lingualrh
+FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh
+FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh
+FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh
+FS_R_Paracentral_Area = smri_area_cdk_paracnrh
+FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh
+FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh
+FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh
+FS_R_Pericalcarine_Area = smri_area_cdk_periccrh
+FS_R_Postcentral_Area = smri_area_cdk_postcnrh
+FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh
+FS_R_Precentral_Area = smri_area_cdk_precnrh
+FS_R_Precuneus_Area = smri_area_cdk_pcrh
+FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh
+FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh
+FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh
+FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh
+FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh
+FS_R_Supramarginal_Area = smri_area_cdk_smrh
+FS_R_Frontalpole_Area = smri_area_cdk_frpolerh
+FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh
+FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh
+FS_R_Insula_Area = 	smri_area_cdk_insularh
+#abcd technically doesn't say gray matter volume for the following varibles
+FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh
+FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh
+FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh
+FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh
+FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh
+FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh
+FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh
+FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh
+FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh
+FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh
+FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh
+FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh
+FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh
+FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh
+FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh
+FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh
+FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh
+FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh
+FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh
+FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh
+FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh
+FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh
+FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh
+FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh
+FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh
+FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh
+FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh
+FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh
+FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh
+FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh
+FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh
+FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh
+FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh
+FS_L_Insula_GrayVol = smri_vol_cdk_insulalh
+FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh
+FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh
+FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh
+FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh
+FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh
+FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh
+FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh
+FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh
+FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh
+FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh
+FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh
+FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh
+FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh
+FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh
+FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh
+FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh
+FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh
+FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh
+FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh
+FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh
+FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh
+FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh
+FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh
+FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh
+FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh
+FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh
+FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh
+FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh
+FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh
+FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh
+FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh
+FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh
+FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh
+FS_R_Insula_GrayVol = smri_vol_cdk_insularh
 

### 2022-02-17 DataSetup.Rmd
Updated variable name changes, added script for running model

--- 2022-02-17 DataSetup.Rmd
@@
   filter(src_subject_id%in%struc_include$src_subject_id) %>% 
-  select(matches("vol|area"))
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))
+
+smri_baseline <- smri_1%>% 
+  filter(eventname=="baseline_year_1_arm_1")
 
@@
 
-```{r}
+Template for renaming sMRI variables
+```{r, eval=F}
+renamed_smri <- smri_baseline %>% 
+  rename(FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
+#FS_BrainSeg_Vol
+#FS_BrainSeg_Vol_No_Vent
+#FS_BrainSeg_Vol_No_Vent_Surf
+FS_LCort_GM_Vol = smri_vol_cdk_totallh,
+FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
+FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
+FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
+#FS_Total_GM_Vol #maybe just calculate?
+FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
+#FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
+#FS_SupraTentorial_No_Vent_Voxel_Count
+#FS_Mask_Vol
+#FS_BrainSegVol_eTIV_Ratio
+#FS_MaskVol_eTIV_Ratio
+FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
+FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
+FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
+FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
+FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
+FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
+FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
+FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
+FS_3rdVent_Vol = smri_vol_scs_3rdventricle,
+FS_4thVent_Vol = smri_vol_scs_4thventricle,
+FS_BrainStem_Vol = smri_vol_scs_bstem,
+FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
+FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
+FS_CSF_Vol = smri_vol_scs_csf,
+FS_L_AccumbensArea_Vol = smri_vol_scs_aal,
+FS_L_VentDC_Vol = smri_vol_scs_vedclh,
+#FS_L_Vessel_Vol
+#FS_L_ChoroidPlexus_Vol
+FS_R_LatVent_Vol = 	smri_vol_scs_ltventriclerh,
+FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
+FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
+# no cerebellum cort vol?
+FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
+FS_R_Caudate_Vol = smri_vol_scs_caudaterh,
+FS_R_Putamen_Vol = smri_vol_scs_putamenrh,
+FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh,
+FS_R_Hippo_Vol = smri_vol_scs_hpusrh,
+FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh,
+FS_R_AccumbensArea_Vol = 	smri_vol_scs_aar,
+FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
+#FS_R_Vessel_Vol
+#FS_R_ChoroidPlexus_Vol
+#FS_OpticChiasm_Vol
+FS_CC_Posterior_Vol = smri_vol_scs_ccps,
+FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps,
+FS_CC_Central_Vol = smri_vol_scs_ccct,
+FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat,
+FS_CC_Anterior_Vol = smri_vol_scs_ccat,
+FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh,
+FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh,
+FS_L_Entorhinal_Area = smri_area_cdk_ehinallh,
+FS_L_Fusiform_Area = smri_area_cdk_fusiformlh,
+FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh,
+FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh,
+FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh,
+FS_L_Lateraloccipital_Area = smri_area_cdk_locclh,
+FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh,
+FS_L_Lingual_Area = smri_area_cdk_linguallh,
+FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh,
+FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh,
+FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh,
+FS_L_Paracentral_Area = smri_area_cdk_paracnlh,
+FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh,
+FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh,
+FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh,
+FS_L_Pericalcarine_Area = smri_area_cdk_pericclh,
+FS_L_Postcentral_Area = smri_area_cdk_postcnlh,
+FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh,
+FS_L_Precentral_Area = smri_area_cdk_precnlh,
+FS_L_Precuneus_Area = smri_area_cdk_pclh,
+FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh,
+FS_L_Superiorparietal_Area = smri_area_cdk_supllh,
+FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh,
+FS_L_Supramarginal_Area = smri_area_cdk_smlh,
+FS_L_Frontalpole_Area = smri_area_cdk_frpolelh,
+FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh,
+FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh,
+FS_L_Insula_Area = smri_area_cdk_insulalh,
+FS_R_Bankssts_Area = smri_area_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh,
+FS_R_Cuneus_Area = smri_area_cdk_cuneusrh,
+FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh,
+FS_R_Fusiform_Area = smri_area_cdk_fusiformrh,
+FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh,
+FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh,
+FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh,
+FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh,
+FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh,
+FS_R_Lingual_Area = smri_area_cdk_lingualrh,
+FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh,
+FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh,
+FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh,
+FS_R_Paracentral_Area = smri_area_cdk_paracnrh,
+FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh,
+FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh,
+FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh,
+FS_R_Pericalcarine_Area = smri_area_cdk_periccrh,
+FS_R_Postcentral_Area = smri_area_cdk_postcnrh,
+FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh,
+FS_R_Precentral_Area = smri_area_cdk_precnrh,
+FS_R_Precuneus_Area = smri_area_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh,
+FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh,
+FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh,
+FS_R_Supramarginal_Area = smri_area_cdk_smrh,
+FS_R_Frontalpole_Area = smri_area_cdk_frpolerh,
+FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh,
+FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh,
+FS_R_Insula_Area = smri_area_cdk_insularh,
+#abcd technically doesn't say gray matter volume for the following varibles
+FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh,
+FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh,
+FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh,
+FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh,
+FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh,
+FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh,
+FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh,
+FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh,
+FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh,
+FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh,
+FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh,
+FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh,
+FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh,
+FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh,
+FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh,
+FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh,
+FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh,
+FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh,
+FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh,
+FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh,
+FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh,
+FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh,
+FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh,
+FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh,
+FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh,
+FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh,
+FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh,
+FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh,
+FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh,
+FS_L_Insula_GrayVol = smri_vol_cdk_insulalh,
+FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh,
+FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh,
+FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh,
+FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh,
+FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh,
+FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh,
+FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh,
+FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh,
+FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh,
+FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh,
+FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh,
+FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh,
+FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh,
+FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh,
+FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh,
+FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh,
+FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh,
+FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh,
+FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh,
+FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh,
+FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh,
+FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh,
+FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh,
+FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh,
+FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh,
+FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh,
+FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh,
+FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh,
+FS_R_Insula_GrayVol = smri_vol_cdk_insularh
+)
+
+#Create empty columns to match feature list
+renamed_smri[,"FS_BrainSeg_Vol"] <- NA
+renamed_smri[,"FS_BrainSeg_Vol_No_Vent"] <- NA
+renamed_smri[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
+renamed_smri[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
+renamed_smri[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
+renamed_smri[,"FS_Mask_Vol"] <- NA
+renamed_smri[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
+renamed_smri[,"FS_MaskVol_eTIV_Ratio"] <- NA
+renamed_smri[,"FS_L_Vessel_Vol"] <- NA
+renamed_smri[,"FS_L_ChoroidPlexus_Vol"] <- NA
+renamed_smri[,"FS_R_Vessel_Vol"] <- NA
+renamed_smri[,"FS_R_ChoroidPlexus_Vol"] <- NA
+renamed_smri[,"FS_OpticChiasm_Vol"] <- NA
+renamed_smri[,"FS_Total_GM_Vol"] <- NA
+
+#renamed_smri <- subset(renamed_smri, select = -c(269,279))
+
+#Convert to numeric
+renamed_smri[,6:279] <- sapply(renamed_smri[,6:279],as.numeric)
+renamed_smri <- renamed_smri %>% 
+  mutate(interview_age = as.numeric(interview_age)/12)
+
+#Template
+FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
+FS_BrainSeg_Vol
+FS_BrainSeg_Vol_No_Vent
+FS_BrainSeg_Vol_No_Vent_Surf
+FS_LCort_GM_Vol = smri_vol_cdk_totallh,
+FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
+FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
+FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
+FS_Total_GM_Vol #maybe just calculate?
+FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
+FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
+FS_SupraTentorial_No_Vent_Voxel_Count
+FS_Mask_Vol
+FS_BrainSegVol_eTIV_Ratio
+FS_MaskVol_eTIV_Ratio
+FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
+FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
+FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
+FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
+FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
+FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
+FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
+FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
+FS_3rdVent_Vol = 	smri_vol_scs_3rdventricle,
+FS_4thVent_Vol = smri_vol_scs_4thventricle,
+FS_BrainStem_Vol = smri_vol_scs_bstem,
+FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
+FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
+FS_CSF_Vol = 	smri_vol_scs_csf,
+FS_L_AccumbensArea_Vol = 	smri_vol_scs_aal,
+FS_L_VentDC_Vol = smri_vol_scs_vedclh,
+FS_L_Vessel_Vol
+FS_L_ChoroidPlexus_Vol
+FS_R_LatVent_Vol = smri_vol_scs_ltventriclerh,
+FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
+FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
+# no cerebellum cort vol?
+FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
+FS_R_Caudate_Vol = smri_vol_scs_caudaterh,
+FS_R_Putamen_Vol = smri_vol_scs_putamenrh,
+FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh,
+FS_R_Hippo_Vol = 	smri_vol_scs_hpusrh,
+FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh,
+FS_R_AccumbensArea_Vol = smri_vol_scs_aar,
+FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
+FS_R_Vessel_Vol
+FS_R_ChoroidPlexus_Vol
+FS_OpticChiasm_Vol
+FS_CC_Posterior_Vol = smri_vol_scs_ccps,
+FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps,
+FS_CC_Central_Vol = smri_vol_scs_ccct,
+FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat,
+FS_CC_Anterior_Vol = smri_vol_scs_ccat,
+FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh,
+FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh,
+FS_L_Entorhinal_Area = smri_area_cdk_ehinallh,
+FS_L_Fusiform_Area = smri_area_cdk_fusiformlh,
+FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh,
+FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh,
+FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh,
+FS_L_Lateraloccipital_Area = smri_area_cdk_locclh,
+FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh,
+FS_L_Lingual_Area = smri_area_cdk_linguallh,
+FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh,
+FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh,
+FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh,
+FS_L_Paracentral_Area = smri_area_cdk_paracnlh,
+FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh,
+FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh,
+FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh,
+FS_L_Pericalcarine_Area = smri_area_cdk_pericclh,
+FS_L_Postcentral_Area = smri_area_cdk_postcnlh,
+FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh,
+FS_L_Precentral_Area = smri_area_cdk_precnlh,
+FS_L_Precuneus_Area = smri_area_cdk_pclh,
+FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh,
+FS_L_Superiorparietal_Area = smri_area_cdk_supllh,
+FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh,
+FS_L_Supramarginal_Area = smri_area_cdk_smlh,
+FS_L_Frontalpole_Area = smri_area_cdk_frpolelh,
+FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh,
+FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh,
+FS_L_Insula_Area = smri_area_cdk_insulalh,
+FS_R_Bankssts_Area = smri_area_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh,
+FS_R_Cuneus_Area = smri_area_cdk_cuneusrh,
+FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh,
+FS_R_Fusiform_Area = smri_area_cdk_fusiformrh,
+FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh,
+FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh,
+FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh,
+FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh,
+FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh,
+FS_R_Lingual_Area = smri_area_cdk_lingualrh,
+FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh,
+FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh,
+FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh,
+FS_R_Paracentral_Area = smri_area_cdk_paracnrh,
+FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh,
+FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh,
+FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh,
+FS_R_Pericalcarine_Area = smri_area_cdk_periccrh,
+FS_R_Postcentral_Area = smri_area_cdk_postcnrh,
+FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh,
+FS_R_Precentral_Area = smri_area_cdk_precnrh,
+FS_R_Precuneus_Area = smri_area_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh,
+FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh,
+FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh,
+FS_R_Supramarginal_Area = smri_area_cdk_smrh,
+FS_R_Frontalpole_Area = smri_area_cdk_frpolerh,
+FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh,
+FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh,
+FS_R_Insula_Area = 	smri_area_cdk_insularh,
+#abcd technically doesn't say gray matter volume for the following varibles
+FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh,
+FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh,
+FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh,
+FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh,
+FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh,
+FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh,
+FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh,
+FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh,
+FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh,
+FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh,
+FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh,
+FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh,
+FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh,
+FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh,
+FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh,
+FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh,
+FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh,
+FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh,
+FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh,
+FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh,
+FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh,
+FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh,
+FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh,
+FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh,
+FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh,
+FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh,
+FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh,
+FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh,
+FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh,
+FS_L_Insula_GrayVol = smri_vol_cdk_insulalh,
+FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh,
+FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh,
+FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh,
+FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh,
+FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh,
+FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh,
+FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh,
+FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh,
+FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh,
+FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh,
+FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh,
+FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh,
+FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh,
+FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh,
+FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh,
+FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh,
+FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh,
+FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh,
+FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh,
+FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh,
+FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh,
+FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh,
+FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh,
+FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh,
+FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh,
+FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh,
+FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh,
+FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh,
+FS_R_Insula_GrayVol = smri_vol_cdk_insularh
 

### 2022-02-17 ExisitingModel.Rmd
Updated variable name changes, added script for running model

---
title: "Existing Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Install XGboost & tidymodels
packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
# You then install this version of the package using
install.packages(packageUrl, repos = NULL, type = 'source')

install.packages("tidymodels")

library(here) # OS agnostic relative paths (relative to project dir)
library(tidyverse) # data wrangling tools and pipes
library(tidymodels) # machine learning metaverse
library(xgboost) # main engine, needs version 1.0.0.2, see above

```

Load Data
```{r}

#load("smri_1.rds")

xgb_mod <- readRDS(
  file = here::here(
    "xgboost_9to19_brain_age_mod.rds"))
```

Predict Brain Age
```{r}
brain_age_df <-
  xgb_mod %>%
  predict(new_data = renamed_smri) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = renamed_smri$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df %>%
  metrics(truth = truth, estimate = .pred)
```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.
# xgb_bias_mod <- lm(.pred ~ truth, data = xgb_validate)

# extract intercept and slope
bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
bias_slope <-  0.55 # xgb_bias_mod$coefficients[["truth"]]

# create bias correct data frame
brain_age_corrected_df <- brain_age_df %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df %>%
  metrics(truth = truth, estimate = corrected_pred)

```

plotting
```{r}
brain_age_actual_df <- as.data.frame(brain_age_df)

ggplot(brain_age_actual_df, aes(x=truth, y=.pred)) +
  geom_point()

ggplot(brain_age_actual_df, aes(x=.pred)) +
  geom_histogram()

ggplot(brain_age_actual_df, aes(x=truth)) +
  geom_histogram()

ggplot(as.data.frame(brain_age_corrected_df), aes(x=truth, y=.pred)) +
  geom_point()

ggplot(as.data.frame(brain_age_corrected_df), aes(x=corrected_pred)) +
  geom_histogram()
```

### 2022-02-18 DataSetup.Rmd
Added code for 2 year follow up

@@
   filter(eventname=="baseline_year_1_arm_1")
+
+smri_followup <- smri_1%>% 
+  filter(eventname=="2_year_follow_up_y_arm_1")
 
@@
 #Create empty columns to match feature list
-renamed_smri[,"FS_BrainSeg_Vol"] <- NA
-renamed_smri[,"FS_BrainSeg_Vol_No_Vent"] <- NA
-renamed_smri[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
-renamed_smri[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
-renamed_smri[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
-renamed_smri[,"FS_Mask_Vol"] <- NA
-renamed_smri[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
-renamed_smri[,"FS_MaskVol_eTIV_Ratio"] <- NA
-renamed_smri[,"FS_L_Vessel_Vol"] <- NA
-renamed_smri[,"FS_L_ChoroidPlexus_Vol"] <- NA
-renamed_smri[,"FS_R_Vessel_Vol"] <- NA
-renamed_smri[,"FS_R_ChoroidPlexus_Vol"] <- NA
-renamed_smri[,"FS_OpticChiasm_Vol"] <- NA
-renamed_smri[,"FS_Total_GM_Vol"] <- NA
+renamed_smri_followup[,"FS_BrainSeg_Vol"] <- NA
+renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent"] <- NA
+renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
+renamed_smri_followup[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
+renamed_smri_followup[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
+renamed_smri_followup[,"FS_Mask_Vol"] <- NA
+renamed_smri_followup[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
+renamed_smri_followup[,"FS_MaskVol_eTIV_Ratio"] <- NA
+renamed_smri_followup[,"FS_L_Vessel_Vol"] <- NA
+renamed_smri_followup[,"FS_L_ChoroidPlexus_Vol"] <- NA
+renamed_smri_followup[,"FS_R_Vessel_Vol"] <- NA
+renamed_smri_followup[,"FS_R_ChoroidPlexus_Vol"] <- NA
+renamed_smri_followup[,"FS_OpticChiasm_Vol"] <- NA
+renamed_smri_followup[,"FS_Total_GM_Vol"] <- NA
 
@@
 #Convert to numeric
-renamed_smri[,6:279] <- sapply(renamed_smri[,6:279],as.numeric)
-renamed_smri <- renamed_smri %>% 
+renamed_smri_followup[,6:279] <- sapply(renamed_smri_followup[,6:279],as.numeric)
+renamed_smri_followup <- renamed_smri_followup %>% 
   mutate(interview_age = as.numeric(interview_age)/12)

### 2022-02-18 ExisitingModel.Rmd
Added code for 2 year follow up

@@
   mutate(gap = .pred - truth)
+
+brain_age_df_followup <-
+  xgb_mod %>%
+  predict(new_data = renamed_smri_followup) %>%
+  mutate(
+    # provide the chronological age at time of scan
+    truth = renamed_smri_followup$interview_age
+  ) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
 ```
@@
 brain_age_df %>%
+  metrics(truth = truth, estimate = .pred)
+
+brain_age_df_followup %>%
   metrics(truth = truth, estimate = .pred)
@@
 
+brain_age_corrected_df_followup <- brain_age_df_followup %>%
+  mutate(
+    # corrected brain age prediction
+    corrected_pred =  (.pred - bias_intercept) / bias_slope
+  ) %>%
+  # corrected corrected brain age gap
+  mutate(corrected_gap = corrected_pred - truth)
+
 #mine
@@
   metrics(truth = truth, estimate = corrected_pred)
+
+brain_age_corrected_df_followup %>%
+  metrics(truth = truth, estimate = corrected_pred)
+
+save(brain_age_df, file="BaselineBrainAge.Rda")
+save(brain_age_corrected_df, file="BaselineBrainAgeCorrected.Rda")
+
+save(brain_age_df_followup, file="FollowUpBrainAge.Rda")
+save(brain_age_corrected_df_followup, file="FollowUpBrainAgeCorrected.Rda")
 
@@
 
+ggplot(brain_age_df_followup, aes(x=.pred)) +
+  geom_histogram()
+
 ggplot(as.data.frame(brain_age_corrected_df), aes(x=truth, y=.pred)) +
@@
   geom_histogram()
+
+ggplot(as.data.frame(brain_age_corrected_df_followup), aes(x=corrected_pred)) +
+  geom_histogram()
 ```

### 2022-02-24 ABCDModel.Rmd
Added ABCD baseline model (initial version)

---
title: "ABCD Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Model Prep (Baseline)
```{r}
# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split_baseline <- initial_split(
  smri_1,
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

df_train_baseline <- training(df_split_baseline)
df_validation_baseline <- testing(df_split_baseline)

# report n training, features, testing
df_train_baseline %>%
  summarise(
    n_training = n(),
    p_features = ncol(df_train_baseline)
  ) %>%
  bind_cols(df_validation_baseline %>%
              summarise(n_testing = n()))

# Pre-processing setup ----------------------------------------------------

# define (what we want to do)
preprocess_recipe_baseline <- df_train_baseline %>%
  # predict scan age by all brain features
  recipe(scan_age ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe_baseline

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_prep_baseline <- juice(preprocess_recipe_baseline)

# apply on validation
df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(df_validation_baseline)

# pre process on all of forbow
frb_prep <- preprocess_recipe_baseline %>% bake(frb_select)

# Cross-validation --------------------------------------------------------

# 10 fold cv repeated 10 times
set.seed(42)
train_cv_baseline <- df_train_prep_baseline %>%
  vfold_cv(
    v = 10, 
    repeats = 10, 
    strata = interview_age
  )

```

Train and Fit Model (Baseline)
```{r}
# Extreme gradient boosting
# Parallel processing -----------------------------------------------------

library(doParallel)
all_cores <- parallel::detectCores(logical = TRUE)
# for 8 core 16 thread machine, good performance running more than 
# physical but less than all logical
registerDoParallel(cores = all_cores - 6)

# Model specification -----------------------------------------------------

boost_mod <- boost_tree(
  mode = "regression", 
  trees = 1500, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
  ) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

# Model parameters --------------------------------------------------------

set.seed(42)

# filled 6 dimensional tuning space
xgboost_grid <- grid_latin_hypercube(
  min_n(), tree_depth(), loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_prep_baseline),
  learn_rate(),
  size = 500 
)

# Model tuning -----------------------------------------------------------

tictoc::tic()

set.seed(42)
xgb_tuned_results <- tune_grid(
  boost_mod,
  interview_age ~ .,
  resamples = train_cv,
  grid = xgboost_grid,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = FALSE,
                         save_pred = TRUE)
)

tictoc::toc()
beepr::beep(2)

# Pick best model ---------------------------------------------------------

xgb_tuned_results %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params <- xgb_tuned_results %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

# Fit and save best model -------------------------------------------------

xgb_final_mod <- boost_mod %>%
  finalize_model(best_xgb_params) %>%
  fit(interview_age ~ .,
      data = df_train_prep)

# save mod
saveRDS(xgb_final_mod, file = here::here(
  "model", "xgboost_abcd_brain_age_mod.rds")
  )

```

### 2022-03-16 ABCDModel.Rmd
Changed variable names to match sample

@@
                "ggpubr", "broom", 
-               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc")
+               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc", "rsample", "tidymodels")
 if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
@@
   # predict scan age by all brain features
-  recipe(scan_age ~ .) %>%
+  recipe(interview_age ~ .) %>%
   # remove near zero variance predictors
@@
 # pre process on all of forbow
-frb_prep <- preprocess_recipe_baseline %>% bake(frb_select)
+#frb_prep <- preprocess_recipe_baseline %>% bake(frb_select)
 
@@
     v = 10, 
-    repeats = 10, 
+    repeats = 10, #10
     strata = interview_age
@@
   interview_age ~ .,
-  resamples = train_cv,
+  resamples = train_cv_baseline,
   grid = xgboost_grid,

### 2022-03-16 model-prep.R

@@ -1,0 +1,61 @@
+# Model prep: split, preprocessing, CV ------------------------------------
+
+# Train / test split ------------------------------------------------------
+
+set.seed(42)
+df_split <- initial_split(
+  df_select,
+  prop = 0.80,
+  # matching age distributions across train and test set
+  strata = "scan_age"
+)
+
+df_train <- training(df_split)
+df_validation <- testing(df_split)
+
+# report n training, features, testing
+df_train %>%
+  summarise(
+    n_training = n(),
+    p_features = ncol(df_train)
+  ) %>%
+  bind_cols(df_validation %>%
+              summarise(n_testing = n()))
+
+# Pre-processing setup ----------------------------------------------------
+
+# define (what we want to do)
+preprocess_recipe <- df_train %>%
+  # predict scan age by all brain features
+  recipe(scan_age ~ .) %>%
+  # remove near zero variance predictors
+  step_nzv(all_predictors()) %>%
+  prep() # where it all gets calculated
+
+preprocess_recipe
+
+
+# Apply pre-processing ----------------------------------------------------
+
+# juice() will work with training data, `bake()` to apply this to our test data
+
+# apply on train (gives processed value)
+df_train_prep <- juice(preprocess_recipe)
+
+# apply on validation
+df_validation_prep <- preprocess_recipe %>% bake(df_validation)
+
+# pre process on all of forbow
+frb_prep <- preprocess_recipe %>% bake(frb_select)
+
+
+# Cross-validation --------------------------------------------------------
+
+# 10 fold cv repeated 10 times
+set.seed(42)
+train_cv <- df_train_prep %>%
+  vfold_cv(
+    v = 10, 
+    repeats = 10, 
+    strata = scan_age
+  )

### 2022-03-16 train-fit-model.R

@@ -1,0 +1,78 @@
+# Extreme gradient boosting
+# Parallel processing -----------------------------------------------------
+
+library(doParallel)
+all_cores <- parallel::detectCores(logical = TRUE)
+# for 8 core 16 thread machine, good performance running more than 
+# physical but less than all logical
+registerDoParallel(cores = all_cores - 6) 
+
+# Model specification -----------------------------------------------------
+
+boost_mod <- boost_tree(
+  mode = "regression", 
+  trees = 1500, 
+  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
+  # randomness
+  sample_size = tune(), mtry = tune(), 
+  # step size
+  learn_rate = tune()
+  ) %>%
+  set_engine("xgboost", 
+             objective = "reg:squarederror")
+
+# Model parameters --------------------------------------------------------
+
+set.seed(42)
+
+# filled 6 dimensional tuning space
+xgboost_grid <- grid_latin_hypercube(
+  min_n(), tree_depth(), loss_reduction(),
+  sample_size = sample_prop(),
+  # has unknown, finalize with data to find max
+  finalize(mtry(), df_train_prep),
+  learn_rate(),
+  size = 500 
+)
+
+# Model tuning -----------------------------------------------------------
+
+tictoc::tic()
+
+set.seed(42)
+xgb_tuned_results <- tune_grid(
+  boost_mod,
+  scan_age ~ .,
+  resamples = train_cv,
+  grid = xgboost_grid,
+  metrics = metric_set(mae, rmse, rsq),
+  control = control_grid(verbose = FALSE,
+                         save_pred = TRUE)
+)
+
+tictoc::toc()
+beepr::beep(2)
+
+# Pick best model ---------------------------------------------------------
+
+xgb_tuned_results %>%
+  # want to minimize MAE
+  show_best("mae") 
+
+# select parsimonious params within one SE of best model
+best_xgb_params <- xgb_tuned_results %>%
+  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth) 
+
+
+# Fit and save best model -------------------------------------------------
+
+xgb_final_mod <- boost_mod %>%
+  finalize_model(best_xgb_params) %>%
+  fit(scan_age ~ .,
+      data = df_train_prep)
+
+
+# save mod
+saveRDS(xgb_final_mod, file = here::here(
+  "model", "xgboost_9to19_brain_age_mod.rds")
+  )

### 2022-03-18 Flux2022.Rmd
Created Rmd for Flux abstract - imports brainage, puberty, and cogntion. Cleans puberty scores, runs descriptives, splits into two sample, runs models, and tests model output.

---
title: "Flux2022"
author: "Lucy Whitmore"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(psych )
```

## Load Data
```{r, include=FALSE}
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))

smri_baseline <- smri_1%>% 
  filter(eventname=="baseline_year_1_arm_1")

```

```{r}
#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname== "baseline_year_1_arm_1") %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(src_subject_id,interview_age, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected) %>% 
  mutate_at(c(2:5), as.numeric)

#Existing Models
corrected_baseline_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/BaselineBrainAgeCorrected.Rda")
```

```{r, eval=F}
#future data import (if we find data)

#Sum scores (categorical)
youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

#Mean Scores & maybe all scores??
youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_mhy01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(src_subject_id, interview_age, sex, eventname,
         pds_y_ss_male_sum, pds_y_ss_male_sum_nm, pds_y_ss_male_mean, 
         pds_y_ss_male_mean_nm, pds_y_ss_male_category, pds_y_ss_male_cat_nm, 
         pds_y_ss_female_sum, pds_y_ss_female_sum_nm, pds_y_ss_female_mean, 
         pds_y_ss_female_mean_nm, pds_y_ss_female_category, pds_y_ss_female_cat_nm)

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_mhp01.txt") %>%
  filter(!collection_title=="collection_title")

```

## Format BrainAge
```{r}
corrected_baseline_brainage$src_subject_id <- smri_baseline$src_subject_id
```
## Format Puberty Data 
```{r}
#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

#Youth-report, removes everyone w/o complete data
youth_pds <- youth_pds %>% 
  mutate(youth_sum = case_when(pds_sex_y == "1" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         pds_sex_y == "2" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)

youth_pds_use <- youth_pds %>% 
  select(src_subject_id, pds_sex_y, youth_sum, youth_mean)

#Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA
parent_pds <- parent_pds %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5)

parent_pds_use <- parent_pds %>% 
  select(src_subject_id, sex, parent_sum, parent_mean)
```

## Make big dataframe
```{r}
#Combine dataframes
brainage_pds <- corrected_baseline_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_use) %>% 
  left_join(parent_pds_use) %>% 
  left_join(cognition)

#Split sample
smp_size <- floor(0.5 * nrow(brainage_pds))

## set the seed to make your partition reproducible
set.seed(123)
sample_ind <- sample(seq_len(nrow(brainage_pds)), size = smp_size)

sample_1 <- brainage_pds[sample_ind, ]
sample_2 <- brainage_pds[-sample_ind, ]
```

## Plot Descriptives
```{r}
ggplot(brainage_pds, aes(x=interview_age)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=corrected_gap)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=youth_mean)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=parent_mean)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=nihtbx_totalcomp_agecorrected)) + 
  geom_histogram()

```

## Split Sample Descriptives
```{r}
#Sample 1 Descriptives
describe(sample_1)

#Sample 2 Descriptives
describe(sample_2)

```

## Run Models
```{r}
# brainage ~ youth sum
baseline_puberty_y_sum_1 <- lm(corrected_gap ~ youth_sum, data = sample_1)
baseline_puberty_y_sum_2 <- lm(corrected_gap ~ youth_sum, data = sample_2)

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2)

# brainage ~ parent sum
baseline_puberty_p_sum_1 <- lm(corrected_gap ~ parent_sum, data = sample_1)
baseline_puberty_p_sum_2 <- lm(corrected_gap ~ parent_sum, data = sample_2)

# brainage ~ parent mean
baseline_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1)
baseline_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2)

baseline_puberty_p_mean_flipped_1 <- lm(parent_mean ~ corrected_gap, data = sample_1)
baseline_puberty_p_mean_flipped_2 <- lm(parent_mean ~ corrected_gap, data = sample_2)

#Cognition
baseline_cognition_1 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_1)
baseline_cognition_2 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_2)

```

## Model Summaries
```{r}
#summary(baseline_puberty_y_sum_1)

summary(baseline_puberty_y_mean_1)
summary(baseline_puberty_y_mean_2)

#summary(baseline_puberty_p_sum)

summary(baseline_puberty_p_mean_1)
summary(baseline_puberty_p_mean_2)

summary(baseline_puberty_p_mean_flipped_1)
summary(baseline_puberty_p_mean_flipped_2)

summary(baseline_cognition_1)
summary(baseline_cognition_2)
```

### 2022-04-05 Flux2022.Rmd
add gitdro submodule

@@
 library(tidyverse)
-library(psych )
+library(psych)
+library(sjPlot)
 ```
@@
 
+#Cognition fluid
+baseline_cognition_fluid_1 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_1)
+baseline_cognition_fluid_2 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_2)
+
+#Cognition crystallized
+baseline_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1)
+baseline_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2)
 ```
@@
 summary(baseline_cognition_2)
+
+summary(baseline_cognition_fluid_1)
+summary(baseline_cognition_fluid_2)
+
+summary(baseline_cognition_crystal_1)
+summary(baseline_cognition_crystal_2)
+
 ```
+
+Model Figures
+```{r}
+plot_model(baseline_puberty_y_mean_1, type="pred", terms = "youth_mean")
+
+plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean")
+
+plot_model(baseline_cognition_1, type="pred", terms = "nihtbx_totalcomp_agecorrected")
+plot_model(baseline_cognition_2, type="pred", terms = "nihtbx_totalcomp_agecorrected")
+
+```

### 2022-04-21 DataSetup.Rmd
Added ABCD baseline model (initial version)
Updated for baseline version, added talapas script

--- 2022-02-17 DataSetup.Rmd
@@
 
+smri_followup <- smri_1%>% 
+  filter(eventname=="2_year_follow_up_y_arm_1")
+
 #T1 intensity, might not need
@@
 
+#parent sum scores abcd_mhp01
+#youth sum scores abcd_mhy01
+
 #Cognition
@@
 ```{r, eval=F}
-renamed_smri <- smri_baseline %>% 
+renamed_smri_baseline <- smri_baseline %>% 
   rename(FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
@@
 #Create empty columns to match feature list
-renamed_smri[,"FS_BrainSeg_Vol"] <- NA
-renamed_smri[,"FS_BrainSeg_Vol_No_Vent"] <- NA
-renamed_smri[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
-renamed_smri[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
-renamed_smri[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
-renamed_smri[,"FS_Mask_Vol"] <- NA
-renamed_smri[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
-renamed_smri[,"FS_MaskVol_eTIV_Ratio"] <- NA
-renamed_smri[,"FS_L_Vessel_Vol"] <- NA
-renamed_smri[,"FS_L_ChoroidPlexus_Vol"] <- NA
-renamed_smri[,"FS_R_Vessel_Vol"] <- NA
-renamed_smri[,"FS_R_ChoroidPlexus_Vol"] <- NA
-renamed_smri[,"FS_OpticChiasm_Vol"] <- NA
-renamed_smri[,"FS_Total_GM_Vol"] <- NA
+renamed_smri_baseline[,"FS_BrainSeg_Vol"] <- NA
+renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent"] <- NA
+renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
+renamed_smri_baseline[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
+renamed_smri_baseline[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
+renamed_smri_baseline[,"FS_Mask_Vol"] <- NA
+renamed_smri_baseline[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
+renamed_smri_baseline[,"FS_MaskVol_eTIV_Ratio"] <- NA
+renamed_smri_baseline[,"FS_L_Vessel_Vol"] <- NA
+renamed_smri_baseline[,"FS_L_ChoroidPlexus_Vol"] <- NA
+renamed_smri_baseline[,"FS_R_Vessel_Vol"] <- NA
+renamed_smri_baseline[,"FS_R_ChoroidPlexus_Vol"] <- NA
+renamed_smri_baseline[,"FS_OpticChiasm_Vol"] <- NA
+renamed_smri_baseline[,"FS_Total_GM_Vol"] <- NA
 
@@
 #Convert to numeric
-renamed_smri[,6:279] <- sapply(renamed_smri[,6:279],as.numeric)
-renamed_smri <- renamed_smri %>% 
+renamed_smri_baseline[,6:279] <- sapply(renamed_smri_baseline[,6:279],as.numeric)
+renamed_smri_baseline <- renamed_smri_baseline %>% 
   mutate(interview_age = as.numeric(interview_age)/12)
 
+save(renamed_smri_baseline, file="renamed_smri_baseline.Rda")
+
+```
+
+Feature list template
+```{r, eval=F}
 #Template

### 2022-04-21 Flux2022.Rmd
Updated for baseline version, added talapas script

@@
 sample_2 <- brainage_pds[-sample_ind, ]
+
+#t test for age differences
+t.test(sample_1$interview_age, sample_2$interview_age)
+
+#Save samples
+
+save(sample_1, file="Sample1.Rda")
+save(sample_2, file = "Sample2.Rda")
+
 ```
@@
 ```{r}
-plot_model(baseline_puberty_y_mean_1, type="pred", terms = "youth_mean")
+plot_model(baseline_puberty_y_mean_1, type="pred", terms = "youth_mean", colors = "#67322e", title = "Predicted Gap") +
+  theme_sjplot()
 
-plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean")
+plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean", title = "Predicted Gap") +
+  theme_sjplot()
 
-plot_model(baseline_cognition_1, type="pred", terms = "nihtbx_totalcomp_agecorrected")
+plot_model(baseline_cognition_1, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
+  theme_sjplot()
+
 plot_model(baseline_cognition_2, type="pred", terms = "nihtbx_totalcomp_agecorrected")

### 2022-04-21 abcd_model_talapas.R

@@ -1,61 +1,14 @@
----
-title: "ABCD Model"
-output: html_document
----
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+library(dplyr)
+library(parallel)
+library(doParallel)
+library(tidymodels)
 
-packages <-  c("tidyverse",
-               "reshape2",
-               "nlme", "lme4",
-               "data.table", "psych",
-               "parallel","lubridate",
-               "ggpubr", "broom", 
-               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc", "rsample", "tidymodels")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-lapply(packages, library, character.only = TRUE)
-```
+#load data
+load("/projects/dcnlab/shared/brainage/df_train_baseline.Rds")
+load("/projects/dcnlab/shared/brainage/df_validation_baseline.Rds")
 
 
-Model Prep (Baseline)
-```{r, eval=F}
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-
-set.seed(42)
-df_split_baseline <- initial_split(
-  smri_1,
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "interview_age"
-)
-
-df_train_baseline <- training(df_split_baseline)
-df_validation_baseline <- testing(df_split_baseline)
-```
-
-
-Data Import/Set-Up
-```{r}
-load("Sample_1.Rds")
-load("Sample_2.Rds")
-load("renamed_smri_baseline.Rda")
-
-
-df_train_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_1$src_subject_id)
-df_validation_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_2$src_subject_id)
-
-save(df_train_baseline, file="df_train_baseline.Rds")
-save(df_validation_baseline, file="df_validation_baseline.Rds")
-
-```
-
-
-```{r}
 # report n training, features, testing
 df_train_baseline %>%
   summarise(
@@ -88,9 +41,6 @@ df_train_prep_baseline <- juice(preprocess_recipe_baseline)
 # apply on validation
 df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(df_validation_baseline)
[…]

### 2022-06-07 ABCDModel.Rmd
Added plots, consolidated analyses and figures into one file

@@
 
-Model Prep (Baseline)
+Data Import/Set-Up
 ```{r}
-# Model prep: split, preprocessing, CV ------------------------------------
+#load("Sample_1.Rds")
+#load("Sample_2.Rds")
+#load("renamed_smri_baseline.Rda")
 
-# Train / test split ------------------------------------------------------
+#df_train_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_1$src_subject_id)
+#df_validation_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_2$src_subject_id)
 
+#save(df_train_baseline, file="df_train_baseline.Rds")
+#save(df_validation_baseline, file="df_validation_baseline.Rds")
+
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/analysis_sample_baseline.Rds")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/training_sample_baseline.Rds")
+```
+
+Test Model
+```{r}
+abcd_brainage_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_brain_age_mod_5_9.rds")
+```
+
+Predict Brain Age (baseline)
+```{r}
+brain_age_df_baseline <-
+  abcd_brainage_model %>%
+  predict(new_data = analysis_sample_baseline) %>%
+  mutate(
+    # provide the chronological age at time of scan
+    truth = analysis_sample_baseline$interview_age
+  ) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
+
+```
+
+Predict Brain Age (follow-up)
+```{r}
+brain_age_df_followup <-
+  xgb_mod %>%
+  predict(new_data = renamed_smri_followup) %>%
+  mutate(
+    # provide the chronological age at time of scan
+    truth = renamed_smri_followup$interview_age
+  ) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
+```
+
+Evaluate accuracy
+```{r}
+# compute common performance metrics: mae, rsq, rmse
+brain_age_df_baseline %>%
+  metrics(truth = truth, estimate = .pred)
+
+brain_age_df_followup %>%
+  metrics(truth = truth, estimate = .pred)
+```
+
+Optional bias correction
+```{r}
+# bias prediction method, steps in comments,
+# previously determined values hardcoded.
+
+#get validation set
 set.seed(42)
 df_split_baseline <- initial_split(
-  smri_1,
+  training_sample_baseline, 
   prop = 0.80,
@@
 
-# report n training, features, testing
-df_train_baseline %>%
-  summarise(
-    n_training = n(),
-    p_features = ncol(df_train_baseline)
+describe(df_train_baseline)
+describe(df_validation_baseline)
+
+#think I need to now run model on validation set??
+abcd_brain_age_correction <-
+  abcd_brainage_model %>%
+  predict(new_data = df_validation_baseline) %>%
+  mutate(
+    # provide the chronological age at time of scan
+    truth = df_validation_baseline$interview_age
   ) %>%
-  bind_cols(df_validation_baseline %>%
-              summarise(n_testing = n()))
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
 
-# Pre-processing setup ----------------------------------------------------
+#get MAE for hold-out 
+abcd_brain_age_correction %>%
+  metrics(truth = truth, estimate = .pred)
 
-# define (what we want to do)
-preprocess_recipe_baseline <- df_train_baseline %>%
-  # predict scan age by all brain features
-  recipe(interview_age ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
+abcd_bias_mod_baseline <- lm(.pred ~ truth, data = abcd_brain_age_correction)
 
-preprocess_recipe_baseline
+# extract intercept and slope
+#abcd_bias_mod_baseline$coefficients[["(Intercept)"]]
+abcd_bias_intercept_baseline <- 7.917289
+#abcd_bias_mod_baseline$coefficients[["truth"]]
+abcd_bias_slope_baseline <-  0.2017998
 
-# Apply pre-processing ----------------------------------------------------
+# create bias correct data frame
+#Baseline
+brain_age_corrected_df_baseline <- brain_age_df_baseline %>%
+  mutate(
+    # corrected brain age prediction
+    corrected_pred =  (.pred - abcd_bias_intercept_baseline) / abcd_bias_slope_baseline
+  ) %>%
+  # corrected corrected brain age gap
+  mutate(corrected_gap = corrected_pred - truth)
 
-# juice() will work with training data, `bake()` to apply this to our test data
+#Follow-up
+brain_age_corrected_df_followup <- brain_age_df_followup %>%
+  mutate(
+    # corrected brain age prediction
+    corrected_pred =  (.pred - bias_intercept) / bias_slope
+  ) %>%
+  # corrected corrected brain age gap
+  mutate(corrected_gap = corrected_pred - truth)
 
-# apply on train (gives processed value)
-df_train_prep_baseline <- juice(preprocess_recipe_baseline)
+#mine
+brain_age_corrected_df_baseline %>%
+  metrics(truth = truth, estimate = .pred)
 
-# apply on validation
-df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(df_validation_baseline)
+brain_age_corrected_df_baseline %>%
+  metrics(truth = truth, estimate = corrected_pred)
 
-# pre process on all of forbow
-#frb_prep <- preprocess_recipe_baseline %>% bake(frb_select)
+brain_age_corrected_df_followup %>%
+  metrics(truth = truth, estimate = corrected_pred)
 
-# Cross-validation --------------------------------------------------------
+save(brain_age_df_baseline, file="ABCDBaselineBrainAge.Rda")
+save(brain_age_corrected_df_baseline, file="ABCDBaselineBrainAgeCorrected.Rda")
 
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv_baseline <- df_train_prep_baseline %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, #10
-    strata = interview_age
-  )
+save(brain_age_df_followup, file="ABCDFollowUpBrainAge.Rda")
+save(brain_age_corrected_df_followup, file="ABCDFollowUpBrainAgeCorrected.Rda")
 
 ```
-
-Train and Fit Model (Baseline)
-```{r}
-# Extreme gradient boosting
-# Parallel processing -----------------------------------------------------
-
-library(doParallel)
-all_cores <- parallel::detectCores(logical = TRUE)
-# for 8 core 16 thread machine, good performance running more than 
-# physical but less than all logical
-registerDoParallel(cores = all_cores - 6)
-
-# Model specification -----------------------------------------------------
-
-boost_mod <- boost_tree(
-  mode = "regression", 
-  trees = 1500, 
-  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
-  # randomness
-  sample_size = tune(), mtry = tune(), 
-  # step size
-  learn_rate = tune()
-  ) %>%
-  set_engine("xgboost", 
-             objective = "reg:squarederror")
-
-# Model parameters --------------------------------------------------------
-
-set.seed(42)
-
-# filled 6 dimensional tuning space
-xgboost_grid <- grid_latin_hypercube(
-  min_n(), tree_depth(), loss_reduction(),
-  sample_size = sample_prop(),
-  # has unknown, finalize with data to find max
-  finalize(mtry(), df_train_prep_baseline),
-  learn_rate(),
-  size = 500 
-)
-
-# Model tuning -----------------------------------------------------------
-
-tictoc::tic()
-
-set.seed(42)
-xgb_tuned_results <- tune_grid(
-  boost_mod,
-  interview_age ~ .,
-  resamples = train_cv_baseline,
-  grid = xgboost_grid,
-  metrics = metric_set(mae, rmse, rsq),
-  control = control_grid(verbose = FALSE,
-                         save_pred = TRUE)
-)
-
-tictoc::toc()
-beepr::beep(2)
-
-# Pick best model ---------------------------------------------------------
-
-xgb_tuned_results %>%
-  # want to minimize MAE
-  show_best("mae")
-
-# select parsimonious params within one SE of best model
-best_xgb_params <- xgb_tuned_results %>%
-  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)
-
-# Fit and save best model -------------------------------------------------
-
-xgb_final_mod <- boost_mod %>%
-  finalize_model(best_xgb_params) %>%
-  fit(interview_age ~ .,
-      data = df_train_prep)
-
-# save mod
-saveRDS(xgb_final_mod, file = here::here(
-  "model", "xgboost_abcd_brain_age_mod.rds")
-  )
-
-```

### 2022-06-07 ABCDModelMultiwave.Rmd
Added plots, consolidated analyses and figures into one file

---
title: "ABCDModelMultiwave"
author: "Lucy Whitmore"
date: "5/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Data Import/Set-Up
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_train_multiwave.Rds")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_test_multiwave.Rds")

```

Test Model
```{r}
abcd_brainage_multiwave_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_multiwave_brain_age_mod.rds")
```

Predict Brain Age (baseline)
```{r}
brain_age_df_multiwave <-
  abcd_brainage_multiwave_model %>%
  predict(new_data = abcd_test_multiwave) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = abcd_test_multiwave$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_multiwave %>%
  metrics(truth = truth, estimate = .pred)
```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.

#get validation set
set.seed(42)
df_split_multiwave <- initial_split(
  abcd_train_multiwave, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

df_validation_multiwave <- testing(df_split_multiwave)

#think I need to now run model on validation set??
abcd_multiwave_brain_age_correction <-
  abcd_brainage_multiwave_model %>%
  predict(new_data = df_validation_multiwave) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_multiwave$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

abcd_bias_mod_multiwave <- lm(.pred ~ truth, data = abcd_multiwave_brain_age_correction)

# extract intercept and slope
#abcd_bias_mod_multiwave$coefficients[["(Intercept)"]]
abcd_bias_intercept_multiwave <- 6.203475
#abcd_bias_mod_multiwave$coefficients[["truth"]]
abcd_bias_slope_multiwave <- 0.4300184

# create bias correct data frame
#Baseline
brain_age_corrected_df_multiwave <- brain_age_df_multiwave %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_multiwave) / abcd_bias_slope_multiwave
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df_multiwave %>%
  metrics(truth = truth, estimate = corrected_pred)

save(brain_age_df_multiwave, file="ABCDMultiwaveBrainAge.Rda")
save(brain_age_corrected_df_multiwave, file="ABCDMultiwaveBrainAgeCorrected.Rda")

```

### 2022-06-07 Flux2022.Rmd
Added plots, consolidated analyses and figures into one file

@@
 #Save samples
-
+save(brainage_pds, file="BrainagePDS.Rda")
 save(sample_1, file="Sample1.Rda")
@@
 ```{r}
-plot_model(baseline_puberty_y_mean_1, type="pred", terms = "youth_mean", colors = "#67322e", title = "Predicted Gap") +
-  theme_sjplot()
+# Youth-Report Puberty
+ggplot(sample_1, aes(x = youth_mean, y = corrected_gap)) +
+  stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
+  xlab("Pubertal Development (Youth-Report)") +
+  ylab("BrainAGE") +
+  ggtitle("Predicted Values") +
+  theme_minimal()
 
+# Parent-Report Puberty
+ggplot(sample_1, aes(x = parent_mean, y = corrected_gap)) +
+    stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
+  xlab("Pubertal Development (Parent-Report)") +
+  ylab("BrainAGE") +
+  ggtitle("Predicted Values") +
+  theme_minimal()
+
+#Cognition
+ggplot(sample_1, aes(x = nihtbx_totalcomp_agecorrected, y = corrected_gap)) +
+    stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
+  xlab("Age-Corrected Cognition Scores") +
+  ylab("BrainAGE") +
+  ggtitle("Predicted Values") +
+  theme_minimal()
+
+#Old Versions
 plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean", title = "Predicted Gap") +

### 2022-06-07 PaperAnalyses&Figures.Rmd
Added plots, consolidated analyses and figures into one file

---
title: "Paper Analyses"
author: "Lucy Whitmore"
date: "5/4/2022"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "ggpubr", 
               "apaTables", "MetBrewer",
               "ggseg", "vip", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```
Load Data
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")

#abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1FollowUp.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2FollowUp.Rda")

#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_baseline.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")

```

# Vlad's Model, Baseline Analyses

#Run Models
```{r}
# brainage ~ youth sum
#baseline_puberty_y_sum_1 <- lm(corrected_gap ~ youth_sum, data = sample_1)
#baseline_puberty_y_sum_2 <- lm(corrected_gap ~ youth_sum, data = sample_2)

sample_1_baseline_months <- sample_1_baseline %>% 
  mutate(interview_age_m = interview_age/12)

sample_2_baseline_months <- sample_2_baseline %>% 
  mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_1_baseline)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_2_baseline)

baseline_puberty_y_mean_1_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_y_mean_2_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_baseline_months)

# brainage ~ parent sum
#baseline_puberty_p_sum_1 <- lm(corrected_gap ~ parent_sum, data = sample_1)
#baseline_puberty_p_sum_2 <- lm(corrected_gap ~ parent_sum, data = sample_2)

# brainage ~ parent mean
baseline_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_baseline)
baseline_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_baseline)

baseline_puberty_p_mean_1_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_p_mean_2_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_2_baseline_months)

#Cognition
baseline_cognition_1 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_1_baseline)
baseline_cognition_2 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_2_baseline)

baseline_cognition_1_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_1_baseline_months)
baseline_cognition_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_2_baseline_months)
#Cognition fluid
#baseline_cognition_fluid_1 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_1)
#baseline_cognition_fluid_2 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_2)

#Cognition crystallized
#baseline_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1)
#baseline_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2)
```

## Model Summaries
```{r}
summary(baseline_puberty_y_mean_1)
summary(baseline_puberty_y_mean_2)

summary(baseline_puberty_p_mean_1)
summary(baseline_puberty_p_mean_2)

summary(baseline_cognition_1)
summary(baseline_cognition_2)
```
Model Summaries of Models w/ age as covariate
```{r}
summary(baseline_puberty_y_mean_1_m)
summary(baseline_puberty_y_mean_2_m)

summary(baseline_puberty_p_mean_1_m)
summary(baseline_puberty_p_mean_2_m)

summary(baseline_cognition_1_m)
summary(baseline_cognition__m2)

```

Figure 4A from vlad’s paper
Make categorical variables for puberty/cognition just for visualization purposes
Geom_density, fill will be different categories

Density plots
X axis brain age gap
y axis density
grouped/split/colored by categorical

3 plots: youth-report puberty, parent-report puberty, cognition
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_baseline <-  brainage_pds_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")

#Non-bias corrected version
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")

```

ggseg ones with weightings for the model
New model only

4 panels:
cortical map and bar graph
subcortical map and bar graph

stacked bar plot with right hemisphere followed by left hemisphere where applicable.
^note: how to decide top 10> top 10 by hemisphere and then add other hemisphere, or top 10 by bilateral calculation?

vip package
```{r}
# Calculate variables of importance
abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
abcd_vi

vip(abcd_model_5_9, num_features = 20)  +
  theme_light()

#Subset so that we don't have to rename all?

# Rename columns to fit with ggseg atlas (ugghhh)

brain_regions(dk)
brain_labels(dk)

brain_labels(aseg)

ggseg(abcd_vi)

#Actually plot
#Cortical
abcd_vi %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = Importance))
#Subcortical
abcd_vi %>%
  ggplot() +
  geom_brain(atlas = aseg, 
             position = position_brain(hemi ~ side),
             aes(fill = Importance))
```

Figure 2
Use subsample, otherwise it’ll look super funky

Scatterplot of scan age/chronological age vs predicted age
2 panels: 1 for validation, 1 for analysis sample
1 or 2 extra panels for analysis sample w/ Vlad's model
```{r}
#subsample data
#sample_n()

ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab= "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(brainage_pds_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regression line
   add.params = list(color = "blue", fill = "lightgray")
   )

ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"),
   xlab= "Scan/Chronological Age",
   ylab = "Corrected Gap"
   ) +
  scale_y_continuous(breaks = c(-5, -2.5, 0, 2.5, 5), 
                     labels = c(-5, -2.5, 0, 2.5, 5))

```

Checking Follow-up Brainage

Load Data
```{r}
corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")

#corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id

```

```{r}
ggpar((corrected_baseline_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Baseline")

ggpar((corrected_followup_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Follow-Up")

ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_followup_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Fit Metrics

#baseline uncorrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = corrected_pred)

#follow-up uncorrected
corrected_followup_brainage %>%
  metrics(truth = truth, estimate = .pred)
#follow-up corrected
corrected_followup_brainage %>%
  metrics(truth = truth, estimate = corrected_pred)

```

Check on talapas model output
```{r}
abcd_model_5_9 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_4_22 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod.rds")
```

Plots & Decriptives for ABCD Model
```{r}
describe(abcd_brainage_baseline)
#describe(abcd_brainage_baseline %>% filter(truth != 9))
#describe(abcd_brainage_baseline %>% filter(between(truth, 9, 9.2)))

#Fit Metrics
#baseline uncorrected
abcd_brainage_baseline %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

ggscatter(abcd_brainage_baseline, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age",
   ylab= "Corrected Gap"
   )

hist(abcd_brainage_baseline$truth)

```

### Follow-Up Analyses

```{r}
# brainage ~ youth mean
followup_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_followup)
followup_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_followup)

# brainage ~ parent mean
followup_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_followup)
followup_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_followup)

#Cognition crystallized
followup_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1_followup)
followup_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2_followup)
```

## Model Summaries
```{r}
summary(followup_puberty_y_mean_1)
summary(followup_puberty_y_mean_2)

summary(followup_puberty_p_mean_1)
summary(followup_puberty_p_mean_2)

summary(followup_cognition_crystal_1)
summary(followup_cognition_crystal_2)
```

#### ABCD Model - Baseline
```{r}
# brainage ~ youth sum
abcd_baseline_puberty_y_sum <- lm(corrected_gap ~ youth_sum, data = abcd_baseline)

# brainage ~ youth mean
abcd_baseline_puberty_y_mean <- lm(corrected_gap ~ youth_mean, data = abcd_baseline)

# brainage ~ parent sum
abcd_baseline_puberty_p_sum <- lm(corrected_gap ~ parent_sum, data = abcd_baseline)

# brainage ~ parent mean
abcd_baseline_puberty_p_mean <- lm(corrected_gap ~ parent_mean, data = abcd_baseline)

#Cognition
abcd_baseline_cognition <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=abcd_baseline)

#Cognition fluid
abcd_baseline_cognition_fluid <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=abcd_baseline)

#Cognition crystallized
abcd_baseline_cognition_crystal <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=abcd_baseline)
```

## Model Summaries
```{r}
summary(abcd_baseline_puberty_y_mean)

summary(abcd_baseline_puberty_y_sum)

summary(abcd_baseline_puberty_p_mean)

summary(abcd_baseline_puberty_p_sum)

summary(abcd_baseline_cognition_crystal)

summary(abcd_baseline_cognition_fluid)

summary(abcd_baseline_cognition)
```
ABCD Baseline Model
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_baseline_categorical_pds <- abcd_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
```
Examining correlation between vlad's model and abcd model
```{r}

abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id

model_compare <- abcd_brainage_baseline %>% 
  rename(.pred_abcd = .pred,
         truth_abcd = truth,
        gap_abcd = gap,
        corrected_pred_abcd = corrected_pred,
        corrected_gap_abcd = corrected_gap) %>% 
    left_join(sample_2)

ggscatter(model_compare, x = "corrected_gap", y = "corrected_gap_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(model_compare, x = "corrected_pred", y = "corrected_pred_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

cor.test(model_compare$truth, model_compare$corrected_gap_abcd)

cor.test(model_compare$truth, model_compare$corrected_gap)

#Correlation between same participants across models
cor(model_compare$corrected_gap, model_compare$corrected_gap_abcd)

```

###########
Combined Wave Model
###########
```{r}
#Fit Metrics
#baseline uncorrected
abcd_brainage_multiwave %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

ggscatter(corrected_multiwave_brainage, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age"
   )

```

### 2022-07-11 DataSetup.Rmd
Cleaned up data setup code and added new variables, added code for calculating difference scores

@@
 ---
-title: "DataSetup"
+title: "FYP Redo"
+author: "Lucy Whitmore"
+date: "5/19/2022"
 output: html_document
@@
 #Load list of included subjects
-struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
+struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
   filter(!collection_title=="collection_title") %>% 
-  select(1:12) %>% 
+  select(1:11) %>% 
   mutate(interview_age = as.numeric(interview_age),
-        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
-        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
-  filter(imgincl_t1w_include == 1)
+        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
+  filter(imgincl_t1w_include == 1) %>% 
+  filter(eventname=="baseline_year_1_arm_1")
+
+struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(1:11) %>% 
+  mutate(interview_age = as.numeric(interview_age),
+        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
+  filter(imgincl_t1w_include == 1) %>% 
+  filter(eventname=="2_year_follow_up_y_arm_1")
 
@@
 #Filter for volume & area measurements
-smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
+smri_baseline <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
+  filter(eventname=="baseline_year_1_arm_1") %>% 
   filter(!collection_title=="collection_title") %>% 
-  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))
+  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
+  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
+  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
 
-smri_baseline <- smri_1%>% 
-  filter(eventname=="baseline_year_1_arm_1")
+smri_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
+  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
+  filter(!collection_title=="collection_title") %>% 
+  filter(src_subject_id%in%struc_include_followup$src_subject_id) %>% 
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
+  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
+  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
 
-smri_followup <- smri_1%>% 
-  filter(eventname=="2_year_follow_up_y_arm_1")
-
-#T1 intensity, might not need
-smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  filter(src_subject_id%in%struc_include$src_subject_id)
-
-#Rename data to fit naming conventions
-#????
+#renamed_smri_baseline<-rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/renamed_smri_baseline.Rda")
 
@@
 
-Load Other Data
+Rename Columns Baseline
 ```{r}
-#Puberty
-youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-#parent sum scores abcd_mhp01
-#youth sum scores abcd_mhy01
-
-#Cognition
-cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)
-
-#Height/weight metrics
-physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
-  filter(!collection_title=="collection_title")  %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)
-
-#Sexual attraction and behavior
-gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-#Grade level, parent report
-grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
-```
-
-Template for renaming sMRI variables
-```{r, eval=F}
 renamed_smri_baseline <- smri_baseline %>% 
@@
 FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
-# no cerebellum cort vol?
+FS_R_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexrh,
 FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
@@
 
-#renamed_smri <- subset(renamed_smri, select = -c(269,279))
-
 #Convert to numeric
-renamed_smri_baseline[,6:279] <- sapply(renamed_smri_baseline[,6:279],as.numeric)
+renamed_smri_baseline[,6:ncol(renamed_smri_baseline)] <- sapply(renamed_smri_baseline[,6:ncol(renamed_smri_baseline)],as.numeric)
 renamed_smri_baseline <- renamed_smri_baseline %>% 
@@
 
+smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180] 
 ```
 
-Feature list template
-```{r, eval=F}
-#Template
-FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
-FS_BrainSeg_Vol
-FS_BrainSeg_Vol_No_Vent
-FS_BrainSeg_Vol_No_Vent_Surf
+Rename Columns Follow-up
+```{r}
+renamed_smri_followup <- smri_followup %>% 
+  rename(FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
+#FS_BrainSeg_Vol
+#FS_BrainSeg_Vol_No_Vent
+#FS_BrainSeg_Vol_No_Vent_Surf
 FS_LCort_GM_Vol = smri_vol_cdk_totallh,
@@
 FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
-FS_Total_GM_Vol #maybe just calculate?
+#FS_Total_GM_Vol #maybe just calculate?
 FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
-FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
-FS_SupraTentorial_No_Vent_Voxel_Count
-FS_Mask_Vol
-FS_BrainSegVol_eTIV_Ratio
-FS_MaskVol_eTIV_Ratio
+#FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
+#FS_SupraTentorial_No_Vent_Voxel_Count
+#FS_Mask_Vol
+#FS_BrainSegVol_eTIV_Ratio
+#FS_MaskVol_eTIV_Ratio
 FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
@@
 FS_L_VentDC_Vol = smri_vol_scs_vedclh,
-FS_L_Vessel_Vol
-FS_L_ChoroidPlexus_Vol
+#FS_L_Vessel_Vol
+#FS_L_ChoroidPlexus_Vol
 FS_R_LatVent_Vol = 	smri_vol_scs_ltventriclerh,
@@
 FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
-FS_R_Vessel_Vol
-FS_R_ChoroidPlexus_Vol
-FS_OpticChiasm_Vol
+#FS_R_Vessel_Vol
+#FS_R_ChoroidPlexus_Vol
+#FS_OpticChiasm_Vol
 FS_CC_Posterior_Vol = smri_vol_scs_ccps,
@@
 FS_R_Insula_GrayVol = smri_vol_cdk_insularh
+)
+
+#Create empty columns to match feature list
+renamed_smri_followup[,"FS_BrainSeg_Vol"] <- NA
+renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent"] <- NA
+renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
+renamed_smri_followup[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
+renamed_smri_followup[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
+renamed_smri_followup[,"FS_Mask_Vol"] <- NA
+renamed_smri_followup[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
+renamed_smri_followup[,"FS_MaskVol_eTIV_Ratio"] <- NA
+renamed_smri_followup[,"FS_L_Vessel_Vol"] <- NA
+renamed_smri_followup[,"FS_L_ChoroidPlexus_Vol"] <- NA
+renamed_smri_followup[,"FS_R_Vessel_Vol"] <- NA
+renamed_smri_followup[,"FS_R_ChoroidPlexus_Vol"] <- NA
+renamed_smri_followup[,"FS_OpticChiasm_Vol"] <- NA
+renamed_smri_followup[,"FS_Total_GM_Vol"] <- NA
+
+#Convert to numeric
+renamed_smri_followup[,6:ncol(renamed_smri_followup)] <- sapply(renamed_smri_followup[,6:ncol(renamed_smri_followup)],as.numeric)
+renamed_smri_followup <- renamed_smri_followup %>% 
+  mutate(interview_age = as.numeric(interview_age)/12)
+
+save(renamed_smri_followup, file="renamed_smri_followup.Rda")
+
+smri_abcd_model_followup <- renamed_smri_followup[, 1:180] 
+```
+
+Load Other Variables
+```{r}
+#Puberty
+#Raw scores, youth
+youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(5:38) %>% 
+  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
+  mutate_at(c(5:28), as.numeric)
+
+#Raw scores, parent
+parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  select(5:27) %>% 
+  select(-c(interview_date, pds_select_language___1)) %>% 
+  mutate_at(c(5:21), as.numeric)
+
+#Cognition Scores (Age-Corrected)
+cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_rawscore, nihtbx_totalcomp_cs, nihtbx_totalcomp_itmcnt) %>%
+  mutate_at(c(2, 4:9), as.numeric)
+
+#nihtbx_totalcomp_rawscore total cognition raw score
+#nihtbx_totalcomp_cs total cognition computed score
+#	nihtbx_totalcomp_itmcnt cognition total composite item count
+# might need to adjust for tasks completed?
+
+social <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  select(-c(subjectkey, interview_date, interview_age, gish2_device)) %>% 
+  mutate_at(c(4:16), as.numeric)
+
+#calculate gish score
+```
+
+## Format Puberty Data 
+```{r}
+#two or more missing values = exclude
+#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al.
+(2021).
+Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4).
+Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)
+
+### Youth-report, removes everyone w/o complete data
+youth_pds[youth_pds == 777] <- NA
+youth_pds[youth_pds == 999] <- NA
+
+# Multiwave
+youth_pds_multiwave <- youth_pds %>% 
+  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
+  mutate(youth_sum = case_when(sex == "M" ~ #male
+                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
+                         sex == "F" ~ #female
+                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
+  mutate(youth_mean = youth_sum/5)  %>% 
+  select(src_subject_id, sex, youth_sum, youth_mean)
+
+youth_pds_baseline <- youth_pds_multiwave %>% 
+  filter(eventname == "baseline_year_1_arm_1")
+
+youth_pds_followup <- youth_pds_multiwave %>% 
+  filter(eventname == "2_year_follow_up_y_arm_1")
+
+#### Parent-report, removes everyone w/o complete data
+parent_pds[parent_pds == 777] <- NA
+parent_pds[parent_pds == 999] <- NA
+
+#Baseline
+
+parent_pds_multiwave <- parent_pds %>% 
+    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
+    mutate(parent_sum = case_when(sex == "M" ~ #male
+                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
+                         sex == "F" ~ #female
+                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
+  mutate(parent_mean = parent_sum/5) %>% 
+  select(src_subject_id, sex, parent_sum, parent_mean)
+
+parent_pds_baseline <- parent_pds_multiwave %>% 
+    filter(eventname == "baseline_year_1_arm_1")
+
+parent_pds_followup <- parent_pds_multiwave%>% 
+    filter(eventname == "2_year_follow_up_y_arm_1")
+
+#### Cognition
+
+#Baseline
+cognition_baseline <- cognition %>% 
+  filter(eventname == "baseline_year_1_arm_1")
+
+#Follow-Up
+cognition_followup <- cognition %>% 
+  filter(eventname == "2_year_follow_up_y_arm_1")
+
+#Multiwave
+cognition_multiwave <- cognition %>% 
+  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1")
+```
+
+Add age to brainAGE estimates
+```{r}
+corrected_baseline_brainage <- brain_age_corrected_df_baseline
+corrected_followup_brainage <- brain_age_corrected_df_followup
+
+corrected_baseline_brainage$src_subject_id <- smri_baseline$src_subject_id
+
+corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
 
 ```
+
+Split Samples for Analysis
+```{r}
+
+brainage_pds_baseline <- corrected_baseline_brainage[, c(6,1:5)] %>% 
+  left_join(youth_pds_baseline) %>% 
+  left_join(parent_pds_baseline) %>% 
+  left_join(cognition_baseline)
+
+#Split sample
+smp_size_baseline <- floor(0.5 * nrow(brainage_pds_baseline))
+
+## set the seed to make your partition reproducible
+set.seed(123)
+sample_ind_baseline <- sample(seq_len(nrow(brainage_pds_baseline)), size = smp_size_baseline)
+
+sample_1_baseline <- brainage_pds_baseline[sample_ind_baseline, ]
+sample_2_baseline <- brainage_pds_baseline[-sample_ind_baseline, ]
+
+#t test for age differences
+t.test(sample_1_baseline$interview_age, sample_2_baseline$interview_age)
+
+#Save samples
+save(brainage_pds_baseline, file="BrainagePDSBaseline.Rda")
+save(sample_1_baseline, file="Sample1Baseline.Rda")
+save(sample_2_baseline, file = "Sample2Baseline.Rda")
+
+# Follow-Up
+brainage_pds_followup <- corrected_followup_brainage[, c(6,1:5)] %>% 
+  left_join(youth_pds_followup) %>% 
+  left_join(parent_pds_followup) %>% 
+  left_join(cognition_followup)
+
+#Split sample
+smp_size_followup <- floor(0.5 * nrow(brainage_pds_followup))
+
+## set the seed to make your partition reproducible
+set.seed(123)
+sample_ind_followup <- sample(seq_len(nrow(brainage_pds_followup)), size = smp_size_followup)
+
+sample_1_followup <- brainage_pds_followup[sample_ind_followup, ]
+sample_2_followup <- brainage_pds_followup[-sample_ind_followup, ]
+
+#t test for age differences
+t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)
+
+#Save samples
+save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
+save(sample_1_followup, file="Sample1FollowUp.Rda")
+save(sample_2_followup, file = "Sample2FollowUp.Rda")
+```
+
+#################
+Make Dataframe for ABCD Model Training & Testing
+################
+```{r}
+
+#smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180]
+
+training_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_1_baseline$src_subject_id) %>% 
+  select(-c(subjectkey, src_subject_id, sex, eventname))
+
+analysis_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_2_baseline$src_subject_id)
+
+save(training_sample_baseline, file="training_sample_baseline.Rds")
+save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")
+
+```
+
+#################
+Make Dataframe of sMRI features for combined wave model
+################
+```{r}
+
+#Pull 1800 total
+#900 from each
+#Pull 900 from baseline
+#filter those 900 out of the followup and then pull 900
+
+## set the seed to make your partition reproducible
+set.seed(42)
+#Pull 900 from baseline
+sample_ind_abcd_baseline <- sample(seq_len(nrow(smri_abcd_model_baseline)), size = 900)
+
+abcd_train_baseline <- smri_abcd_model_baseline[sample_ind_abcd_baseline, ]
+abcd_test_baseline <- smri_abcd_model_baseline[-sample_ind_abcd_baseline, ]
+
+#filter baseline participants out of follow-up
+smri_abcd_model_followup_filtered <- smri_abcd_model_followup %>% 
+  filter(!src_subject_id %in% abcd_train_baseline$src_subject_id)
+
+set.seed(42)
+#Pull 900 from followup
+sample_ind_abcd_followup <- sample(seq_len(nrow(smri_abcd_model_followup_filtered )), size = 900)
+
+abcd_train_followup <- smri_abcd_model_followup[sample_ind_abcd_followup, ]
+abcd_test_followup <- smri_abcd_model_followup[-sample_ind_abcd_followup, ]
+
+abcd_train_multiwave <- rbind(abcd_train_baseline, abcd_train_followup) %>% 
+  select(-c(subjectkey, src_subject_id, sex, eventname))
+save(abcd_train_multiwave, file="abcd_train_multiwave.Rds")
+
+abcd_test_multiwave <- rbind(abcd_test_baseline, abcd_test_followup)
+save(abcd_test_multiwave, file="abcd_test_multiwave.Rds")
+
+```
+
+##############
+Add Other Variable to Multiwave Test
+##############
+Add age to brainAGE estimates
+```{r}
+#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDMultiwaveBrainAgeCorrected.Rda")
+
+corrected_multiwave_brainage <- brain_age_corrected_df_multiwave
+
+corrected_multiwave_brainage$src_subject_id <- abcd_test_multiwave$src_subject_id
+
+#combine waves of youth pds
+
+#combine waves of parent pds
+
+#combine waves of cognition
+
+#change
+brainage_pds_multiwave <- corrected_multiwave_brainage[, c(6,1:5)] %>% 
+  left_join(youth_pds_multiwave) %>% 
+  left_join(parent_pds_multiwave) %>% 
+  left_join(cognition_multiwave)
+
+```

### 2022-07-11 DifferenceScores.Rmd
Cleaned up data setup code and added new variables, added code for calculating difference scores

---
title: "ChangeScores"
author: "Lucy Whitmore"
date: "7/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load Structural Data
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/renamed_smri_followup.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/renamed_smri_baseline.Rda")

smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180] 
smri_abcd_model_followup <- renamed_smri_followup[, 1:180]

```

Load data
```{r}
#Load smri_baseline

#Load smri_followup

smri_difference <- left_join(smri_abcd_model_baseline, smri_abcd_model_followup, by ="src_subject_id")
#could right join to reduce NAs

```

for loop - doesn't work very well
```{r}
difference1 <- c(1, 2, 3, 4, 5)
difference2 <- c(2, 4, 6, 8, 10)
difference3 <- c(1, 3, 5, 7, 9)
difference4 <- c(1, 2, 3, 4, 5)
difference5 <- c(5, 4, 3, 2, 1)
difference6 <- c(1, 2, 3, 5, 8)

difference_df <- data.frame(difference1, difference2, difference3, difference4, difference5, difference6)
difference_df

for (i in 1:(ncol(difference_df)/2)){
  print(i)
  print(i + ncol(difference_df)/2)
  difference_df[ncol(difference_df)+1] <-  difference_df[i + ncol(difference_df)/2] - difference_df[i]
  #print(dif)
}
```

other version
```{r}
#make empty dataframe that only has ids/numbers
#loop through columns, subtract as usual
#rbind (or append, or join, or whatever) the new difference score column onto the new data frame
#end up with dataframe of only ids and difference scores
diff_df <- data.frame(c("id1", "id2", "id3", "id4", "id5"))

for (i in 1:(ncol(difference_df)/2)){
  #print(i)
  #print(i + ncol(difference_df)/2)
  print((difference_df[i + ncol(difference_df)/2] - difference_df[i]))
  cbind(diff_df, (difference_df[i + ncol(difference_df)/2] - difference_df[i]))  
  #print(dif)
}
diff_df
```

Calculate difference scores format
# Stack Exchange - this works!
```{r}

difference1.x <- c(1, 2, 3, 4, 5)
difference2.x <- c(2, 4, 6, 8, 10)
difference3.x <- c(1, 3, 5, 7, 9)
difference1.y <- c(1, 2, 3, 4, 5)
difference2.y <- c(5, 4, 3, 2, 1)
difference3.y <- c(1, 2, 3, 5, 8)

difference_df_xy <- data.frame(difference1.x, difference2.x, difference3.x, difference1.y, difference2.y, difference3.y)
difference_df_xy

# Stack Exchange - this works!
difference_df_xy %>%
 mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))

#need to subtract, then divide by age gap
```

Calculate difference scores
```{r}
smri_difference_calculated <- smri_difference %>%
 select(-c(subjectkey.x,subjectkey.y, sex.x,sex.y, eventname.x,eventname.y)) %>% 
 mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))

#need to subtract, then divide by age gap
```

Test to make sure this worked
```{r}
smri_difference_calculated$interview_age.y[1] - smri_difference_calculated$interview_age.x[1]
smri_difference_calculated$interview_age.y_diff[1]

smri_difference_calculated$FS_L_Fusiform_Area.y[1] - smri_difference_calculated$FS_L_Fusiform_Area.x[1]
smri_difference_calculated$FS_L_Fusiform_Area.y_diff[1]

#test with NA
smri_difference_calculated$interview_age.y[3] - smri_difference_calculated$interview_age.x[3]
smri_difference_calculated$interview_age.y_diff[3]

smri_difference_calculated$FS_L_Fusiform_Area.y[3] - smri_difference_calculated$FS_L_Fusiform_Area.x[3]
smri_difference_calculated$FS_L_Fusiform_Area.y_diff[3]

#stopifnot(smri_difference_calculated$FS_L_Fusiform_Area.y[3] - smri_difference_calculated$FS_L_Fusiform_Area.x[3] == smri_difference_calculated$FS_L_Fusiform_Area.y_diff[3])

```

Prep for training
```{r}
# Select only difference columns
smri_difference_calculated_subset <- smri_difference_calculated %>% 
  select(src_subject_id, ends_with("diff")) 
  #divide all brain columns by interview age

smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2])

smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id

smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff

smri_difference_calculated_scaled <- smri_difference_calculated_scaled
select(src_subject_id, interview_age_diff, everything())

# Should we be predicting the change in age?
set.seed(42)
smri_difference_calculated <- smri_difference_calculated %>% 
#split for testing
select(-c(src_subject_id))

```

### 2022-07-11 PaperAnalyses&Figures.Rmd
Cleaned up data setup code and added new variables, added code for calculating difference scores

@@
 # brainage ~ youth mean
-baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_1_baseline)
-baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_2_baseline)
+baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_baseline)
+baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_baseline)
 
@@
 baseline_puberty_y_mean_2_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_baseline_months)
+
+#baseline_puberty_y_mean_1_pred <- lm(corrected_pred ~ youth_mean, data = sample_1_baseline)
+#baseline_puberty_y_mean_2_pred <- lm(corrected_pred ~ youth_mean, data = sample_2_baseline)
+
+#baseline_puberty_y_mean_1_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
+#baseline_puberty_y_mean_2_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_2_baseline_months)
 
@@
 
+#summary(baseline_puberty_y_mean_1_pred)
+#summary(baseline_puberty_y_mean_2_pred)
+
 summary(baseline_puberty_p_mean_1)
@@
 summary(baseline_cognition_1_m)
-summary(baseline_cognition__m2)
+summary(baseline_cognition_2_m)
+
+```
+
+Basic Prediction Plots
+```{r}
+
+df <- ggpredict(baseline_puberty_y_mean_1_m, terms = c("youth_mean","interview_age_m"))
+ggplot(df, aes(x, predicted)) + 
+    geom_line(aes(linetype=group, color=group)) +
+    geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=group), alpha=0.15) +
+    scale_linetype_manual(values = c("solid", "dashed", "dotted"))
+
+plot_model(baseline_puberty_y_mean_1_m, type="pred", terms = "youth_mean", title = "Predicted Gap", colors = "#cc5968") +
+  xlab("Youth-Report Puberty") +
+  ylab("Predicted Gap") +
+  theme_sjplot()
+
+plot_model(baseline_puberty_y_mean_2_m, type="pred", terms = "youth_mean", title = "Predicted Gap") +
+  xlab("Youth-Report Puberty") +
+  ylab("Predicted Gap") +
+  theme_sjplot()
+
+plot_model(baseline_puberty_p_mean_1_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
+  xlab("Parent-Report Puberty") +
+  ylab("Predicted Gap") +
+  theme_sjplot()
+
+plot_model(baseline_puberty_p_mean_2_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
+    xlab("Parent-Report Puberty") +
+    ylab("Predicted Gap") +
+  theme_sjplot()
+
+plot_model(baseline_cognition_1_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
+   xlab("Age-Corrected Cognition Score") +
+    ylab("Predicted Gap") +
+  theme_sjplot()
+
+plot_model(baseline_cognition_2_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
+    xlab("Age-Corrected Cognition Score") +
+    ylab("Predicted Gap") +
+  theme_sjplot()
 

### 2022-07-11 multiwave_model_talapas.R

@@ -1,4 +1,3 @@
-
 library(dplyr)
 library(parallel)
 library(doParallel)
@@ -7,44 +6,44 @@ library(tidymodels)
 #load data
 
 
-load("/projects/dcnlab/shared/brainage/training_sample_baseline.Rds")
+load("/projects/dcnlab/shared/brainage/abcd_train_multiwave.Rds")
 
 # Model prep: split, preprocessing, CV ------------------------------------
 
 # Train / test split ------------------------------------------------------
 set.seed(42)
-df_split_baseline <- initial_split(
-  training_sample_baseline, 
+df_split <- initial_split(
+  abcd_train_multiwave, 
   prop = 0.80,
   # matching age distributions across train and test set
   strata = "interview_age"
 )
 
-df_train_baseline <- training(df_split_baseline)
-df_validation_baseline <- testing(df_split_baseline)
+df_train <- training(df_split)
+df_validation <- testing(df_split)
 
 # report n training, features, testing
 #df_train_baseline %>% 
-df_train_baseline%>%
+df_train%>%
   summarise(
     n_training = n(),
-   p_features = ncol(df_train_baseline)
+    p_features = ncol(df_train)
   ) %>%
-  bind_cols(df_validation_baseline %>%
+  bind_cols(df_validation %>%
               summarise(n_testing = n()))
 
 # Pre-processing setup ----------------------------------------------------
 
 # define (what we want to do)
-preprocess_recipe_baseline <- df_train_baseline %>%
-#preprocess_recipe_baseline <- train_test %>%
+preprocess_recipe <- df_train %>%
+  #preprocess_recipe_baseline <- train_test %>%
   # predict scan age by all brain features
   recipe(interview_age ~ .) %>%
   # remove near zero variance predictors
   step_nzv(all_predictors()) %>%
   prep() # where it all gets calculated
 
-preprocess_recipe_baseline
+preprocess_recipe
 
 
 # Apply pre-processing ----------------------------------------------------
@@ -52,18 +51,17 @@ preprocess_recipe_baseline
 # juice() will work with training data, `bake()` to apply this to our test data
 
 # apply on train (gives processed value)
-df_train_prep_baseline <- juice(preprocess_recipe_baseline)
+df_train_prep <- juice(preprocess_recipe)
 
 # apply on validation
[…]

### 2022-11-07 DataSetup.Rmd
Updated code for difference score model, updated plottimng code.

@@
 ```{r}
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
+
 #Puberty
@@
   filter(!collection_title=="collection_title")  %>% 
-  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_rawscore, nihtbx_totalcomp_cs, nihtbx_totalcomp_itmcnt) %>%
-  mutate_at(c(2, 4:9), as.numeric)
+  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_uncorrected) %>% 
+  mutate_at(c(2, 4:7), as.numeric)
 
@@
 
-social <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
+gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
   filter(!collection_title=="collection_title")  %>% 
-  select(-c(subjectkey, interview_date, interview_age, gish2_device)) %>% 
-  mutate_at(c(4:16), as.numeric)
+  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
+   select(src_subject_id, sex, contains("gish2")) %>% 
+  select(-c(gish2_device, abcd_gish2y01_id))%>% 
+  mutate_at(c(3:15), as.numeric)
 
@@
   mutate(youth_mean = youth_sum/5)  %>% 
-  select(src_subject_id, sex, youth_sum, youth_mean)
+  select(src_subject_id, eventname, sex, youth_sum, youth_mean)
 
 youth_pds_baseline <- youth_pds_multiwave %>% 
-  filter(eventname == "baseline_year_1_arm_1")
+  filter(eventname == "baseline_year_1_arm_1") %>% 
+  select(-eventname)
 
 youth_pds_followup <- youth_pds_multiwave %>% 
-  filter(eventname == "2_year_follow_up_y_arm_1")
+  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
+  select(-eventname)
 
@@
 #Baseline
-
 parent_pds_multiwave <- parent_pds %>% 
@@
   mutate(parent_mean = parent_sum/5) %>% 
-  select(src_subject_id, sex, parent_sum, parent_mean)
+  select(src_subject_id, eventname, sex, parent_sum, parent_mean)
 
 parent_pds_baseline <- parent_pds_multiwave %>% 
-    filter(eventname == "baseline_year_1_arm_1")
+    filter(eventname == "baseline_year_1_arm_1") %>% 
+  select(-eventname)
 
 parent_pds_followup <- parent_pds_multiwave%>% 
-    filter(eventname == "2_year_follow_up_y_arm_1")
+    filter(eventname == "2_year_follow_up_y_arm_1")%>% 
+  select(-eventname)
 
@@
 cognition_baseline <- cognition %>% 
-  filter(eventname == "baseline_year_1_arm_1")
+  filter(eventname == "baseline_year_1_arm_1") %>% 
+  select(-eventname) %>%
+  mutate(cog_zscore = scale(nihtbx_totalcomp_uncorrected))
 
@@
 cognition_followup <- cognition %>% 
-  filter(eventname == "2_year_follow_up_y_arm_1")
+  filter(eventname == "2_year_follow_up_y_arm_1")%>% 
+  select(-eventname)
 
@@
   filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1")
+
+#Social/GISH
+gish[gish== 777] <- NA
+gish[gish == 999] <- NA
+
+# sum scores
+gish %>% 
+  mutate(gishSum = rowSums(across(c(3:15)), na.rm=T))
+
 ```
@@
 corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
-
 ```
@@
 save(sample_2_followup, file = "Sample2FollowUp.Rda")
+
+#Add Gish to follow-up
+```
+
+Attach maturity variables to ABCD model estimates
+```{r}
+abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
+abcd_brainage_baseline$src_subject_id <- sample_2_baseline$src_subject_id
+
+abcd_brainage_baseline <- abcd_brainage_baseline[, c(6,1:5)]
+
+abcd_baseline <- sample_2_baseline %>% 
+  select(1, 7:17)
+
+abcd_baseline <- left_join(abcd_baseline, abcd_brainage_baseline)
+
+save(abcd_baseline, file="ABCDBaselineSample.Rda")
+
+#double check that sample 2s remain consistent
+
+abcd_baseline_ids <- abcd_brainage_baseline$src_subject_id
+
+og_sample_2 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp_old/Sample2Baseline.Rda")
+og_sample_2_ids <- og_sample_2$src_subject_id
+
+id_compare<-data.frame(abcd_baseline_ids,og_sample_2_ids)
+sum(ifelse(id_compare$abcd_baseline_ids==id_compare$og_sample_2_ids,1,0))
+
+#they're the same!
+
 ```

### 2022-11-07 DifferenceScores.Rmd
Updated code for difference score model, updated plottimng code.

@@
 
-smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2])
+smri_difference_calculated_subset[, -(1:2)] <- sweep(smri_difference_calculated_subset[, -(1:2)], 1, smri_difference_calculated_subset[, 2], "/")
 
-smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id
+# testing
+#difference_df_scaled<- difference_df[,-(1:2)] %>% sapply(`/`, difference_df[,2])
 
-smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff
+#smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2]) 
+#try lapply instead
 
-smri_difference_calculated_scaled <- smri_difference_calculated_scaled
-select(src_subject_id, interview_age_diff, everything())
+#difference_df_scaled<- difference_df %>%  
+#  mutate(across(everything(), -c(difference1, difference2) ~ . / !! difference2))
+
+#difference_df_scaled<- difference_df %>% 
+ # mutate(across(!difference1 & !difference2), ~ . / !! difference2)
+
+#difference_df_scaled<- difference_df %>% 
+ # mutate(across(.cols = -c(difference1, difference2), . / difference2(., na.rm = FALSE)))
+
+#difference_df_scaled<- difference_df %>%
+#  mutate_at(vars(-difference1 -difference2),
+  #          ~ . / !! difference2
+#  )
+
+#difference_df_scale <- difference_df
+#difference_df_scale[, -(1:2)] <- sweep(difference_df_scale[, -(1:2)], 1, difference_df_scale[, 2], "/")
+
+####
+
+#smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id
+
+#smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff
 
 # Should we be predicting the change in age?
-set.seed(42)
-smri_difference_calculated <- smri_difference_calculated %>% 
-#split for testing
-select(-c(src_subject_id))
+
+#probably filter out people w/o age_diff scores
+smri_difference_calculated_filtered <- smri_difference_calculated_subset %>% 
+  filter(!is.na(interview_age.y_diff))
+
+#calculate midpoint age, use change scores to predict midpoint
+smri_difference_calculated_filtered <- smri_difference_calculated_filtered %>% 
+  left_join(smri_difference %>%  select(src_subject_id, interview_age.x, interview_age.y)) %>% 
+  rowwise() %>% 
+  mutate(age_midpoint = mean(c(interview_age.x, interview_age.y))) %>% 
+  ungroup()
+
+smri_difference_calculated_filtered <- smri_difference_calculated_filtered %>% 
+  ungroup()
+
+save(smri_difference_calculated_filtered, file="smri_difference_scores.Rda")
+
+# select just subject_id, brain columns, age midpoint
+
+#load("smri_difference_scores.Rda")
+
+#1000 people for training
+
+#Split sample
+smp_size_diff <- floor(0.13 * nrow(smri_difference_calculated_filtered)) #training sample of 1000 participants
+
+## set the seed to make your partition reproducible
+set.seed(123)
+sample_ind_diff <- sample(seq_len(nrow(smri_difference_calculated_filtered)), size = smp_size_diff)
+
+sample_train_diff <- smri_difference_calculated_filtered[sample_ind_diff, ]
+sample_test_diff <- smri_difference_calculated_filtered[-sample_ind_diff, ]
+
+# select just subject_id, brain columns, age midpoint
+sample_train_diff <- sample_train_diff %>% 
+select(-c(src_subject_id, interview_age.y_diff, interview_age.x, interview_age.y))
+
+save(sample_train_diff, file="sample_train_diff.Rda")
+save(sample_test_diff, file="sample_test_diff.Rda")
 

### 2022-11-07 ExisitingModel.Rmd
Updated file path to reflect submodule
Updated code for difference score model, updated plottimng code.

@@
 #Install XGboost & tidymodels
-packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
+#packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
 # You then install this version of the package using
-install.packages(packageUrl, repos = NULL, type = 'source')
+#install.packages(packageUrl, repos = NULL, type = 'source')
 
-install.packages("tidymodels")
+#install.packages("tidymodels")
 
@@
   file = here::here(
-    "xgboost_9to19_brain_age_mod.rds"))
+    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))
 ```
@@
 ```{r}
-brain_age_df <-
+brain_age_df_baseline <-
   xgb_mod %>%
-  predict(new_data = renamed_smri) %>%
+  predict(new_data = renamed_smri_baseline) %>%
   mutate(
     # provide the chronological age at time of scan
-    truth = renamed_smri$interview_age
+    truth = renamed_smri_baseline$interview_age
   ) %>%
@@
 # compute common performance metrics: mae, rsq, rmse
-brain_age_df %>%
+brain_age_df_baseline %>%
   metrics(truth = truth, estimate = .pred)
@@
 # create bias correct data frame
-brain_age_corrected_df <- brain_age_df %>%
+brain_age_corrected_df_baseline <- brain_age_df_baseline %>%
   mutate(
@@
 #mine
-brain_age_corrected_df %>%
+brain_age_corrected_df_baseline %>%
   metrics(truth = truth, estimate = corrected_pred)
@@
 
-save(brain_age_df, file="BaselineBrainAge.Rda")
-save(brain_age_corrected_df, file="BaselineBrainAgeCorrected.Rda")
+save(brain_age_df_baseline, file="BaselineBrainAge.Rda")
+save(brain_age_corrected_df_baseline, file="BaselineBrainAgeCorrected.Rda")
 

### 2022-11-07 PaperAnalyses&Figures.Rmd
Updated code for difference score model, updated plottimng code.

@@
 ```{r}
+# BrainAGE and Behavioral Data - Vlad's Model, Baseline
 load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")
 
-#abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
+# BrainAGE and Behavioral Data - Vlad's Model, Follow-Up
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSFollowUp.Rda")
 
-abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
+# BrainAGE and Behavioral Data - ABCD Model, Baseline
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineSample.Rda")
+
+abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
+
+#abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
 
@@
 
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
+
 load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
 load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")
+
+# Load Demographics
+demographics <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/pdem02.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, contains("demo_race"), demo_ethn_v2, demo_ethn2_v2, demo_comb_income_v2)
+
+  #join once by baseline IDs, then by follow-up (only recorded at one wave)
+
+demographics_baseline <- left_join(brainage_pds_baseline, demographics)
+
+demographics_followup <- left_join(brainage_pds_followup, demographics)
+
+demographics_baseline_abcdmodel <- left_join(abcd_baseline, demographics)
+
+demographics_long <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_lpds01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, eventname, demo_comb_income_v2_l)
+
+demographics_baseline <- left_join(demographics_baseline, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
+
+demographics_followup <- left_join(demographics_followup, demographics_long %>% filter(eventname == "2_year_follow_up_y_arm_1"))
+
+demographics_baseline_abcdmodel <- left_join(demographics_baseline_abcdmodel, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
 
@@
 baseline_cognition_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_2_baseline_months)
+
+#standard scores
+baseline_cognition_standard_1_m <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=sample_1_baseline_months)
+baseline_cognition_standard_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=sample_2_baseline_months)
+
+#z-scores
+baseline_cognition_zscore_1_m <- lm(corrected_gap ~ cog_zscore + interview_age_m, data=sample_1_baseline_months)
+baseline_cognition_zscore_2_m <- lm(corrected_gap ~ cog_zscore + interview_age_m, data=sample_2_baseline_months)
+
 #Cognition fluid
@@
 summary(baseline_cognition_2)
+
+summary(baseline_cognition_zscore_1_m)
+summary(baseline_cognition_zscore_2_m)
+
 ```
@@
 summary(baseline_cognition_2_m)
+
+summary(baseline_cognition_standard_1_m)
+summary(baseline_cognition_standard_2_m)
 
@@
 
+Color Palettes:
+- Darker Green for Model 1 Baseline
+#CBDFBD
+#BED7AC
+#A9CA91
+#88B668
+- Lighter green for Model 1 follow-uop
+#DEE7B1
+#D4E09B 
+#CAD982
+#BDCF63
+- Pink for Model 2 baseline
+#F5BBA3
+#F19C79
+#ED865A
+#A44A3F
+
+Color Palettes v2
+233D4D
+E76F51
+FCCA46
+A1C181
+619B8A
+
+Green for M1 Baseline
+CCDDBB
+A4C185  --
+A1C181
+7CA352
+526D37
+
+Blue for M1 Follow-Up
+
+Red/Yellow for m2 baseline
+
 3 plots: youth-report puberty, parent-report puberty, cognition
@@
          parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
-         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))
-
+         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
+         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
+         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
 #don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
@@
   filter(!is.na(youth_pds_cat)) %>% 
-  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))),  #,'#526D37'
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS") +
+  theme(text = element_text(size = 24)) 
+ggsave(filename="baseline_youth_pds_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
   filter(!is.na(parent_pds_cat)) %>% 
-ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS") +
+  theme(text = element_text(size = 24)) 
+ggsave(filename="baseline_parent_pds_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
   filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette = c('#7CA352',
+ '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Standardized Cognition")  +
+  theme(text = element_text(size = 24))
+
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cognition_cat)) %>% 
   ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
+
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cognition_cat_small)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition (.5 sd")
+
+#Cognition Quartiles
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#7CA352',
+ '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)")  +
+  theme(text = element_text(size = 24)) 
+ggsave(filename="baseline_cognition_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
+
+```
+
+Uncorrected Version
+```{r}
 
@@
   filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
-  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
+  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")
 
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")
 ```
 
-ggseg ones with weightings for the model
-New model only
+Demographics for Baseline Sample
+```{r}
+table(brainage_pds_baseline$sex)
 
-4 panels:
-cortical map and bar graph
-subcortical map and bar graph
+#white
+table(demographics_baseline$demo_race_a_p___10)
 
-stacked bar plot with right hemisphere followed by left hemisphere where applicable.
-^note: how to decide top 10> top 10 by hemisphere and then add other hemisphere, or top 10 by bilateral calculation?
+#black/african-american
+table(demographics_baseline$demo_race_a_p___11)
 
-vip package
-```{r}
-# Calculate variables of importance
-abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
-abcd_vi
+#native american
+table(demographics_baseline$demo_race_a_p___12)
 
-vip(abcd_model_5_9, num_features = 20)  +
-  theme_light()
+#alaska native
+table(demographics_baseline$demo_race_a_p___13)
 
-#Subset so that we don't have to rename all?
+#native hawaiian
+table(demographics_baseline$demo_race_a_p___14)
 
-# Rename columns to fit with ggseg atlas (ugghhh)
+#guamian
+table(demographics_baseline$demo_race_a_p___15)
 
-brain_regions(dk)
-brain_labels(dk)
+#samoan
+table(demographics_baseline$demo_race_a_p___16)
 
-brain_labels(aseg)
+#other pacific islander
+table(demographics_baseline$demo_race_a_p___17)
 
-ggseg(abcd_vi)
+#asian indian
+table(demographics_baseline$demo_race_a_p___18)
 
-#Actually plot
-#Cortical
-abcd_vi %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = Importance))
-#Subcortical
-abcd_vi %>%
-  ggplot() +
-  geom_brain(atlas = aseg, 
-             position = position_brain(hemi ~ side),
-             aes(fill = Importance))
+#chinese
+table(demographics_baseline$demo_race_a_p___19)
+
+#filipino
+table(demographics_baseline$demo_race_a_p___20)
+
+#japanese
+table(demographics_baseline$demo_race_a_p___21)
+
+#korean
+table(demographics_baseline$demo_race_a_p___22)
+
+#vietnamese
+table(demographics_baseline$demo_race_a_p___23)
+
+#other asian
+table(demographics_baseline$demo_race_a_p___24)
+
+#other race
+table(demographics_baseline$demo_race_a_p___25)
+
+#don't know
+table(demographics_baseline$demo_race_a_p___99)
+
+#hispanic or latino
+table(demographics_baseline$demo_ethn_v2)
+
+#family income
+table(demographics_baseline$demo_comb_income_v2)
+
+#Overall values
+#Non-hispanic white
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Non-hispanic Black
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Hispanic
+demographics_baseline %>% 
+  filter(demo_ethn_v2 == 1)
+
+#Asian
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Native American Alaska Native
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
+
+#Multiracial
+
+# mutate column: true if participant marked any asian category
+# mutate column: true if participant marked native american/alaska native
+# sum white, black, asian, native american/alaska native, and additional races
+
+demographics_baseline %>% 
+  mutate_at(c(18:33), as.numeric) %>% 
+  mutate(
+    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
+                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
+         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
+                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
+         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>%
+  filter(race_total > 1)
+
+#additional races
+#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.
+
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+
 ```
@@
 
-ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_pred",
-   color = "#00AFBB", # Points color, shape and size
-   add = "reg.line",  # Add regressin line
-   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
-   conf.int = TRUE, # Add confidence interval
-   xlab= "Scan/Chronological Age", ylab = "Predicted Age"
-   )
+ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#A1C181")+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  #geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 1 (Baseline)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="baseline_truth_corrected_pred_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
 ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_gap",
-   color = "#00AFBB", # Points color, shape and size
+   color = "#CBDFBD", # Points color, shape and size
    add = "reg.line",  # Add regressin line
@@
 
+ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#A1C181")+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 1 (Baseline)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="baseline_truth_corrected_gap_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
 ```
@@
 corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")
+
+load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/BrainagePDSFollowUp.Rda")
 
@@
 
-ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_pred",
-   color = "#00AFBB", # Points color, shape and size
-   add = "reg.line",  # Add regressin line
-   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
-   conf.int = TRUE # Add confidence interval
-   )
+# Actual vs Corrected Predictions, w/ perfect correlation line
 
-ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_gap",
-   color = "#00AFBB", # Points color, shape and size
-   add = "reg.line",  # Add regressin line
-   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
-   conf.int = TRUE # Add confidence interval
-   )
+ggplot(brainage_pds_followup, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#619B8A")+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 1 (Follow-Up)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="followup_truth_corrected_pred_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Age Bias - truth vs corrected gap
+ggplot(brainage_pds_followup, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#619B8A")+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 1 (Follow-Up)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="followup_truth_corrected_gap_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
 #follow-up uncorrected
-corrected_followup_brainage %>%
+#corrected_followup_brainage %>%
+#  metrics(truth = truth, estimate = .pred)
+#follow-up corrected
+#corrected_followup_brainage %>%
+#  metrics(truth = truth, estimate = corrected_pred)
+
+#follow-up uncorrected
+brain_age_corrected_df_followup %>%
   metrics(truth = truth, estimate = .pred)
 #follow-up corrected
-corrected_followup_brainage %>%
+brain_age_corrected_df_followup%>%
   metrics(truth = truth, estimate = corrected_pred)
@@
 
+### Follow-Up Analyses
+
+```{r}
+#Convert ages to months
+sample_1_followup_months <- sample_1_followup %>% 
+  mutate(interview_age_m = interview_age/12)
+
+sample_2_followup_months <- sample_2_followup %>% 
+  mutate(interview_age_m = interview_age/12)
+
+# brainage ~ youth mean
+followup_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_1_followup_months)
+followup_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_followup_months)
+
+# brainage ~ parent mean
+followup_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_1_followup_months)
+followup_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_2_followup_months)
+
+#Cognition crystallized
+followup_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1_followup)
+followup_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2_followup)
+
+# brainage ~ GISH2
+```
+
+## Model Summaries
+```{r}
+summary(followup_puberty_y_mean_1)
+summary(followup_puberty_y_mean_2)
+
+summary(followup_puberty_p_mean_1)
+summary(followup_puberty_p_mean_2)
+
+summary(followup_cognition_crystal_1)
+summary(followup_cognition_crystal_2)
+```
+
+# Density Plots for Follow-Up
+```{r}
+#Youth-Report Puberty
+#make categorical variable for puberty
+#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
+categorical_pds_followup <-  brainage_pds_followup %>% 
+  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))
+
+#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
+
+ggpar((categorical_pds_followup %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")  +
+  theme(text = element_text(size = 24))
+
+ggsave(filename="followup_youth_pds_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Parent-Report Puberty
+ggpar((categorical_pds_followup %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")  +
+  theme(text = element_text(size = 24))
+
+ggsave(filename="followup_parent_pds_mod1.png",
+                          width=7, height=4, units='in', dpi=300)
+```
+
+Uncorrected version
+```{r}
+#Non-bias corrected version
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")
+
+#Parent-Report Puberty
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")
+
+#Cognition
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
+  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")
+
+ggpar((categorical_pds_baseline %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")
+
+```
+
+Demographics for Follow-Up
+```{r}
+table(brainage_pds_followup$sex)
+
+#white
+table(demographics_followup$demo_race_a_p___10)
+
+#black/african-american
+table(demographics_followup$demo_race_a_p___11)
+
+#native american
+table(demographics_followup$demo_race_a_p___12)
+
+#alaska native
+table(demographics_followup$demo_race_a_p___13)
+
+#native hawaiian
+table(demographics_followup$demo_race_a_p___14)
+
+#guamian
+table(demographics_followup$demo_race_a_p___15)
+
+#samoan
+table(demographics_followup$demo_race_a_p___16)
+
+#other pacific islander
+table(demographics_followup$demo_race_a_p___17)
+
+#asian indian
+table(demographics_followup$demo_race_a_p___18)
+
+#chinese
+table(demographics_followup$demo_race_a_p___19)
+
+#filipino
+table(demographics_followup$demo_race_a_p___20)
+
+#japanese
+table(demographics_followup$demo_race_a_p___21)
+
+#korean
+table(demographics_followup$demo_race_a_p___22)
+
+#vietnamese
+table(demographics_followup$demo_race_a_p___23)
+
+#other asian
+table(demographics_followup$demo_race_a_p___24)
+
+#other race
+table(demographics_followup$demo_race_a_p___25)
+
+#don't know
+table(demographics_followup$demo_race_a_p___99)
+
+#hispanic or latino
+table(demographics_followup$demo_ethn_v2)
+
+#income
+table(demographics_followup$demo_comb_income_v2_l)
+
+#Overall values
+
+#Non-hispanic white
+demographics_followup %>% 
+  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Non-hispanic Black
+demographics_followup %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Hispanic
+demographics_followup %>% 
+  filter(demo_ethn_v2 == 1)
+
+#Asian
+demographics_followup %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Native American Alaska Native
+demographics_followup %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
+
+#Multiracial
+# mutate column: true if participant marked any asian category
+# mutate column: true if participant marked native american/alaska native
+# sum white, black, asian, native american/alaska native, and additional races
+
+demographics_followup %>% 
+  mutate_at(c(17:32), as.numeric) %>% 
+  mutate(
+    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
+                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
+         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
+                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
+         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>%
+  filter(race_total > 1)
+
+#additional races
+#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.
+
+demographics_followup %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+
+```
+
+##### ABCD Model #####
+
 Check on talapas model output
@@
 abcd_model_5_9 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")
+
+abcd_model_5_9 <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")
+
+abcd_model_multiwave <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_multiwave_brain_age_mod.rds")
+
+xgb.save(abcd_model_5_9, 'abcd_model_5_9.json')
+
+save(abcd_model_5_9, abcd_model_5_9.RData)
+
+xgb.load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")
+
+xgb.serialize(abcd_model_5_9)
+
+xgb.Booster.complete(abcd_model_5_9)
+
+config <- xgb.config(abcd_model_5_9)
+print(config)
+
+bst <- xgboost::xgb.Booster("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")
+
+abcd_model_59 <- readRDS("~/Desktop/xgboost_abcd_brain_age_mod_5_9.rds")
 
@@
 
-ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_pred",
-   color = "#00AFBB", # Points color, shape and size
-   add = "reg.line",  # Add regressin line
-   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
-   conf.int = TRUE, # Add confidence interval
-   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
-   )
+# Actual vs Corrected Predictions, w/ perfect correlation line
 
-ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_gap",
-   color = "#00AFBB", # Points color, shape and size
-   add = "reg.line",  # Add regressin line
-   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
-   conf.int = TRUE, # Add confidence interval
-   xlab = "Scan/Chronological Age",
-   ylab= "Corrected Gap"
-   )
+ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#E76F51")+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 2 (Baseline)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="baseline_truth_corrected_pred_mod2.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Age Bias - truth vs corrected gap
+ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#E76F51")+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 2 (Baseline)") +
+  theme_classic() + 
+  theme(text = element_text(size = 22))  
+ggsave(filename="baseline_truth_corrected_gap_mod2.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
 ```
+ggseg ones with weightings for the model
+New model only
 
-### Follow-Up Analyses
+4 panels:
+cortical map and bar graph
+subcortical map and bar graph
 
+stacked bar plot with right hemisphere followed by left hemisphere where applicable.
+^note: how to decide top 10> top 10 by hemisphere and then add other hemisphere, or top 10 by bilateral calculation?
+
+vip package
 ```{r}
-# brainage ~ youth mean
-followup_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_followup)
-followup_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_followup)
+# Calculate variables of importance
+abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
+abcd_vi
 
-# brainage ~ parent mean
-followup_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_followup)
-followup_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_followup)
+abcd_vi <- vi(abcd_model_59) #xgb_final_mod, xgboost_abcd_brain_age_mod
+abcd_vi
 
-#Cognition crystallized
-followup_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1_followup)
-followup_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2_followup)
-```
+xgboost::xgb.importance(model=abcd_model_5_9)
 
-## Model Summaries
-```{r}
-summary(followup_puberty_y_mean_1)
-summary(followup_puberty_y_mean_2)
+vip(abcd_model_5_9, num_features = 20)  +
+  theme_light()
 
-summary(followup_puberty_p_mean_1)
-summary(followup_puberty_p_mean_2)
+vip(abcd_model_59, num_features = 20)  +
+  theme_light()
 
-summary(followup_cognition_crystal_1)
-summary(followup_cognition_crystal_2)
+#Subset so that we don't have to rename all?
+
+# Rename columns to fit with ggseg atlas (ugghhh)
+
+brain_regions(dk)
+brain_labels(dk)
+
+brain_labels(aseg)
+
+ggseg(abcd_vi)
+
+#Actually plot
+#Cortical
+abcd_vi %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = Importance))
+#Subcortical
+abcd_vi %>%
+  ggplot() +
+  geom_brain(atlas = aseg, 
+             position = position_brain(hemi ~ side),
+             aes(fill = Importance))
 ```
@@
 ```{r}
+abcd_baseline_months <- abcd_baseline %>% 
+  mutate(interview_age_m = interview_age/12)
+
 # brainage ~ youth sum
-abcd_baseline_puberty_y_sum <- lm(corrected_gap ~ youth_sum, data = abcd_baseline)
+abcd_baseline_puberty_y_sum <- lm(corrected_gap ~ youth_sum + interview_age_m, data = abcd_baseline_months)
 
 # brainage ~ youth mean
-abcd_baseline_puberty_y_mean <- lm(corrected_gap ~ youth_mean, data = abcd_baseline)
+abcd_baseline_puberty_y_mean <- lm(corrected_gap ~ youth_mean+ interview_age_m, data = abcd_baseline_months)
+#abcd_baseline_puberty_y_mean_noage <- lm(corrected_gap ~ youth_mean, data = abcd_baseline_months)
 
 # brainage ~ parent sum
-abcd_baseline_puberty_p_sum <- lm(corrected_gap ~ parent_sum, data = abcd_baseline)
+abcd_baseline_puberty_p_sum <- lm(corrected_gap ~ parent_sum + interview_age_m, data = abcd_baseline_months)
 
 # brainage ~ parent mean
-abcd_baseline_puberty_p_mean <- lm(corrected_gap ~ parent_mean, data = abcd_baseline)
+abcd_baseline_puberty_p_mean <- lm(corrected_gap ~ parent_mean + interview_age_m, data = abcd_baseline_months)
+#abcd_baseline_puberty_p_mean_noage <- lm(corrected_gap ~ parent_mean, data = abcd_baseline_months)
 
 #Cognition
-abcd_baseline_cognition <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=abcd_baseline)
+abcd_baseline_cognition <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected+ interview_age_m, data=abcd_baseline_months)
+
+abcd_baseline_cognition_standard <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=abcd_baseline_months)
+
+abcd_baseline_cognition_standard_noage <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected, data=abcd_baseline_months)
 
 #Cognition fluid
-abcd_baseline_cognition_fluid <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=abcd_baseline)
+#abcd_baseline_cognition_fluid <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=abcd_baseline_months)
 
 #Cognition crystallized
-abcd_baseline_cognition_crystal <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=abcd_baseline)
+#abcd_baseline_cognition_crystal <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=abcd_baseline_months)
 ```
@@
 
-summary(abcd_baseline_puberty_y_sum)
+#summary(abcd_baseline_puberty_y_sum)
 
@@
 
-summary(abcd_baseline_puberty_p_sum)
+#summary(abcd_baseline_puberty_p_sum)
 
-summary(abcd_baseline_cognition_crystal)
+#summary(abcd_baseline_cognition_crystal)
 
-summary(abcd_baseline_cognition_fluid)
+#summary(abcd_baseline_cognition_fluid)
 
 summary(abcd_baseline_cognition)
+
+summary(abcd_baseline_cognition_standard)
+
 ```
+Model Plots
+```{r}
+#Cognition
+ggPredict(abcd_baseline_cognition_standard,se=TRUE)
+
+```
+
 ABCD Baseline Model
@@
          parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
-         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))
+         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
+         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
+         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
 
@@
 
+#Youth-report puberty
 ggpar((abcd_baseline_categorical_pds %>% 
   filter(!is.na(youth_pds_cat)) %>% 
-  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")  +
+  theme(text = element_text(size = 24))
+
+ggsave(filename="baseline_youth_pds_mod2.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
   filter(!is.na(parent_pds_cat)) %>% 
-ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")  +
+  theme(text = element_text(size = 24)) 
+ggsave(filename="baseline_parent_pds_mod2.png",
+                          width=7, height=4, units='in', dpi=300)
 
 ggpar((abcd_baseline_categorical_pds %>% 
-  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
-  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette=c('#FCCA46',
+ '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores") +
+  theme(text = element_text(size = 24))
+
+ggpar((abcd_baseline_categorical_pds %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores) .5 SD") +
+  theme(text = element_text(size = 24))
+
+#Cognition Quartiles
+ggpar((abcd_baseline_categorical_pds %>% 
+  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#FCCA46',
+ '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)")  +
+  theme(text = element_text(size = 24)) 
+ggsave(filename="baseline_cognition_mod2.png",
+                          width=7, height=4, units='in', dpi=300)
+
 ```
+
+Demographics for ABCD Baseline
+```{r}
+table(demographics_baseline_abcdmodel$sex)
+
+#white
+table(demographics_baseline_abcdmodel$demo_race_a_p___10)
+
+#black/african-american
+table(demographics_baseline_abcdmodel$demo_race_a_p___11)
+
+#native american
+table(demographics_baseline_abcdmodel$demo_race_a_p___12)
+
+#alaska native
+table(demographics_baseline_abcdmodel$demo_race_a_p___13)
+
+#native hawaiian
+table(demographics_baseline_abcdmodel$demo_race_a_p___14)
+
+#guamian
+table(demographics_baseline_abcdmodel$demo_race_a_p___15)
+
+#samoan
+table(demographics_baseline_abcdmodel$demo_race_a_p___16)
+
+#other pacific islander
+table(demographics_baseline_abcdmodel$demo_race_a_p___17)
+
+#asian indian
+table(demographics_baseline_abcdmodel$demo_race_a_p___18)
+
+#chinese
+table(demographics_baseline_abcdmodel$demo_race_a_p___19)
+
+#filipino
+table(demographics_baseline_abcdmodel$demo_race_a_p___20)
+
+#japanese
+table(demographics_baseline_abcdmodel$demo_race_a_p___21)
+
+#korean
+table(demographics_baseline_abcdmodel$demo_race_a_p___22)
+
+#vietnamese
+table(demographics_baseline_abcdmodel$demo_race_a_p___23)
+
+#other asian
+table(demographics_baseline_abcdmodel$demo_race_a_p___24)
+
+#other race
+table(demographics_baseline_abcdmodel$demo_race_a_p___25)
+
+#don't know
+table(demographics_baseline_abcdmodel$demo_race_a_p___99)
+
+#hispanic or latino
+table(demographics_baseline_abcdmodel$demo_ethn_v2)
+
+#income
+table(demographics_baseline_abcdmodel$demo_comb_income_v2)
+
+# calculate overall values
+#Overall values
+#Non-hispanic white
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Non-hispanic Black
+demographics_baseline_abcdmodel%>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Hispanic
+demographics_baseline_abcdmodel %>% 
+  filter(demo_ethn_v2 == 1)
+
+#Asian
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Native American Alaska Native
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
+
+#Multiracial
+#Multiracial
+
+# mutate column: true if participant marked any asian category
+# mutate column: true if participant marked native american/alaska native
+# sum white, black, asian, native american/alaska native, and additional races
+
+demographics_baseline_abcdmodel %>% 
+  mutate_at(c(18:33), as.numeric) %>% 
+  mutate(
+    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
+                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
+         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
+                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
+         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>%
+  filter(race_total > 1)
+
+#additional races
+#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.
+
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+
+```
+
 Examining correlation between vlad's model and abcd model

### 2022-12-05 DataSetupOld.Rmd
Updated plots, updated difference score model (save method)

--- 2022-04-21 DataSetup.Rmd
@@
   filter(src_subject_id%in%struc_include$src_subject_id) %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
+  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
+  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
 
@@
 #T1 intensity, might not need
-smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  filter(src_subject_id%in%struc_include$src_subject_id)
+#smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
+ # filter(!collection_title=="collection_title") %>% 
+ # filter(src_subject_id%in%struc_include$src_subject_id)
 
@@
 #????
+ogfeatures <- rio::import("~/Documents/FYP/feature_list.txt")
 
@@
 #Convert to numeric
-renamed_smri_baseline[,6:279] <- sapply(renamed_smri_baseline[,6:279],as.numeric)
+renamed_smri_baseline[,6:ncol(renamed_smri_baseline)] <- sapply(renamed_smri_baseline[,6:ncol(renamed_smri_baseline)],as.numeric)
 renamed_smri_baseline <- renamed_smri_baseline %>% 
@@
 save(renamed_smri_baseline, file="renamed_smri_baseline.Rda")
+
+smri_abcd_model <- renamed_smri_baseline[, 1:180] 
+```
+
+Rename sMRI for follow-up
+```{r}
+
+renamed_smri_followup <- smri_followup %>% 
+  rename(FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
+#FS_BrainSeg_Vol
+#FS_BrainSeg_Vol_No_Vent
+#FS_BrainSeg_Vol_No_Vent_Surf
+FS_LCort_GM_Vol = smri_vol_cdk_totallh,
+FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
+FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
+FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
+#FS_Total_GM_Vol #maybe just calculate?
+FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
+#FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
+#FS_SupraTentorial_No_Vent_Voxel_Count
+#FS_Mask_Vol
+#FS_BrainSegVol_eTIV_Ratio
+#FS_MaskVol_eTIV_Ratio
+FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
+FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
+FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
+FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
+FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
+FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
+FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
+FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
+FS_3rdVent_Vol = smri_vol_scs_3rdventricle,
+FS_4thVent_Vol = smri_vol_scs_4thventricle,
+FS_BrainStem_Vol = smri_vol_scs_bstem,
+FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
+FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
+FS_CSF_Vol = smri_vol_scs_csf,
+FS_L_AccumbensArea_Vol = smri_vol_scs_aal,
+FS_L_VentDC_Vol = smri_vol_scs_vedclh,
+#FS_L_Vessel_Vol
+#FS_L_ChoroidPlexus_Vol
+FS_R_LatVent_Vol = 	smri_vol_scs_ltventriclerh,
+FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
+FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
+FS_R_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexrh,
+FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
+FS_R_Caudate_Vol = smri_vol_scs_caudaterh,
+FS_R_Putamen_Vol = smri_vol_scs_putamenrh,
+FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh,
+FS_R_Hippo_Vol = smri_vol_scs_hpusrh,
+FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh,
+FS_R_AccumbensArea_Vol = 	smri_vol_scs_aar,
+FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
+#FS_R_Vessel_Vol
+#FS_R_ChoroidPlexus_Vol
+#FS_OpticChiasm_Vol
+FS_CC_Posterior_Vol = smri_vol_scs_ccps,
+FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps,
+FS_CC_Central_Vol = smri_vol_scs_ccct,
+FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat,
+FS_CC_Anterior_Vol = smri_vol_scs_ccat,
+FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh,
+FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh,
+FS_L_Entorhinal_Area = smri_area_cdk_ehinallh,
+FS_L_Fusiform_Area = smri_area_cdk_fusiformlh,
+FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh,
+FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh,
+FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh,
+FS_L_Lateraloccipital_Area = smri_area_cdk_locclh,
+FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh,
+FS_L_Lingual_Area = smri_area_cdk_linguallh,
+FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh,
+FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh,
+FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh,
+FS_L_Paracentral_Area = smri_area_cdk_paracnlh,
+FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh,
+FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh,
+FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh,
+FS_L_Pericalcarine_Area = smri_area_cdk_pericclh,
+FS_L_Postcentral_Area = smri_area_cdk_postcnlh,
+FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh,
+FS_L_Precentral_Area = smri_area_cdk_precnlh,
+FS_L_Precuneus_Area = smri_area_cdk_pclh,
+FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh,
+FS_L_Superiorparietal_Area = smri_area_cdk_supllh,
+FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh,
+FS_L_Supramarginal_Area = smri_area_cdk_smlh,
+FS_L_Frontalpole_Area = smri_area_cdk_frpolelh,
+FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh,
+FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh,
+FS_L_Insula_Area = smri_area_cdk_insulalh,
+FS_R_Bankssts_Area = smri_area_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh,
+FS_R_Cuneus_Area = smri_area_cdk_cuneusrh,
+FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh,
+FS_R_Fusiform_Area = smri_area_cdk_fusiformrh,
+FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh,
+FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh,
+FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh,
+FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh,
+FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh,
+FS_R_Lingual_Area = smri_area_cdk_lingualrh,
+FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh,
+FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh,
+FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh,
+FS_R_Paracentral_Area = smri_area_cdk_paracnrh,
+FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh,
+FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh,
+FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh,
+FS_R_Pericalcarine_Area = smri_area_cdk_periccrh,
+FS_R_Postcentral_Area = smri_area_cdk_postcnrh,
+FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh,
+FS_R_Precentral_Area = smri_area_cdk_precnrh,
+FS_R_Precuneus_Area = smri_area_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh,
+FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh,
+FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh,
+FS_R_Supramarginal_Area = smri_area_cdk_smrh,
+FS_R_Frontalpole_Area = smri_area_cdk_frpolerh,
+FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh,
+FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh,
+FS_R_Insula_Area = smri_area_cdk_insularh,
+#abcd technically doesn't say gray matter volume for the following varibles
+FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh,
+FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh,
+FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh,
+FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh,
+FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh,
+FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh,
+FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh,
+FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh,
+FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh,
+FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh,
+FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh,
+FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh,
+FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh,
+FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh,
+FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh,
+FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh,
+FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh,
+FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh,
+FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh,
+FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh,
+FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh,
+FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh,
+FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh,
+FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh,
+FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh,
+FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh,
+FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh,
+FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh,
+FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh,
+FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh,
+FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh,
+FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh,
+FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh,
+FS_L_Insula_GrayVol = smri_vol_cdk_insulalh,
+FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh,
+FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh,
+FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh,
+FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh,
+FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh,
+FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh,
+FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh,
+FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh,
+FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh,
+FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh,
+FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh,
+FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh,
+FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh,
+FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh,
+FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh,
+FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh,
+FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh,
+FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh,
+FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh,
+FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh,
+FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh,
+FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh,
+FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh,
+FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh,
+FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh,
+FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh,
+FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh,
+FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh,
+FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh,
+FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh,
+FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh,
+FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh,
+FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh,
+FS_R_Insula_GrayVol = smri_vol_cdk_insularh
+)
+
+#Create empty columns to match feature list
+renamed_smri_baseline[,"FS_BrainSeg_Vol"] <- NA
+renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent"] <- NA
+renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
+renamed_smri_baseline[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
+renamed_smri_baseline[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
+renamed_smri_baseline[,"FS_Mask_Vol"] <- NA
+renamed_smri_baseline[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
+renamed_smri_baseline[,"FS_MaskVol_eTIV_Ratio"] <- NA
+renamed_smri_baseline[,"FS_L_Vessel_Vol"] <- NA
+renamed_smri_baseline[,"FS_L_ChoroidPlexus_Vol"] <- NA
+renamed_smri_baseline[,"FS_R_Vessel_Vol"] <- NA
+renamed_smri_baseline[,"FS_R_ChoroidPlexus_Vol"] <- NA
+renamed_smri_baseline[,"FS_OpticChiasm_Vol"] <- NA
+renamed_smri_baseline[,"FS_Total_GM_Vol"] <- NA
+
+#renamed_smri <- subset(renamed_smri, select = -c(269,279))
+
+#Convert to numeric
+renamed_smri_baseline[,6:ncol(renamed_smri_baseline)] <- sapply(renamed_smri_baseline[,6:ncol(renamed_smri_baseline)],as.numeric)
+renamed_smri_baseline <- renamed_smri_baseline %>% 
+  mutate(interview_age = as.numeric(interview_age)/12)
+
+save(renamed_smri_baseline, file="renamed_smri_baseline.Rda")
+
+smri_abcd_model <- renamed_smri_baseline[, 1:180]
+
+```
+
+Split Samples
+```{r}
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1.Rda")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2.Rda")
+load("renamed_smri_baseline.Rda")
+
+training_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_1$src_subject_id) %>% 
+  select(-c(subjectkey, src_subject_id, sex, eventname))
+
+analysis_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_2$src_subject_id)
+
+save(training_sample_baseline, file="training_sample_baseline.Rds")
+save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")
+
+```
+
+Make dataframe for ML testing
+```{r, eval = F}
+train_test <- training_sample_baseline %>% 
+  #select(1:15) %>% 
+  slice_head(n=20)
+
+#validation_test <- df_validation_baseline %>% 
+  #select(1:15) %>% 
+#  slice_head(n=20)
+
+save(train_test, file="train_test.Rds")
+#save(validation_test, file="validation_test.Rds")
 
@@
 FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
+
 # no cerebellum cort vol?
@@
 ```
+
+```{r}
+#Puberty
+#Raw scores, youth
+youth_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
+  select(5:38) %>% 
+  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
+  mutate_at(c(5:28), as.numeric)
+
+#Raw scores, parent
+parent_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  filter(eventname== "2_year_follow_up_y_arm_1") %>% 
+  select(5:27) %>% 
+  select(-c(interview_date, pds_select_language___1)) %>% 
+  mutate_at(c(5:21), as.numeric)
+
+#Cognition Scores (Age-Corrected)
+cognition_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
+  select(src_subject_id,interview_age, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected) %>% 
+  mutate_at(c(2:5), as.numeric)
+
+#Existing Models
+corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")
+```
+
+## Format BrainAge
+```{r}
+corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
+```
+## Format Puberty Data 
+```{r}
+#two or more missing values = exclude
+#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al.
+(2021).
+Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4).
+Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)
+
+youth_pds_followup[youth_pds_followup == 777] <- NA
+youth_pds_followup[youth_pds_followup == 999] <- NA
+
+#Youth-report, removes everyone w/o complete data
+youth_pds_followup <- youth_pds_followup %>% 
+  mutate(youth_sum = case_when(sex == "M" ~ #male
+                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
+                         sex == "F" ~ #female
+                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
+  mutate(youth_mean = youth_sum/5)
+
+youth_pds_use_followup <- youth_pds_followup %>% 
+  select(src_subject_id, sex, youth_sum, youth_mean)
+
+#Parent-report, removes everyone w/o complete data
+parent_pds_followup[parent_pds_followup == 777] <- NA
+parent_pds_followup[parent_pds_followup == 999] <- NA
+parent_pds_followup <- parent_pds_followup %>% 
+    mutate(parent_sum = case_when(sex == "M" ~ #male
+                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
+                         sex == "F" ~ #female
+                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
+  mutate(parent_mean = parent_sum/5)
+
+parent_pds_use_followup <- parent_pds_followup %>% 
+  select(src_subject_id, sex, parent_sum, parent_mean)
+```
+
+## Make big dataframe
+```{r}
+brainage_pds_followup <- corrected_followup_brainage[, c(6,1:5)] %>% 
+  left_join(youth_pds_use_followup) %>% 
+  left_join(parent_pds_use_followup) %>% 
+  left_join(cognition_followup)
+
+#Split sample
+smp_size_followup <- floor(0.5 * nrow(brainage_pds_followup))
+
+## set the seed to make your partition reproducible
+set.seed(123)
+sample_ind_followup <- sample(seq_len(nrow(brainage_pds_followup)), size = smp_size_followup)
+
+sample_1_followup <- brainage_pds_followup[sample_ind_followup, ]
+sample_2_followup <- brainage_pds_followup[-sample_ind_followup, ]
+
+#t test for age differences
+t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)
+
+#Save samples
+save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
+save(sample_1_followup, file="Sample1FollowUp.Rda")
+save(sample_2_followup, file = "Sample2FollowUp.Rda")
+
+```
+
+Attach baseline pds/cognition data to sample 2 / abcd model brainage
+
+## Format BrainAge
+```{r}
+abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id
+
+abcd_brainage_baseline <- abcd_brainage_baseline[, c(6,1:5)]
+```
+
+```{r}
+abcd_baseline <- sample_2 %>% 
+  select(1, 7:16)
+
+abcd_baseline <- left_join(abcd_baseline, abcd_brainage_baseline)
+
+str(abcd_baseline)
+
+save(abcd_baseline, file="abcd_baseline.Rda")
+```
+
+#################
+Make Dataframe of sMRI features for combined wave model
+################
+```{r}
+
+```

### 2022-12-05 DiffScoreModel.Rmd
Updated plots, updated difference score model (save method)

---
title: "DiffScoreModel"
author: "Lucy Whitmore"
date: "7/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc", "rsample", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)

packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"

packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.5.2.1.tar.gz"

#packageUrl <- "~/Downloads/xgboost_1.0.0.2.tar.gz"

# You then install this version of the package using
install.packages(packageUrl, repos = NULL, type = 'source')
```

Load data
```{r}
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/sample_test_diff.Rda")
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/sample_train_diff.Rda")

load("~/Documents/FYP/sample_test_diff.Rda")
load("~/Documents/FYP/sample_train_diff.Rda")
```

Load model
```{r}
#abcd_brainage_diff_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_diff_brain_age_mod.rds")

abcd_brainage_diff_model <- xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod"))

#abcd_brainage_diff_model_complete <- xgb.Booster.complete(abcd_brainage_diff_model)

#abcd_brainage_multiwave_model <- readRDS("~/Documents/FYP/xgboost_abcd_multiwave_brain_age_mod.rds")

#xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod.rds"))

#xgboost::xgb.load(readRDS("xgboost_abcd_multiwave_brain_age_mod.rds"))

```

Predict Brain Age (baseline)
```{r}
sample_test_diff <- sample_test_diff %>% 
  select(-c(src_subject_id, interview_age.y_diff, interview_age.x, interview_age.y))

brain_age_df_diff <- abcd_brainage_diff_model %>%
  predict(new_data = sample_test_diff) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = sample_test_diff$age_midpoint
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_diff  %>%
  metrics(truth = truth, estimate = .pred)

```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.

#get validation set
set.seed(42)
df_split_diff <- initial_split(
  sample_train_diff, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

df_train_diff <- training(df_split_diff)
df_validation_diff <- testing(df_split_diff)

describe(df_train_diff)
describe(df_validation_diff)

#think I need to now run model on validation set??
abcd_brain_age_correction_diff <-
  abcd_brainage_diff_model %>%
  predict(new_data = df_validation_diff) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction_diff %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod_diff <- lm(.pred ~ truth, data = abcd_brain_age_correction_diff)

# extract intercept and slope
#abcd_bias_mod_diffcoefficients[["(Intercept)"]]
abcd_bias_intercept_diff <- 
#abcd_bias_mod_diff$coefficients[["truth"]]
abcd_bias_slope_diff <-

# create bias correct data frame
#Baseline
brain_age_corrected_df_diff <- brain_age_df_diff %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_diff) / abcd_bias_slope_diff
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = .pred)

brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = corrected_pred)

save(brain_age_df_diff, file="ABCDBaselineBrainAge.Rda")
save(brain_age_corrected_df_diff, file="ABCDBaselineBrainAgeCorrected.Rda")

```

### 2022-12-05 PaperAnalyses&Figures.Rmd
Updated plots, updated difference score model (save method)

@@
 ```{r}
-#Youth-Report Puberty
+
 #make categorical variable for puberty
@@
 
-ggpar((categorical_pds_baseline %>% 
+#Youth-Report Puberty
+youth_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
   filter(!is.na(youth_pds_cat)) %>% 
   ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))),  #,'#526D37'
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS") +
-  theme(text = element_text(size = 24)) 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend = 8) +
+  theme(text = element_text(size = 12)) 
 ggsave(filename="baseline_youth_pds_mod1.png",
@@
 #Parent-Report Puberty
-ggpar((categorical_pds_baseline %>% 
+parent_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
   filter(!is.na(parent_pds_cat)) %>% 
 ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS") +
-  theme(text = element_text(size = 24)) 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8) +
+  theme(text = element_text(size = 12)) 
 ggsave(filename="baseline_parent_pds_mod1.png",
@@
 #Cognition Quartiles
-ggpar((categorical_pds_baseline %>% 
+cog_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
   filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
   ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#7CA352',
- '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)")  +
-  theme(text = element_text(size = 24)) 
+ '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
 ggsave(filename="baseline_cognition_mod1.png",
                           width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_baseline_mod1_plot, parent_pds_baseline_mod1_plot, cog_baseline_mod1_plot, 
+          labels = c("A", "B", "C"),
+          ncol = 2, nrow = 2)
+
+ggsave(filename = "distributions_baseline_mod1.png",
+                          width=7, height=5, units='in', dpi=300)
 
@@
 
-ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_pred)) + 
+baseline_pred_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_pred)) + 
   geom_jitter(color="#A1C181")+
@@
   #geom_smooth(method=lm, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 1 (Baseline)") +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 14))  
 ggsave(filename="baseline_truth_corrected_pred_mod1.png",
@@
 
-ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_gap)) + 
+baseline_gap_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_gap)) + 
   geom_jitter(color="#A1C181")+
   geom_smooth(method=lm, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 1 (Baseline)") +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 14))  
 ggsave(filename="baseline_truth_corrected_gap_mod1.png",
                           width=7, height=4, units='in', dpi=300)
+
+ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+
+ggsave(filename = "pred_bias_baseline_mod1.png",
+                          width=10, height=5, units='in', dpi=300)
 ```
@@
 
-ggplot(brainage_pds_followup, aes(x=truth, y=corrected_pred)) + 
+followup_pred_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_pred)) + 
   geom_jitter(color="#619B8A")+
   geom_abline(a=0, b=1, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 1 (Follow-Up)") +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 14))  
 ggsave(filename="followup_truth_corrected_pred_mod1.png",
@@
 # Age Bias - truth vs corrected gap
-ggplot(brainage_pds_followup, aes(x=truth, y=corrected_gap)) + 
+followup_gap_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_gap)) + 
   geom_jitter(color="#619B8A")+
   geom_smooth(method=lm, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 1 (Follow-Up)") +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 14))  
 ggsave(filename="followup_truth_corrected_gap_mod1.png",
@@
 
+#Panel Plots
+ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, followup_pred_mod1_plot, followup_gap_mod1_plot, 
+          labels = c("A", "B", "C", "D"),
+          ncol = 2, nrow = 2)
+
+ggsave(filename = "pred_bias_mod1.png",
+                          width=10, height=10, units='in', dpi=300)
+
 #Fit Metrics
@@
 
-ggpar((categorical_pds_followup %>% 
+youth_pds_followup_mod1_plot <- ggpar((categorical_pds_followup %>% 
   filter(!is.na(youth_pds_cat)) %>% 
   ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")  +
-  theme(text = element_text(size = 24))
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
 
@@
 #Parent-Report Puberty
-ggpar((categorical_pds_followup %>% 
+parent_pds_followup_mod1_plot <-ggpar((categorical_pds_followup %>% 
   filter(!is.na(parent_pds_cat)) %>% 
 ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")  +
-  theme(text = element_text(size = 24))
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
 
@@
                           width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_followup_mod1_plot, parent_pds_followup_mod1_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+ggsave(filename = "distributions_followup_mod1.png",
+                          width=10, height=4, units='in', dpi=300)
+
 ```
@@
 
-ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_pred)) + 
+baseline_pred_mod2_plot <- ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_pred)) + 
   geom_jitter(color="#E76F51")+
   geom_abline(a=0, b=1, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Predicted Age", title = "Age v Pred Age - Model 2 (Baseline)") +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 12))  
 ggsave(filename="baseline_truth_corrected_pred_mod2.png",
@@
 # Age Bias - truth vs corrected gap
-ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_gap)) + 
+baseline_gap_mod2_plot <- ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_gap)) + 
   geom_jitter(color="#E76F51")+
   geom_smooth(method=lm, color='#233D4D') +
-  labs(x = "Scan/Chronological Age", y ="Corrected Gap", title = "Age Bias - Model 2 (Baseline)") +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
   theme_classic() + 
-  theme(text = element_text(size = 22))  
+  theme(text = element_text(size = 12))  
 ggsave(filename="baseline_truth_corrected_gap_mod2.png",
                           width=7, height=4, units='in', dpi=300)
+
+#Panel Plots
+ggarrange(baseline_pred_mod2_plot, baseline_gap_mod2_plot,
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+
+ggsave(filename = "pred_bias_mod2.png",
+                          width=10, height=5, units='in', dpi=300)
 
@@
 # Calculate variables of importance
-abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
-abcd_vi
-
-abcd_vi <- vi(abcd_model_59) #xgb_final_mod, xgboost_abcd_brain_age_mod
-abcd_vi
-
-xgboost::xgb.importance(model=abcd_model_5_9)
+#abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
+#abcd_vi
+
+#abcd_vi <- vi(abcd_model_59) #xgb_final_mod, xgboost_abcd_brain_age_mod
+#abcd_vi
+
+#xgboost::xgb.importance(model=abcd_brainage_model)
+#xgboost::xgb.importance(model=abcd_vi)
+
+abcd_brainage_model <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_brain_age_mod_5_9.rds")
+
+#This works!
+#abcd_vi <- vi(abcd_brainage_model) #xgb_final_mod, xgboost_abcd_brain_age_mod
+#abcd_vi
+#write_csv(abcd_vi, "abcd_vi.csv")
 
@@
 
+vip(abcd_brainage_model, num_features = 20)
+ggsave(filename="abcd_model_vip.png",
+                          width=7, height=4, units='in', dpi=300)
+
 #Subset so that we don't have to rename all?
@@
 
-brain_regions(dk)
+abcd_vi_renamed <- rio::import("~/Documents/FYP/abcd_vi_renamed.csv")
+abcd_vi_renamed_dk <- rio::import("~/Documents/FYP/abcd_vi_renamed_dk.csv")
+
+#check labels
 brain_labels(dk)
-
 brain_labels(aseg)
 
-ggseg(abcd_vi)
-
-#Actually plot
-#Cortical
-abcd_vi %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = Importance))
+#Cortical - Volume
+dk_vi_vol <- abcd_vi_renamed_dk %>%
+  filter(grepl('Vol', Variable)) %>% 
+  filter(label != "") %>% 
+  select(Importance,label)
+
+dk_vi_vol <- data.frame(dk_vi_vol)
+
+ggplot(dk_vi_vol) +
+  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
+  labs(title = "Cortical Volume")
+ggsave(filename="abcd_model_dk_vol_importance.png",
+                         width=8, height=4, units='in', dpi=300)
+
+#Cortical - Area
+dk_vi_area <- abcd_vi_renamed_dk %>%
+  filter(grepl('Area', Variable)) %>% 
+  filter(label != "") %>% 
+  select(Importance,label)
+
+dk_vi_area <- data.frame(dk_vi_area)
+
+ggplot(dk_vi_area) +
+  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
+  labs(title = "Cortical Area")
+ggsave(filename="abcd_model_dk_area_importance.png",
+                         width=8, height=4, units='in', dpi=300)
+
 #Subcortical
-abcd_vi %>%
-  ggplot() +
-  geom_brain(atlas = aseg, 
-             position = position_brain(hemi ~ side),
-             aes(fill = Importance))
+aseg_vi <- abcd_vi_renamed %>%
+  filter(label != "") %>% 
+  select(Importance,label)
+
+aseg_vi <- data.frame(aseg_vi)
+
+ggplot(aseg_vi) +
+  geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+ggsave(filename="abcd_model_aseg_importance_coronal.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggplot(aseg_vi) +
+  geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+ggsave(filename="abcd_model_aseg_importance_sagittal.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#troubleshooting names
+aseg_labs <- brain_labels(aseg) |> 
+  as.character()
+aseg_vi$label[!aseg_vi$label %in% aseg_labs]
+aseg_labs[!aseg_labs %in% aseg_vi$label]
+
+dk_labs <- brain_labels(dk) |> 
+  as.character()
+dk_vi_area$label[!dk_vi_area$label %in% dk_labs]
+dk_vi_vol$label[!dk_vi_vol$label %in% dk_labs]
+
+dk_labs[!dk_labs %in% dk_vi_area$label]
+dk_labs[!dk_labs %in% dk_vi_vol$label]
+
+#'#FCCA46','#F3B5A5', '#E76F51', '#B43718'
 ```
@@
 #Youth-report puberty
-ggpar((abcd_baseline_categorical_pds %>% 
+youth_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
   filter(!is.na(youth_pds_cat)) %>% 
   ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")  +
-  theme(text = element_text(size = 24))
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS",  font.legend = 8)  +
+  theme(text = element_text(size = 12))
 
@@
 #Parent-Report Puberty
-ggpar((abcd_baseline_categorical_pds %>% 
+parent_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
   filter(!is.na(parent_pds_cat)) %>% 
 ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
-  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")  +
-  theme(text = element_text(size = 24)) 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
 ggsave(filename="baseline_parent_pds_mod2.png",
@@
 #Cognition Quartiles
-ggpar((abcd_baseline_categorical_pds %>% 
+cog_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
   filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
   ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#FCCA46',
- '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)")  +
-  theme(text = element_text(size = 24)) 
+ '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
 ggsave(filename="baseline_cognition_mod2.png",
@@
 
+ggarrange(youth_pds_baseline_mod2_plot, parent_pds_baseline_mod2_plot, cog_baseline_mod2_plot, 
+          labels = c("A", "B", "C"),
+          ncol = 2, nrow = 2)
+ggsave(filename = "distributions_baseline_mod2.png",
+                          width=7, height=5, units='in', dpi=300)
 ```

### 2022-12-05 abcd_diff_model_talapas.R

@@ -1,0 +1,157 @@
+
+library(dplyr)
+library(parallel)
+library(doParallel)
+library(tidymodels)
+library(xgboost)
+
+#load data
+
+
+load("/projects/dcnlab/shared/brainage/sample_train_diff.Rda")
+
+# Model prep: split, preprocessing, CV ------------------------------------
+
+# Train / test split ------------------------------------------------------
+set.seed(42)
+df_split_diff <- initial_split(
+  sample_train_diff, 
+  prop = 0.80,
+  # matching age distributions across train and test set
+  strata = "age_midpoint"
+)
+
+df_train_diff <- training(df_split_diff)
+df_validation_diff <- testing(df_split_diff)
+
+# report n training, features, testing
+#df_train_baseline %>% 
+df_train_diff%>%
+  summarise(
+    n_training = n(),
+    p_features = ncol(df_train_diff)
+  ) %>%
+  bind_cols(df_validation_diff %>%
+              summarise(n_testing = n()))
+
+# Pre-processing setup ----------------------------------------------------
+
+# define (what we want to do)
+preprocess_recipe_diff <- df_train_diff %>%
+#preprocess_recipe_baseline <- train_test %>%
+  # predict scan age by all brain features
+  recipe(age_midpoint ~ .) %>%
+  # remove near zero variance predictors
+  step_nzv(all_predictors()) %>%
+  prep() # where it all gets calculated
+
+preprocess_recipe_diff
+
+
+# Apply pre-processing ----------------------------------------------------
+
+# juice() will work with training data, `bake()` to apply this to our test data
+
+# apply on train (gives processed value)
+df_train_prep_diff <- juice(preprocess_recipe_diff)
+
+# apply on validation
+df_validation_prep_diff <- preprocess_recipe_diff %>% bake(df_validation_diff)
+#df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(validation_test)
+
+
+# Cross-validation --------------------------------------------------------
+
+# 10 fold cv repeated 10 times
+set.seed(42)
+train_cv_diff <- df_train_prep_diff %>%
+  vfold_cv(
+    v = 10, 
+    repeats = 10, 
+    strata = age_midpoint
+  )
+
+
+#Train and Fit Model (Baseline)
+
+# Extreme gradient boosting
+# Parallel processing -----------------------------------------------------
+
[…]

### 2023-03-01 ABCDModel.Rmd
Cleaned up code, added comments & rearranged sections, updated plotting code

@@
 Predict Brain Age (follow-up)
-```{r}
+(for future analyses, no current follow-model)
+```{r, eval = F}
 brain_age_df_followup <-
@@
   metrics(truth = truth, estimate = .pred)
+```
 
-brain_age_df_followup %>%
-  metrics(truth = truth, estimate = .pred)
+```{r, eval=F}
+#brain_age_df_followup %>%
+#  metrics(truth = truth, estimate = .pred)
 ```
@@
 #Follow-up
-brain_age_corrected_df_followup <- brain_age_df_followup %>%
-  mutate(
+#brain_age_corrected_df_followup <- brain_age_df_followup %>%
+#  mutate(
     # corrected brain age prediction
-    corrected_pred =  (.pred - bias_intercept) / bias_slope
-  ) %>%
+#    corrected_pred =  (.pred - bias_intercept) / bias_slope
+#  ) %>%
   # corrected corrected brain age gap
-  mutate(corrected_gap = corrected_pred - truth)
+#  mutate(corrected_gap = corrected_pred - truth)
 
@@
 
-brain_age_corrected_df_followup %>%
-  metrics(truth = truth, estimate = corrected_pred)
+#brain_age_corrected_df_followup %>%
+#  metrics(truth = truth, estimate = corrected_pred)
 
@@
 
-save(brain_age_df_followup, file="ABCDFollowUpBrainAge.Rda")
-save(brain_age_corrected_df_followup, file="ABCDFollowUpBrainAgeCorrected.Rda")
+#save(brain_age_df_followup, file="ABCDFollowUpBrainAge.Rda")
+#save(brain_age_corrected_df_followup, file="ABCDFollowUpBrainAgeCorrected.Rda")
 

### 2023-03-01 DataSetup.Rmd
Cleaned up code, added comments & rearranged sections, updated plotting code

@@
 ```{r}
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
-
 #Puberty
@@
 
-gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
-  filter(!collection_title=="collection_title")  %>% 
-  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
-   select(src_subject_id, sex, contains("gish2")) %>% 
-  select(-c(gish2_device, abcd_gish2y01_id))%>% 
-  mutate_at(c(3:15), as.numeric)
+#gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
+#  filter(!collection_title=="collection_title")  %>% 
+ # filter(eventname == "2_year_follow_up_y_arm_1") %>% 
+#   select(src_subject_id, sex, contains("gish2")) %>% 
+ # select(-c(gish2_device, abcd_gish2y01_id))%>% 
+ # mutate_at(c(3:15), as.numeric)
 
@@
 ```
+**At this point, switch to the ExistingModel.Rmd file, where you can use renamed_smri_baseline and renamed_smri_followup to make predictions.**
 
-Add age to brainAGE estimates
+Add age to brainAGE estimates - Exisiting Model
 ```{r}
+#Don't need to run the load lines if you just ran ExistingModel.Rmd, these will already be loaded
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
+
+#make renamed copies of predictions
 corrected_baseline_brainage <- brain_age_corrected_df_baseline
@@
 
+#add subject ids to predictions
 corrected_baseline_brainage$src_subject_id <- smri_baseline$src_subject_id
@@
 
-#Add Gish to follow-up
+#Add Gish to follow-up at a later date
 ```
+
+#################
+Make Dataframe for ABCD Model Training & Testing
+################
+```{r}
+
+#can run the line below if you need to reload smri_abcd_model_baseline
+#smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180]
+
+training_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_1_baseline$src_subject_id) %>% 
+  select(-c(subjectkey, src_subject_id, sex, eventname))
+
+analysis_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_2_baseline$src_subject_id)
+
+save(training_sample_baseline, file="training_sample_baseline.Rds")
+save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")
+
+```
+
+**Switch over to ABCDModel.Rmd to get predictions from ABCD-trained model**
 
@@
 ```{r}
+#if you just ran ABCDModel.Rmd, comment out the next line
 abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
@@
 #double check that sample 2s remain consistent
-
 abcd_baseline_ids <- abcd_brainage_baseline$src_subject_id
@@
 
-#################
-Make Dataframe for ABCD Model Training & Testing
-################
-```{r}
-
-#smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180]
-
-training_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_1_baseline$src_subject_id) %>% 
-  select(-c(subjectkey, src_subject_id, sex, eventname))
-
-analysis_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_2_baseline$src_subject_id)
-
-save(training_sample_baseline, file="training_sample_baseline.Rds")
-save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")
-
-```
+**Below section is for future analyses**
 

### 2023-03-01 PaperAnalyses&Figures.Rmd
Cleaned up code, added comments & rearranged sections, updated plotting code

@@
 
+# ABCD Model Baseline BrainAGE estimates
 abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
@@
 #abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
+
+#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_baseline.Rda")
+#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
+
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
+
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")
 
@@
 load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2FollowUp.Rda")
-
-#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_baseline.Rda")
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
-
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")
-
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")
 
@@
   #join once by baseline IDs, then by follow-up (only recorded at one wave)
-
 demographics_baseline <- left_join(brainage_pds_baseline, demographics)
@@
 demographics_baseline_abcdmodel <- left_join(demographics_baseline_abcdmodel, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
-
-```
-
-# Vlad's Model, Baseline Analyses
-
-#Run Models
+```
+
+**Vlad's Model, Baseline Analyses**
+Run Models
 ```{r}
@@
 
-## Model Summaries
+Model Summaries
 ```{r}
@@
 ```
+
 Model Summaries of Models w/ age as covariate
@@
 Figure 2
-Use subsample, otherwise it’ll look super funky
-
 Scatterplot of scan age/chronological age vs predicted age
@@
 baseline_pred_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_pred)) + 
-  geom_jitter(color="#A1C181")+
+  geom_jitter(color="#A1C181", alpha=.2) +
   geom_abline(a=0, b=1, color='#233D4D') +
@@
 baseline_gap_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_gap)) + 
-  geom_jitter(color="#A1C181")+
+  geom_jitter(color="#A1C181", alpha=.2)+
   geom_smooth(method=lm, color='#233D4D') +
@@
 
+Plots
 ```{r}
@@
 followup_pred_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_pred)) + 
-  geom_jitter(color="#619B8A")+
+  geom_jitter(color="#619B8A", alpha=.2)+
   geom_abline(a=0, b=1, color='#233D4D') +
@@
 followup_gap_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_gap)) + 
-  geom_jitter(color="#619B8A")+
+  geom_jitter(color="#619B8A", alpha=.2)+
   geom_smooth(method=lm, color='#233D4D') +
@@
                           width=10, height=10, units='in', dpi=300)
-
-#Fit Metrics
-
+```
+
+Fit Metrics
+```{r}
 #baseline uncorrected
@@
 
-```
-
-### Follow-Up Analyses
+# Correlation between gap and age
+cor(brainage_pds_baseline$truth, brainage_pds_baseline$corrected_gap)
+cor(brainage_pds_followup$truth, brainage_pds_followup$corrected_gap)
+```
+
+**Follow-Up Analyses**
 
@@
 
-##### ABCD Model #####
+**ABCD Model**
 
@@
 
-Plots & Decriptives for ABCD Model
+Plots & Descriptives for ABCD Model
 ```{r}
@@
 
+# Check age v gap
+cor(abcd_baseline$truth, abcd_baseline$corrected_gap)
+
+# Plots
 ggscatter(abcd_brainage_baseline, x = "truth", y = ".pred",
@@
 
-baseline_pred_mod2_plot <- ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_pred)) + 
-  geom_jitter(color="#E76F51")+
+baseline_pred_mod2_plot <- ggplot(abcd_baseline, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#E76F51", alpha=.25)+
   geom_abline(a=0, b=1, color='#233D4D') +
@@
 # Age Bias - truth vs corrected gap
-baseline_gap_mod2_plot <- ggplot(abcd_brainage_baseline, aes(x=truth, y=corrected_gap)) + 
-  geom_jitter(color="#E76F51")+
+baseline_gap_mod2_plot <- ggplot(abcd_baseline, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#E76F51", alpha=.25)+
   geom_smooth(method=lm, color='#233D4D') +
@@
 
-#### ABCD Model - Baseline
+ABCD Model - Baseline
 ```{r}
@@
 
-#summary(abcd_baseline_puberty_y_sum)
-
 summary(abcd_baseline_puberty_p_mean)
 
-#summary(abcd_baseline_puberty_p_sum)
-
-#summary(abcd_baseline_cognition_crystal)
-
-#summary(abcd_baseline_cognition_fluid)
-
 summary(abcd_baseline_cognition)
@@
 summary(abcd_baseline_cognition_standard)
-
 ```
@@
 
-ABCD Baseline Model
+ABCD Baseline Model - Density Plots
 ```{r}
@@
 
-###########
-Combined Wave Model
-###########
+Checking internal consistency of puberty means
+```{r}
+#Puberty
+#Raw scores, youth
+youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(5:38) %>% 
+  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
+  mutate_at(c(5:28), as.numeric)
+
+#Raw scores, parent
+parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
+  filter(!collection_title=="collection_title")  %>% 
+  select(5:27) %>% 
+  select(-c(interview_date, pds_select_language___1)) %>% 
+  mutate_at(c(5:21), as.numeric)
+
+#two or more missing values = exclude
+#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al.
+(2021).
+Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4).
+Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)
+
+### Youth-report, removes everyone w/o complete data
+youth_pds[youth_pds == 777] <- NA
+youth_pds[youth_pds == 999] <- NA
+
+#### Parent-report, removes everyone w/o complete data
+parent_pds[parent_pds == 777] <- NA
+parent_pds[parent_pds == 999] <- NA
+
+# Calculting internal consistency
+# Use alpha output
+
+# Baseline, youth-report, male
+youth_pds %>% 
+  #filter for baseline wave, male, has BrainAGE estimate
+  filter(eventname == "baseline_year_1_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_m4_y","pds_m5_y")) %>% 
+  psych::omega(nfactors=3)
+
+# Baseline, youth-report, female
+youth_pds %>% 
+  #filter for baseline wave, female, has BrainAGE estimate
+  filter(eventname == "baseline_year_1_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_f4_2_y","pds_f5_y")) %>% 
+  psych::omega(nfactors=3)
+
+# Baseline, parent-report, male                           
+parent_pds %>% 
+  #filter for baseline wave, male, has BrainAGE estimate
+  filter(eventname == "baseline_year_1_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m4_p", "pds_m5_p")) %>% 
+  psych::omega(nfactors=3)
+
+# Baseline, parent-report, female                           
+parent_pds %>% 
+  #filter for baseline wave, female, has BrainAGE estimate
+  filter(eventname == "baseline_year_1_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")) %>% 
+  psych::omega(nfactors=3)
+
+# Follow-up, youth-report, male
+youth_pds %>% 
+  #filter for baseline wave, male, has BrainAGE estimate
+  filter(eventname == "2_year_follow_up_y_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_m4_y","pds_m5_y")) %>% 
+  psych::omega(nfactors=3)
+
+# Follow-up, youth-report, female
+youth_pds %>% 
+  #filter for baseline wave, female, has BrainAGE estimate
+  filter(eventname == "2_year_follow_up_y_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_f4_2_y","pds_f5_y")) %>% 
+  psych::omega(nfactors=3)
+
+# Follow-up, parent-report, male                           
+parent_pds %>% 
+  #filter for baseline wave, male, has BrainAGE estimate
+  filter(eventname == "2_year_follow_up_y_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m4_p", "pds_m5_p")) %>% 
+  psych::omega(nfactors=3)
+
+# Follow-up, parent-report, female                           
+parent_pds %>% 
+  #filter for baseline wave, female, has BrainAGE estimate
+  filter(eventname == "2_year_follow_up_y_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
+  #select values being used for mean calculation (male)
+  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")) %>% 
+  psych::omega(nfactors=3)
+
+```
+
+Check correlation between youth and parent report puberty
+```{r}
+# baseline youth & parent report
+cor(brainage_pds_baseline$youth_mean, brainage_pds_baseline$parent_mean, use="complete.obs")
+
+# follow-up youth and parent report
+cor(brainage_pds_followup$youth_mean, brainage_pds_followup$parent_mean, use="complete.obs")
+
+# abcd model baseline
+cor(abcd_baseline$youth_mean, abcd_baseline$parent_mean, use="complete.obs")
+
+```
+
+**Combined Wave Model**
+
 ```{r}

### 2023-03-01 DiffScoreModel.Rmd
Updated bias metrics, updated names

@@
 
-abcd_brainage_diff_model <- xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod"))
+#abcd_brainage_diff_model <- xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod"))
+
+abcd_brainage_diff_model <- readRDS("xgboost_abcd_diff_brain_age_mod_updated.rds")
+
+#https://stackoverflow.com/questions/73506085/old-json-xgboost-model-format-will-be-discontinued-in-version-2-3-warning
 
@@
     # provide the chronological age at time of scan
-    truth = df_validation_baseline$interview_age
+    truth = df_validation_diff$age_midpoint
   ) %>%
@@
 # extract intercept and slope
-#abcd_bias_mod_diffcoefficients[["(Intercept)"]]
-abcd_bias_intercept_diff <- 
+#abcd_bias_mod_diff$coefficients[["(Intercept)"]]
+abcd_bias_intercept_diff <- 9.582399
 #abcd_bias_mod_diff$coefficients[["truth"]]
-abcd_bias_slope_diff <-
+abcd_bias_slope_diff <-  0.1240839
 
@@
 
-save(brain_age_df_diff, file="ABCDBaselineBrainAge.Rda")
-save(brain_age_corrected_df_diff, file="ABCDBaselineBrainAgeCorrected.Rda")
+save(brain_age_df_diff, file="ABCDBrainAgeDiff.Rda")
+save(brain_age_corrected_df_diff, file="ABCDBrainAgeDiffCorrected.Rda")
 

### 2023-03-01 ABCDDiffScoreAnalyses.Rmd
Created script for difference score analyses

---
title: "DiffScoreAnalyses"
author: "Lucy Whitmore"
date: "12/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(vip)
```

Import Data
```{r}

load("~/Volumes/DevBrainLab/Lucy/fyp/ABCDBrainAgeDiffCorrected.Rda")
#brain_age_corrected_df_diff

```

Descriptive Plots
```{r}
hist(brain_age_corrected_df_diff$corrected_gap)

ggpar((brain_age_corrected_df_diff %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Difference Scores")

# Actual vs Corrected Predictions, w/ perfect correlation line

diff_pred_plot <- ggplot(brain_age_corrected_df_diff, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A")+
  geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="diff_truth_corrected_pred_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

# Age Bias - truth vs corrected gap
diff_gap_plot <- ggplot(brain_age_corrected_df_diff, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#619B8A")+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="followup_truth_corrected_gap_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

ggscatter(brain_age_corrected_df_diff, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Panel Plots
ggarrange(diff_pred_plot, diff_gap_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

#ggsave(filename = "pred_bias_mod1.png",
#                          width=10, height=10, units='in', dpi=300)

#Fit Metrics

#baseline uncorrected
brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = corrected_pred)

```

Variable Importance
```{r}

abcd_diff_vi <- vi(xgb_final_mod) #xgb_final_mod, xgboost_abcd_brain_age_mod
#abcd_vi
write_csv(abcd_diff_vi, "abcd_diff_vi.csv")

vip(xgb_final_mod, num_features = 20)  +
  theme_light()

```

### 2023-05-30 DataSetup.Rmd
Cleaned up file organization

@@
 #t test for age differences
-t.test(sample_1_baseline$interview_age, sample_2_baseline$interview_age)
+t_apa(t_test(sample_1_baseline$interview_age, sample_2_baseline$interview_age))
+
+#chi-sq for sex differences
+#chisq_apa(chisq.test(sample_1_baseline$sex, sample_2_baseline$sex))
+
+prop <- c(length(which(sample_2_baseline$sex == "F")) / length(which(sample_2_baseline$sex == "F" | sample_2_baseline$sex == "M")), length(which(sample_2_baseline$sex == "M")) / length(which(sample_2_baseline$sex == "F" | sample_2_baseline$sex == "M")))
+
+chisq_apa(chisq.test(x = table(sample_1_baseline$sex), p = prop))
 
@@
 #t test for age differences
-t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)
+t_apa(t_test(sample_1_followup$interview_age, sample_2_followup$interview_age))
+
+#chi-sq for sex differences
+
+props <- c(length(which(sample_2_followup$sex == "F")) / length(which(sample_2_followup$sex == "F" | sample_2_followup$sex == "M")), length(which(sample_2_followup$sex == "M")) / length(which(sample_2_followup$sex == "F" | sample_2_followup$sex == "M")))
+
+chisq_apa(chisq.test(x = table(sample_1_followup$sex), p = props))
 

### 2023-05-30 Difference Scores /ABCDDiffScoreAnalyses.Rmd
Cleaned up file organization

--- 2023-03-01 ABCDDiffScoreAnalyses.Rmd
@@
 
-load("~/Volumes/DevBrainLab/Lucy/fyp/ABCDBrainAgeDiffCorrected.Rda")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/Diff Scores/ABCDBrainAgeDiffCorrected.Rda")
 #brain_age_corrected_df_diff
@@
   geom_jitter(color="#619B8A")+
-  geom_abline(a=0, b=1, color='#233D4D') +
+ # geom_abline(a=0, b=1, color='#233D4D') +
   labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
@@
   #                        width=7, height=4, units='in', dpi=300)
+
+diff_pred_plot
+
+cor(brain_age_corrected_df_diff$truth, brain_age_corrected_df_diff$corrected_pred)
+
+# uncorrected 
+ggplot(brain_age_corrected_df_diff, aes(x=truth, y=.pred)) + 
+  geom_jitter(color="#619B8A")+
+ # geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
+  theme_classic() + 
+  theme(text = element_text(size = 14))
 

### 2023-05-30 Difference Scores /ABCDDiffScoreAnalyses_Updated.Rmd
Cleaned up file organization

---
title: "DiffScoreAnalyses"
author: "Lucy Whitmore"
date: "12/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(vip)
```

Import Data
```{r}
# make sure to move this file to the server
load("/Volumes/devbrainlab/Lucy/BrainAGE/Diff Scores/Corrected_Diff_BrainAGE.Rda.Rda")
#brain_age_corrected_df_diff

```

Descriptive Plots
```{r}
hist(brain_age_diff_corrected_df$corrected_gap)

ggpar((brain_age_diff_corrected_df %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Difference Scores")

# Actual vs Corrected Predictions, w/ perfect correlation line

diff_pred_plot <- ggplot(brain_age_diff_corrected_df, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="diff_truth_corrected_pred_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

diff_pred_plot

cor(brain_age_diff_corrected_df$truth, brain_age_diff_corrected_df$corrected_pred)

cor.test(brain_age_diff_corrected_df$truth, brain_age_diff_corrected_df$corrected_pred)

# uncorrected 
ggplot(brain_age_diff_corrected_df, aes(x=truth, y=.pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# Age Bias - truth vs corrected gap
diff_gap_plot <- ggplot(brain_age_diff_corrected_df, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#619B8A")+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="followup_truth_corrected_gap_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

ggscatter(brain_age_diff_corrected_df, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Panel Plots
ggarrange(diff_pred_plot, diff_gap_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

#ggsave(filename = "pred_bias_mod1.png",
#                          width=10, height=10, units='in', dpi=300)

#Fit Metrics

#baseline uncorrected
brain_age_diff_corrected_df %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
brain_age_diff_corrected_df %>%
  metrics(truth = truth, estimate = corrected_pred)

```

Variable Importance
```{r}
final_xgb_diff %>%
  fit(data = df_train) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

#abcd_diff_vi <- vi(xgb_final_mod)

#write_csv(abcd_diff_vi, "abcd_diff_vi.csv")

#vip(xgb_final_mod, num_features = 20)  +
#  theme_light()

```

### 2023-05-30 Difference Scores /DifferenceScores.Rmd
Cleaned up file organization

--- 2022-11-07 DifferenceScores.Rmd
@@
   select(src_subject_id, ends_with("diff"))
-  #divide all brain columns by interview age
 
+#divide all brain columns by change in interview age
 smri_difference_calculated_subset[, -(1:2)] <- sweep(smri_difference_calculated_subset[, -(1:2)], 1, smri_difference_calculated_subset[, 2], "/")
@@
 ```
+
+############################################################
+MAY 2023 Update - New Approach
+###########################################################
+
+Updates to this version: Predicting T2 age from difference scores (scaled by time interval, in delta interview age)
+
+Load Structural Data
+```{r}
+#Load list of included subjects
+struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(1:11) %>% 
+  mutate(interview_age = as.numeric(interview_age),
+        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
+  filter(imgincl_t1w_include == 1) %>% 
+  filter(eventname=="baseline_year_1_arm_1")
+
+struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(1:11) %>% 
+  mutate(interview_age = as.numeric(interview_age),
+        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
+  filter(imgincl_t1w_include == 1) %>% 
+  filter(eventname=="2_year_follow_up_y_arm_1")
+
+#Load structural data
+#Filter for volume & area measurements
+smri_baseline_v2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
+  filter(eventname=="baseline_year_1_arm_1") %>% 
+  filter(!collection_title=="collection_title") %>% 
+  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area|thick")) %>% 
+  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
+  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
+
+smri_followup_v2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
+  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
+  filter(!collection_title=="collection_title") %>% 
+  filter(src_subject_id%in%struc_include_followup$src_subject_id) %>% 
+  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area|thick")) %>% 
+  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
+  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
+
+#renamed_smri_baseline<-rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/renamed_smri_baseline.Rda")
+
+```
+
+Load and join data
+```{r}
+
+#Convert to numeric
+smri_baseline_v2[,6:ncol(smri_baseline_v2)] <- sapply(smri_baseline_v2[,6:ncol(smri_baseline_v2)],as.numeric)
+
+# Calculate age in years
+smri_baseline_v2 <- smri_baseline_v2 %>% 
+  mutate(interview_age = as.numeric(interview_age)/12)
+
+#Convert to numeric
+smri_followup_v2[,6:ncol(smri_followup_v2)] <- sapply(smri_followup_v2[,6:ncol(smri_followup_v2)],as.numeric)
+
+# calculate age in years
+smri_followup_v2 <- smri_followup_v2 %>% 
+  mutate(interview_age = as.numeric(interview_age)/12)
+
+smri_difference_v2 <- left_join(smri_baseline_v2, smri_followup_v2, by ="src_subject_id")
+```
+
+Calculate difference scores
+```{r}
+smri_difference_calculated_v2 <- smri_difference_v2 %>%
+ select(-c(subjectkey.x,subjectkey.y, sex.x,sex.y, eventname.x,eventname.y)) %>% 
+ mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
+ rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))
+
+#need to subtract, then divide by age gap
+```
+
+Set up data
+```{r}
+smri_difference_v2 <- smri_difference_calculated_v2 %>% 
+  select(src_subject_id, interview_age.y, ends_with("diff"))
+
+#divide all brain columns by change in interview age
+# this gets us per year change (not a %, but numeric change per year)
+smri_difference_v2[, -(1:3)] <- sweep(smri_difference_v2[, -(1:3)], 1, smri_difference_v2[, 3], "/")
+
+smri_difference_v2 <- smri_difference_v2 %>% 
+  filter(!is.na(interview_age.y_diff))
+
+#1000 people for training
+
+#Split sample
+smp_size_diff_v2 <- floor(0.25 * nrow(smri_difference_v2)) #training sample of 1000 participants
+
+## set the seed to make your partition reproducible
+set.seed(123)
+sample_ind_diff_v2 <- sample(seq_len(nrow(smri_difference_v2)), size = smp_size_diff_v2)
+
+sample_train_diff_v2 <- smri_difference_v2[sample_ind_diff_v2, ]
+sample_test_diff_v2 <- smri_difference_v2[-sample_ind_diff_v2, ]
+
+# select just subject_id, brain columns, age midpoint
+sample_train_diff_v2 <- sample_train_diff_v2 %>% 
+select(-c(src_subject_id, interview_age.y_diff))
+
+save(sample_train_diff_v2, file="sample_train_diff_v2.Rda")
+save(sample_test_diff_v2, file="sample_test_diff_v2.Rda")
+```

### 2023-05-30 Difference Scores /UpdatedDiffScoreModel.Rmd
Cleaned up file organization

---
title: "Train BrainAGE with tidymodels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```

```{r}
library(tidymodels)
library(xgboost)
```

```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/Difference Scores/sample_train_diff_v2.Rda")

#training_subset_test <- training_sample_baseline[1:100,]

# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split <- initial_split(
  sample_train_diff_v2, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age.y"
)
df_train <- training(df_split)
df_validation <- testing(df_split)

```

# Pre-processing setup ----------------------------------------------------

```{r}
# define (what we want to do)
preprocess_recipe <- df_train %>%
  # predict scan age by all brain features
  recipe(interview_age.y ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_prep <- juice(preprocess_recipe)

# apply on validation
df_validation_prep <- preprocess_recipe %>% bake(df_validation)

```

```{r}
boost_mod <- boost_tree(
  mode = "regression", 
  trees = 150, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

```

```{r}
set.seed(42)

xgboost_grid <- grid_latin_hypercube(
  min_n(), 
  tree_depth(), 
  loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_prep),
  learn_rate(),
  size = 500 
)

```

```{r}
xgb_wf <- workflow() %>%
  add_formula(interview_age.y ~ .) %>%
  add_model(boost_mod)

xgb_wf
```

```{r}
set.seed(42)

train_cv <- df_train_prep %>%
  vfold_cv(
    v = 10, 
    repeats = 10, 
    strata = interview_age.y
  )
```

```{r}
doParallel::registerDoParallel()

set.seed(42)

xgb_tuned_results <- tune_grid(
  xgb_wf,
  resamples = train_cv,
  grid = xgboost_grid,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = TRUE,
                         save_pred = TRUE)
)

xgb_tuned_results
```

```{r}
xgb_tuned_results %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params <- xgb_tuned_results %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

```

Finalize workflow and fit final model
```{r}
final_xgb_diff <- finalize_workflow(
  xgb_wf,
  best_xgb_params
)

final_xgb_diff

fit_workflow_diff <- fit(final_xgb_diff, df_train_prep)

#xgb_final_mod_test <- boost_mod_test %>%
 # finalize_model(best_xgb_params_test) %>%
 # fit(interview_age ~ .,
#      data = df_train_prep)

```

Variable importance, ignore
```{r}
library(vip)

final_xgb_diff %>%
  fit(data = df_train) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

```

```{r}
final_res <- last_fit(final_xgb_diff, df_split)

collect_metrics(final_res)
```

Save models
```{r}
# three different save formats just in case
# 1. Rds
save(fit_workflow_diff, file = "fit_workflow_diff.rds")
# 2. rda
save(fit_workflow_diff, file = "fit_workflow_2.rda")
# 3. raw xbg object, which can then be loaded with xgb.load
model_obj_diff <- fit_workflow_diff$fit$fit$fit
xgb.save(model_obj_diff, "model_obj_diff")
```

```{r}

model_obj <- fit_workflow$fit$fit$fit
xgb.save(model_obj, "model_obj")

xgb.save.raw(model_obj, "model_obj_raw")
```

Predict
redo this part as needed when the model's done
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/Difference Scores/sample_test_diff_v2.Rda")

# dou
analysis_sample_diff <- sample_test_diff_v2 %>% 
  select(-c(src_subject_id, interview_age.y_diff))

#workflow_predict <- predict(fit_workflow, analysis_sample)

brain_age_diff <- fit_workflow_diff %>%
  predict(new_data = analysis_sample_diff) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = analysis_sample_diff$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Alternate Prediction, if using raw model object 
```{r}
brain_age_df_test <-
  model_obj %>%
  predict(newdata = as.matrix(analysis_sample %>% select(-c(interview_age))))

brain_age_df_test <- as.data.frame(brain_age_df_test) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = analysis_sample_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

```

Bias Correction
```{r}
#run model on validation set
brain_age_diff_correction <-
  fit_workflow_diff %>%
  predict(new_data = df_validation) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
brain_age_diff_correction %>%
  metrics(truth = truth, estimate = .pred)

diff_bias_mod <- lm(.pred ~ truth, data = brain_age_diff_correction)

# extract intercept and slope
#diff_bias_mod$coefficients[["(Intercept)"]]
diff_bias_intercept <- 10.27455
#diff_bias_mod$coefficients[["truth"]]
diff_bias_slope <- 0.1411708

# create bias correct data frame
#Baseline
brain_age_diff_corrected_df <- brain_age_diff %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - diff_bias_intercept) / diff_bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

```

Save corrected estimates
```{r}
save(brain_age_diff_corrected_df, file="Corrected_Diff_BrainAGE.Rda")
```

### 2023-05-30 Difference Scores /OldVersions/abcd_diff_model_talapas.R

@@ -1,0 +1,174 @@
+
+library(dplyr)
+library(parallel)
+library(doParallel)
+library(tidymodels)
+library(xgboost)
+
+#load data
+
+
+load("/projects/dcnlab/shared/brainage/sample_train_diff.Rda")
+
+# Model prep: split, preprocessing, CV ------------------------------------
+
+# Train / test split ------------------------------------------------------
+set.seed(42)
+df_split_diff <- initial_split(
+  sample_train_diff, 
+  prop = 0.80,
+  # matching age distributions across train and test set
+  strata = "age_midpoint"
+)
+
+df_train_diff <- training(df_split_diff)
+df_validation_diff <- testing(df_split_diff)
+
+# report n training, features, testing
+#df_train_baseline %>% 
+df_train_diff%>%
+  summarise(
+    n_training = n(),
+    p_features = ncol(df_train_diff)
+  ) %>%
+  bind_cols(df_validation_diff %>%
+              summarise(n_testing = n()))
+
+# Pre-processing setup ----------------------------------------------------
+
+# define (what we want to do)
+preprocess_recipe_diff <- df_train_diff %>%
+#preprocess_recipe_baseline <- train_test %>%
+  # predict scan age by all brain features
+  recipe(age_midpoint ~ .) %>%
+  # remove near zero variance predictors
+  step_nzv(all_predictors()) %>%
+  prep() # where it all gets calculated
+
+preprocess_recipe_diff
+
+
+# Apply pre-processing ----------------------------------------------------
+
+# juice() will work with training data, `bake()` to apply this to our test data
+
+# apply on train (gives processed value)
+df_train_prep_diff <- juice(preprocess_recipe_diff)
+
+# apply on validation
+df_validation_prep_diff <- preprocess_recipe_diff %>% bake(df_validation_diff)
+#df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(validation_test)
+
+
+# Cross-validation --------------------------------------------------------
+
+# 10 fold cv repeated 10 times
+set.seed(42)
+train_cv_diff <- df_train_prep_diff %>%
+  vfold_cv(
+    v = 10, 
+    repeats = 10, 
+    strata = age_midpoint
+  )
+
+
+#Train and Fit Model (Baseline)
+
+# Extreme gradient boosting
[…]

### 2023-05-30 abcd_diff_model_talapas.R
More file cleanup

@@ -1,157 +1,0 @@
-
-library(dplyr)
-library(parallel)
-library(doParallel)
-library(tidymodels)
-library(xgboost)
-
-#load data
-
-
-load("/projects/dcnlab/shared/brainage/sample_train_diff.Rda")
-
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-set.seed(42)
-df_split_diff <- initial_split(
-  sample_train_diff, 
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "age_midpoint"
-)
-
-df_train_diff <- training(df_split_diff)
-df_validation_diff <- testing(df_split_diff)
-
-# report n training, features, testing
-#df_train_baseline %>% 
-df_train_diff%>%
-  summarise(
-    n_training = n(),
-    p_features = ncol(df_train_diff)
-  ) %>%
-  bind_cols(df_validation_diff %>%
-              summarise(n_testing = n()))
-
-# Pre-processing setup ----------------------------------------------------
-
-# define (what we want to do)
-preprocess_recipe_diff <- df_train_diff %>%
-#preprocess_recipe_baseline <- train_test %>%
-  # predict scan age by all brain features
-  recipe(age_midpoint ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
-
-preprocess_recipe_diff
-
-
-# Apply pre-processing ----------------------------------------------------
-
-# juice() will work with training data, `bake()` to apply this to our test data
-
-# apply on train (gives processed value)
-df_train_prep_diff <- juice(preprocess_recipe_diff)
-
-# apply on validation
-df_validation_prep_diff <- preprocess_recipe_diff %>% bake(df_validation_diff)
-#df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(validation_test)
-
-
-# Cross-validation --------------------------------------------------------
-
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv_diff <- df_train_prep_diff %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, 
-    strata = age_midpoint
-  )
-
-
-#Train and Fit Model (Baseline)
-
-# Extreme gradient boosting
-# Parallel processing -----------------------------------------------------
-
[…]

### 2023-05-30 multiwave_model_talapas.R

@@ -1,149 +1,0 @@
-library(dplyr)
-library(parallel)
-library(doParallel)
-library(tidymodels)
-
-#load data
-
-
-load("/projects/dcnlab/shared/brainage/abcd_train_multiwave.Rds")
-
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-set.seed(42)
-df_split <- initial_split(
-  abcd_train_multiwave, 
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "interview_age"
-)
-
-df_train <- training(df_split)
-df_validation <- testing(df_split)
-
-# report n training, features, testing
-#df_train_baseline %>% 
-df_train%>%
-  summarise(
-    n_training = n(),
-    p_features = ncol(df_train)
-  ) %>%
-  bind_cols(df_validation %>%
-              summarise(n_testing = n()))
-
-# Pre-processing setup ----------------------------------------------------
-
-# define (what we want to do)
-preprocess_recipe <- df_train %>%
-  #preprocess_recipe_baseline <- train_test %>%
-  # predict scan age by all brain features
-  recipe(interview_age ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
-
-preprocess_recipe
-
-
-# Apply pre-processing ----------------------------------------------------
-
-# juice() will work with training data, `bake()` to apply this to our test data
-
-# apply on train (gives processed value)
-df_train_prep <- juice(preprocess_recipe)
-
-# apply on validation
-df_validation_prep <- preprocess_recipe_baseline %>% bake(df_validation)
-
-
-# Cross-validation --------------------------------------------------------
-
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv <- df_train_prep %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, 
-    strata = interview_age
-  )
-
-
-#Train and Fit Model (Baseline)
-
-# Extreme gradient boosting
-# Parallel processing -----------------------------------------------------
-
-all_cores <- parallel::detectCores(logical = TRUE)
-# for 8 core 16 thread machine, good performance running more than 
-# physical but less than all logical
-registerDoParallel(cores = all_cores - 6) 
-
[…]

### 2023-05-30 UpdatedModels/ABCDModelTraining.Rmd
Added updated model training script

---
title: "Train BrainAGE with tidymodels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```

```{r}
library(tidymodels)
library(xgboost)
```

```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/training_sample_baseline.Rds")

#training_subset_test <- training_sample_baseline[1:100,]

# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split <- initial_split(
  training_sample_baseline, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)
df_train <- training(df_split)
df_validation <- testing(df_split)

```

# Pre-processing setup ----------------------------------------------------

```{r}
# define (what we want to do)
preprocess_recipe <- df_train %>%
  # predict scan age by all brain features
  recipe(interview_age ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_prep <- juice(preprocess_recipe)

# apply on validation
df_validation_prep <- preprocess_recipe %>% bake(df_validation)

```

```{r}
boost_mod_test <- boost_tree(
  mode = "regression", 
  trees = 150, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

```

```{r}
set.seed(42)

xgboost_grid_test <- grid_latin_hypercube(
  min_n(), 
  tree_depth(), 
  loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_prep),
  learn_rate(),
  size = 500 
)

```

```{r}
xgb_wf <- workflow() %>%
  add_formula(interview_age ~ .) %>%
  add_model(boost_mod_test)

xgb_wf
```

```{r}
set.seed(42)

train_cv <- df_train_prep %>%
  vfold_cv(
    v = 10, 
    repeats = 10, 
    strata = interview_age
  )
```

```{r}
doParallel::registerDoParallel()

set.seed(42)

xgb_tuned_results_test <- tune_grid(
  xgb_wf,
  resamples = train_cv,
  grid = xgboost_grid_test,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = TRUE,
                         save_pred = TRUE)
)

xgb_tuned_results_test
```

```{r}
xgb_tuned_results_test %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params_test <- xgb_tuned_results_test %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

```

Finalize workflow and fit final model
```{r}
final_xgb <- finalize_workflow(
  xgb_wf,
  best_xgb_params_test
)

final_xgb

fit_workflow <- fit(final_xgb, df_train_prep)

#xgb_final_mod_test <- boost_mod_test %>%
 # finalize_model(best_xgb_params_test) %>%
 # fit(interview_age ~ .,
#      data = df_train_prep)

```

Variable importance, ignore
```{r}
library(vip)

final_xgb %>%
  fit(data = df_train) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

```

```{r}
final_res <- last_fit(final_xgb, df_split)

collect_metrics(final_res)
```

Save models
```{r}
# three different save formats just in case
# 1. Rds
save(fit_workflow, file = "fit_workflow.rds")
# 2. rda
save(fit_workflow, file = "fit_workflow_2.rda")
# 3. raw xbg object, which can then be loaded with xgb.load
model_obj <- fit_workflow$fit$fit$fit
xgb.save(model_obj, "model_obj")
```

Model bundling, ignore for now
```{r}
# Testing out bundling
#mod_bundle <- bundle::bundle(fit_workflow)
#save(mod_bundle, file = "xgb_workflow_bundles.rds")

#save(mod_bundle, file = "xgb_workflow.rds")

# in a new R session:
#mod_bundle <- readRDS("path/to/file.rds")
#mod_new <- unbundle(mod_bundle)

#xgb.save.raw(fit_workflow$fit$fit, raw_format = "deprecated")

#model_obj <- fit_workflow$fit$fit$fit
#xgb.save(model_obj, "model_obj")

#xgb.save.raw(model_obj, "model_obj_raw")
```

Predict
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/analysis_sample_baseline.Rds")

analysis_sample <- analysis_sample_baseline %>% 
  select(-c(subjectkey, src_subject_id, sex, eventname))

brain_age_df_test <-
  model_obj %>%
  predict(newdata = as.matrix(analysis_sample %>% select(-c(interview_age))))

brain_age_df_test <- as.data.frame(brain_age_df_test) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = analysis_sample_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#workflow_predict <- predict(fit_workflow, analysis_sample)

brain_age_df_workflow <- fit_workflow %>%
  predict(new_data = analysis_sample) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = analysis_sample$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Bias Correction
```{r}
#run model on validation set
abcd_brain_age_correction <-
  fit_workflow %>%
  predict(new_data = df_validation) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod <- lm(.pred ~ truth, data = abcd_brain_age_correction)

# extract intercept and slope
#
abcd_bias_intercept <- abcd_bias_mod$coefficients[["(Intercept)"]]

abcd_bias_slope <-  abcd_bias_mod$coefficients[["truth"]]

# create bias correct data frame
#Baseline
brain_age_corrected_df <- brain_age_df %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept) / abcd_bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

```

### 2023-05-30 model-prep.R
Removed old training files

@@ -1,61 +1,0 @@
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-
-set.seed(42)
-df_split <- initial_split(
-  df_select,
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "scan_age"
-)
-
-df_train <- training(df_split)
-df_validation <- testing(df_split)
-
-# report n training, features, testing
-df_train %>%
-  summarise(
-    n_training = n(),
-    p_features = ncol(df_train)
-  ) %>%
-  bind_cols(df_validation %>%
-              summarise(n_testing = n()))
-
-# Pre-processing setup ----------------------------------------------------
-
-# define (what we want to do)
-preprocess_recipe <- df_train %>%
-  # predict scan age by all brain features
-  recipe(scan_age ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
-
-preprocess_recipe
-
-
-# Apply pre-processing ----------------------------------------------------
-
-# juice() will work with training data, `bake()` to apply this to our test data
-
-# apply on train (gives processed value)
-df_train_prep <- juice(preprocess_recipe)
-
-# apply on validation
-df_validation_prep <- preprocess_recipe %>% bake(df_validation)
-
-# pre process on all of forbow
-frb_prep <- preprocess_recipe %>% bake(frb_select)
-
-
-# Cross-validation --------------------------------------------------------
-
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv <- df_train_prep %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, 
-    strata = scan_age
-  )

### 2023-05-30 train-fit-model.R

@@ -1,78 +1,0 @@
-# Extreme gradient boosting
-# Parallel processing -----------------------------------------------------
-
-library(doParallel)
-all_cores <- parallel::detectCores(logical = TRUE)
-# for 8 core 16 thread machine, good performance running more than 
-# physical but less than all logical
-registerDoParallel(cores = all_cores - 6) 
-
-# Model specification -----------------------------------------------------
-
-boost_mod <- boost_tree(
-  mode = "regression", 
-  trees = 1500, 
-  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
-  # randomness
-  sample_size = tune(), mtry = tune(), 
-  # step size
-  learn_rate = tune()
-  ) %>%
-  set_engine("xgboost", 
-             objective = "reg:squarederror")
-
-# Model parameters --------------------------------------------------------
-
-set.seed(42)
-
-# filled 6 dimensional tuning space
-xgboost_grid <- grid_latin_hypercube(
-  min_n(), tree_depth(), loss_reduction(),
-  sample_size = sample_prop(),
-  # has unknown, finalize with data to find max
-  finalize(mtry(), df_train_prep),
-  learn_rate(),
-  size = 500 
-)
-
-# Model tuning -----------------------------------------------------------
-
-tictoc::tic()
-
-set.seed(42)
-xgb_tuned_results <- tune_grid(
-  boost_mod,
-  scan_age ~ .,
-  resamples = train_cv,
-  grid = xgboost_grid,
-  metrics = metric_set(mae, rmse, rsq),
-  control = control_grid(verbose = FALSE,
-                         save_pred = TRUE)
-)
-
-tictoc::toc()
-beepr::beep(2)
-
-# Pick best model ---------------------------------------------------------
-
-xgb_tuned_results %>%
-  # want to minimize MAE
-  show_best("mae") 
-
-# select parsimonious params within one SE of best model
-best_xgb_params <- xgb_tuned_results %>%
-  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth) 
-
-
-# Fit and save best model -------------------------------------------------
-
-xgb_final_mod <- boost_mod %>%
-  finalize_model(best_xgb_params) %>%
-  fit(scan_age ~ .,
-      data = df_train_prep)
-
-
-# save mod
-saveRDS(xgb_final_mod, file = here::here(
-  "model", "xgboost_9to19_brain_age_mod.rds")
-  )

### 2023-05-30 abcd_model_talapas.R
removed another old training code

@@ -1,151 +1,0 @@
-
-library(dplyr)
-library(parallel)
-library(doParallel)
-library(tidymodels)
-
-#load data
-
-
-load("/projects/dcnlab/shared/brainage/training_sample_baseline.Rds")
-
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-set.seed(42)
-df_split_baseline <- initial_split(
-  training_sample_baseline, 
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "interview_age"
-)
-
-df_train_baseline <- training(df_split_baseline)
-df_validation_baseline <- testing(df_split_baseline)
-
-# report n training, features, testing
-#df_train_baseline %>% 
-df_train_baseline%>%
-  summarise(
-    n_training = n(),
-   p_features = ncol(df_train_baseline)
-  ) %>%
-  bind_cols(df_validation_baseline %>%
-              summarise(n_testing = n()))
-
-# Pre-processing setup ----------------------------------------------------
-
-# define (what we want to do)
-preprocess_recipe_baseline <- df_train_baseline %>%
-#preprocess_recipe_baseline <- train_test %>%
-  # predict scan age by all brain features
-  recipe(interview_age ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
-
-preprocess_recipe_baseline
-
-
-# Apply pre-processing ----------------------------------------------------
-
-# juice() will work with training data, `bake()` to apply this to our test data
-
-# apply on train (gives processed value)
-df_train_prep_baseline <- juice(preprocess_recipe_baseline)
-
-# apply on validation
-df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(df_validation_baseline)
-#df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(validation_test)
-
-
-# Cross-validation --------------------------------------------------------
-
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv_baseline <- df_train_prep_baseline %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, 
-    strata = interview_age
-  )
-
-
-#Train and Fit Model (Baseline)
-
-# Extreme gradient boosting
-# Parallel processing -----------------------------------------------------
-
[…]

### 2023-05-31 ABCDModel.Rmd
Updated import

@@
                "data.table", "psych",
-               "parallel","lubridate",
+               "parallel","lubridate", "here",
                "ggpubr", "broom", 
@@
 ```{r}
-abcd_brainage_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_brain_age_mod_5_9.rds")
+abcd_brainage_model <- readRDS(here("UpdatedModels", "abcd_fit_workflow.rds"))
 ```

### 2023-09-14 Revisions_Analyses&Figures.Rmd
New script for paper revisions

---
title: "Revisions Analyses & Figures"
author: "Lucy Whitmore"
date: "7/26/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install combat
```{r}
library(devtools)
install_github("jfortin1/neuroCombatData")
install_github("jfortin1/neuroCombat_Rpackage")

library(tidymodels) # machine learning metaverse
library(xgboost) # main engine, needs version 1.0.0.2, see above
library(effectsize)
```

Load everything (uncorrected for scanner)
```{r}
# BrainAGE and Behavioral Data - Vlad's Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")

# BrainAGE and Behavioral Data - Vlad's Model, Follow-Up
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSFollowUp.Rda")

# BrainAGE and Behavioral Data - ABCD Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineSample.Rda")
```

Load Structural Data
```{r}
#Load list of included subjects
struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="baseline_year_1_arm_1")

struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#Load structural data
#Filter for volume & area measurements

##### Need to also select scanner/site #####
smri_baseline <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="baseline_year_1_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

smri_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id %in% struc_include_followup$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

```

# combat/site standardization
```{r}
# load raw data
# filter for quality data
# keep scanner site

#https://github.com/Jfortin1/neuroCombat_Rpackage/tree/2ccab2b42c59163717f012c70a54d20e8757b1c7
# can realign to virtual site or to reference site

data.harmonized <- neuroCombat(dat=dat, batch=batch) # batch is the site/scanner
# maybe need to add age (and gender?) as a covariate, mod =

# maybe run without empirical bayes because number of features is smaller than number of participants?
#data.harmonized <- neuroCombat(dat=dat, batch=batch, eb=FALSE)

# Load data harmonized using LongCombat - longCombat.R
load("Volumes/devbrainlab/Lucy/BrainAGE/fyp/harmonized_mri_baseline.Rda")
load("Volumes/devbrainlab/Lucy/BrainAGE/fyp/harmonized_mri_followup.Rda")

```

Data Cleaning for harmonized data
```{r}
mod_1 <- readRDS(
  file = here::here(
    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))

# check for any columns that aren't in model
names.use <- names(model1_features_baseline)[!(names(model1_features_baseline) %in% mod_1$fit$feature_names)]

# baseline
# remove those columns
model1_features_baseline_clean <- model1_features_baseline %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_baseline = as.matrix(model1_features_baseline_clean)
newdata_baseline = newdata_baseline[ , mod_1$fit$feature_names]

### Follow-Up
# remove those columns
model1_features_followup_clean <- model1_features_followup %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_followup = as.matrix(model1_features_followup_clean)
newdata_followup = newdata_followup[ , mod_1$fit$feature_names]
```

# rerun predictions on harmonized data
```{r}
# drop scanner site variable from data

# predict brainage - baseline, model 1

# Predict
# This prediction method gives the output as a vector, we'll fix that later on
brainage_mod1_baseline <-
  mod_1$fit %>%
  predict(newdata = newdata_baseline)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_baseline <- as.data.frame(brainage_mod1_baseline) %>%
  rename(.pred = brainage_mod1_baseline) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_baseline$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# bias correction
# Slope and intercept are hardcoded
bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
bias_slope <- 0.55 # xgb_bias_mod$coefficients[["truth"]]

# create bias corrected data frame
brain_age_df_mod1_baseline_corrected <- brain_age_df_mod1_baseline %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_baseline_corrected$src_subject_id <- harmonized_mri_baseline$src_subject_id

# predict brainage - followup, model 1
brainage_mod1_followup <-
  mod_1$fit %>%
  predict(newdata = newdata_followup)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_followup <- as.data.frame(brainage_mod1_followup) %>%
  rename(.pred = brainage_mod1_followup) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_followup$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# create bias corrected data frame
brain_age_df_mod1_followup_corrected <- brain_age_df_mod1_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_followup_corrected$src_subject_id <- harmonized_mri_followup$src_subject_id

# predict brainage - baseline, model 2
#brainage <-
 # fit_workflow %>%
#  predict(new_data = analysis_sample_baseline) %>%
#  mutate(
    # provide the chronological age at time of scan
#    truth = analysis_sample_baseline$interview_age
 # ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
#  mutate(gap = .pred - truth)

# predict brainage - follow-up, model 2 (3?)

#followup_brainage <-
#  fit_workflow_followup %>%
#  predict(new_data = analysis_sample_baseline) %>%
#  mutate(
    # provide the chronological age at time of scan
#    truth = analysis_sample_baseline$interview_age
 # ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
 # mutate(gap = .pred - truth)
```
# model features and importance
```{r}
# vip plot of feature importance
library(vip)

vip(abcd_model_obj_combat, num_features = 20)

vi(abcd_model_obj_combat)

# ggseg importance plots
# Rename columns to fit with ggseg atlas (ugghhh)

# select only name and atlas name, then join with new vip estimates to get around renaming by hand
abcd_vi_renamed <- rio::import("~/Documents/FYP/abcd_vi_renamed.csv") %>% 
  select(Variable, brain_labels_dk, label)

abcd_vi <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed)

abcd_vi_renamed_dk <- rio::import("~/Documents/FYP/abcd_vi_renamed_dk.csv") %>% 
  select(Variable, label, aseg_label)

abcd_vi_dk <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed_dk)

#check labels
brain_labels(dk)
brain_labels(aseg)

#Cortical - Volume
dk_vi_vol <- abcd_vi_dk %>%
  filter(grepl('Vol', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_vol <- data.frame(dk_vi_vol)

ggplot(dk_vi_vol) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Volume")
ggsave(filename="abcd_model_dk_vol_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Cortical - Area
dk_vi_area <- abcd_vi_renamed_dk %>%
  filter(grepl('Area', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_area <- data.frame(dk_vi_area)

ggplot(dk_vi_area) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Area")
ggsave(filename="abcd_model_dk_area_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Subcortical
aseg_vi <- abcd_vi_renamed %>%
  filter(label != "") %>% 
  select(Importance,label)

aseg_vi <- data.frame(aseg_vi)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_coronal.png",
                          width=7, height=4, units='in', dpi=300)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_sagittal.png",
                          width=7, height=4, units='in', dpi=300)

#troubleshooting names
aseg_labs <- brain_labels(aseg) |> 
  as.character()
aseg_vi$label[!aseg_vi$label %in% aseg_labs]
aseg_labs[!aseg_labs %in% aseg_vi$label]

dk_labs <- brain_labels(dk) |> 
  as.character()
dk_vi_area$label[!dk_vi_area$label %in% dk_labs]
dk_vi_vol$label[!dk_vi_vol$label %in% dk_labs]

dk_labs[!dk_labs %in% dk_vi_area$label]
dk_labs[!dk_labs %in% dk_vi_vol$label]

```

# attach pds and cognition measures to BrainAGE estimates
```{r}
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_uncorrected) %>% 
  mutate_at(c(2, 4:7), as.numeric)

#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

### Youth-report, removes everyone w/o complete data
youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

# Multiwave
youth_pds_multiwave <- youth_pds %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)  %>% 
  select(src_subject_id, eventname, sex, youth_sum, youth_mean)

youth_pds_baseline <- youth_pds_multiwave %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

youth_pds_followup <- youth_pds_multiwave %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  select(-eventname)

#### Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA

#Baseline
parent_pds_multiwave <- parent_pds %>% 
    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5) %>% 
  select(src_subject_id, eventname, sex, parent_sum, parent_mean)

parent_pds_baseline <- parent_pds_multiwave %>% 
    filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

parent_pds_followup <- parent_pds_multiwave%>% 
    filter(eventname == "2_year_follow_up_y_arm_1")%>% 
  select(-eventname)

#### Cognition

#Baseline
cognition_baseline <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname) %>%
  mutate(cog_zscore = scale(nihtbx_totalcomp_uncorrected))

#### Create Dataframes for Analysis #####
# join brainage, pds, cognition - change names
brainage_pds_baseline_combat <- brain_age_df_mod1_baseline_corrected[, c(6,1:5)] %>% 
  left_join(youth_pds_baseline) %>% 
  left_join(parent_pds_baseline) %>% 
  left_join(cognition_baseline)

brainage_pds_followup_combat <- brain_age_df_mod1_followup_corrected[, c(6,1:5)] %>% 
  left_join(youth_pds_followup) %>% 
  left_join(parent_pds_followup)

save(brainage_pds_baseline_combat, file="brainage_pds_baseline_combat.Rda")
save(brainage_pds_followup_combat, file="brainage_pds_followup_combat.Rda")

# abcd model, baseline
abcd_brainage_pds_baseline_combat <- brain_age_corrected_df_combat %>% 
    left_join(youth_pds_baseline) %>% 
    left_join(parent_pds_baseline) %>% 
    left_join(cognition_baseline)

save(abcd_brainage_pds_baseline_combat, file="abcd_brainage_pds_baseline_combat.Rda")

# abcd model, follow-up
abcd_brainage_pds_followup_combat <- brain_age_corrected_df_combat_followup %>% 
    left_join(youth_pds_followup) %>% 
    left_join(parent_pds_followup)

save(abcd_brainage_pds_followup_combat, file="abcd_brainage_pds_followup_combat.Rda")

load("~/Documents/FYP/abcd_brainage_pds_followup_combat.Rda")

```

# rerun analyses without splitting
```{r}
## Baseline - Model 1
# get age in months
# mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
baseline_puberty_y_mod1 <- lm(corrected_gap ~ youth_mean + truth, data = brainage_pds_baseline_combat)

# brainage ~ parent mean
baseline_puberty_p_mod1 <- lm(corrected_gap ~ parent_mean + truth, data = brainage_pds_baseline_combat)

#standard scores - check that this is actually the one we're trying to use
baseline_cognition_mod1 <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + truth, data=brainage_pds_baseline_combat)

## Followup - Model 1
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
followup_puberty_y_mod1 <- lm(corrected_gap ~ youth_mean + truth, data = brainage_pds_followup_combat)

# brainage ~ parent mean
followup_puberty_p_mod1 <- lm(corrected_gap ~ parent_mean + truth, data = brainage_pds_followup_combat)

## Baseline - Model 2
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
baseline_puberty_y_mod2 <- lm(corrected_gap ~ youth_mean + truth, data = abcd_brainage_pds_baseline_combat)

# brainage ~ parent mean
baseline_puberty_p_mod2 <- lm(corrected_gap ~ parent_mean + truth, data = abcd_brainage_pds_baseline_combat)

#standard scores - check that this is actually the one we're trying to use
baseline_cognition_mod2 <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + truth, data=abcd_brainage_pds_baseline_combat)

## Followup - Model 2
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
followup_puberty_y_mod2 <- lm(corrected_gap ~ youth_mean + truth, data = abcd_brainage_pds_followup_combat)

# brainage ~ parent mean
followup_puberty_p_mod2 <- lm(corrected_gap ~ parent_mean + truth, data = abcd_brainage_pds_followup_combat)

```

# model summaries
```{r}
summary(baseline_puberty_y_mod1)
summary(baseline_puberty_p_mod1)
summary(baseline_cognition_mod1)

summary(followup_puberty_y_mod1)
summary(followup_puberty_p_mod1)

summary(baseline_puberty_y_mod2)
summary(baseline_puberty_p_mod2)
summary(baseline_cognition_mod2)

summary(followup_puberty_y_mod2)
summary(followup_puberty_p_mod2)

```

# maybe make a table of results
```{r}

```

# standardize effect sizes
```{r}
library(effectsize)
# standardize(model) # updated model, refit with standardized data
# standardize_parameters(models) # table of standardized coefficients from model

# Baseline - Model 1
standardize_parameters(baseline_puberty_y_mod1)
standardize_parameters(baseline_puberty_p_mod1)
standardize_parameters(baseline_cognition_mod1)

# Follow-up - Model 1
standardize_parameters(followup_puberty_y_mod1)
standardize_parameters(followup_puberty_p_mod1)

# Baseline - Model 2
standardize_parameters(baseline_puberty_y_mod2)
standardize_parameters(baseline_puberty_p_mod2)
standardize_parameters(baseline_cognition_mod2)

# Followup - Model 2
standardize_parameters(followup_puberty_y_mod2)
standardize_parameters(followup_puberty_p_mod2)
standardize(followup_puberty_p_mod2)
```

# alternate standardization
```{r}
install.packages("lm.beta")
library(lm.beta)

summary(lm.beta(followup_puberty_y_mod2))

```

# sex differences
```{r}
# model with sex as covariate?

```

# Change plot scaling - Model 1
```{r}
# figure 1 and 4, put actual and chronological on the same scale
# maybe will also need to rerun other plots too, for combat + grid plotting

# Baseline
#baseline_pred_mod1_plot <- 
ggplot(brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#A1C181", alpha=.2) +
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(5, 20) + # 5, 20
  ylim(5, 20) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="baseline_truth_corrected_pred_mod1.png",
            #              width=7, height=4, units='in', dpi=300)

# Follow-Up
ggplot(brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A", alpha=.2)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(6, 21) + # 5, 20
  ylim(6, 21) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

```

# Change plot scaling - Model 2
```{r}
ggplot(abcd_brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#E76F51", alpha=.25)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(2, 18) + # 5, 20
  ylim(2, 18) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 12))

ggplot(abcd_brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#BDB5D5", alpha=.25)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(1, 22) + # 5, 20
  ylim(1, 22) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 12))

```

# Fig 1 with line of identity, jitter fixed
```{r}
# fig 1 has different jitter for ab vs. cd. x axis scaling of ab and cd should be equal
# add line of identity to fig 1 to vizualize bias - already done for AB, but CD has a regression line, do they want a 1:1 line on that too?
```

# Plot of model 1 & 2 prediction overlaid (baseline)
```{r}
# overlay model 1 on new predictions in fig 4

# try combining into one dataframe and then grouping by model
# mutate each dataframe to add column for model
# then rbind
both_mods_baseline <- rbind(brainage_pds_baseline_combat %>% mutate(model = 1), abcd_brainage_pds_baseline_combat %>% mutate(model = 2))

both_mods_baseline <- both_mods_baseline %>% 
  mutate(model = as.factor(model))

ggplot(both_mods_baseline, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
     # geom_point() +
  geom_jitter(alpha=.2) +
  #geom_abline(a=0, b=1, color='#233D4D') +
  #xlim(8, 15) + # 5, 20
 # ylim(8, 15) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# change colors
# decide if it should only show the same participants
```
# Plot of model 1 & 2 prediction overlaid (followup)
```{r}
# overlay model 1 on new predictions in fig 4

# try combining into one dataframe and then grouping by model
# mutate each dataframe to add column for model
# then rbind
both_mods_followup <- rbind(brainage_pds_followup_combat %>% mutate(model = 1), abcd_brainage_pds_followup_combat %>% mutate(model = 2))

both_mods_followup <- both_mods_followup %>% 
  mutate(model = as.factor(model))

ggplot(both_mods_followup, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
     # geom_point() +
  geom_jitter(alpha=.2) +
  #geom_abline(a=0, b=1, color='#233D4D') +
  #xlim(8, 15) + # 5, 20
 # ylim(8, 15) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# change colors
# decide if it should only show the same participants
```

### Rerun old figures
```{r}
# Distrubutions of puberty/cognition by model

```

```{r}
# Prediction Plots

```

```{r}
# Model feature - baseline

# vip

# ggseg

```

### New longitudinal analyses ###

figure out how many baseline training v analysis are in follow-up
```{r}
# check how many from training sample are in follow-up

renamed_harmonized_mri %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")  %>% 
  select(-c(eventname, mri_info_deviceserialnumber)) %>% 
  filter(src_subject_id %in% model2_features$src_subject_id)

#3,739

# check how many from analysis are in follow-up

renamed_harmonized_mri %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")  %>% 
  select(-c(eventname, mri_info_deviceserialnumber)) %>% 
  filter(src_subject_id %in% model2_analysis$src_subject_id)

# 3,693

# total: 7432/7695
```

```{r}
# correlation between gaps from waves, or something more complicated?
# maybe multilevel model so that it's nested by participant? - 
# correlation for vlad's models

#correlation for abcd models

# figure - probably spaghetti plot, show direction of gap change

# for each set of models, look at change in brainage vs chronological age?

```

### 2023-09-19 Revisions_Analyses&Figures.Rmd
Added updated plots, demographics, variable importance code

@@
 library(effectsize)
+library(psych)
+library(ggpubr)
+library(vip)
 ```
@@
 ```
+
 # model features and importance
@@
 
+## from model code
+
+final_xgb_combat %>%
+  fit(data = df_train) %>%
+  extract_fit_parsnip() %>%
+  vip(geom = "point")
+
+fit_workflow_combat %>%
+  fit(data = df_train) %>%
+  extract_fit_parsnip() %>%
+  vi()
+
+###
+
+vip(abcd_model_obj_combat, num_features = 175)
+
 vip(abcd_model_obj_combat, num_features = 20)
 
-vi(abcd_model_obj_combat)
+vi_updated <- vi(abcd_model_obj_combat)
+
+vi(fit_workflow_combat)
+
+vi_updated2 <- vi(extract_fit_parsnip(fit_workflow_combat))
+
+#vi_model(abcd_model_obj_combat)
 
@@
 #Cortical - Area
-dk_vi_area <- abcd_vi_renamed_dk %>%
+dk_vi_area <- abcd_vi_dk %>%
   filter(grepl('Area', Variable)) %>% 
@@
 summary(lm.beta(followup_puberty_y_mod2))
+```
+
+# Descriptive Table
+```{r}
+describe(brainage_pds_baseline_combat)
+
+describe(brainage_pds_followup_combat)
+
+describe(abcd_brainage_pds_baseline_combat)
+
+describe(abcd_brainage_pds_followup_combat)
+```
+# New Demographics Column for Follow- Up (Model 2)
+```{r}
+nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "F")) / nrow(abcd_brainage_pds_followup_combat)
+nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "M")) / nrow(abcd_brainage_pds_followup_combat)
+
+demographics <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/pdem02.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, contains("demo_race"), demo_ethn_v2, demo_ethn2_v2, demo_comb_income_v2)
+
+demographics_baseline_abcdmodel <- left_join(abcd_baseline, demographics)
+
+demographics_long <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_lpds01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, eventname, demo_comb_income_v2_l)
+
+demographics_baseline_abcdmodel <- left_join(demographics_baseline_abcdmodel, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
+
+#family income
+table(demographics_baseline$demo_comb_income_v2)
+
+#Overall values
+#Non-hispanic white
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Non-hispanic Black
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Hispanic
+demographics_baseline %>% 
+  filter(demo_ethn_v2 == 1)
+
+demographics_baseline %>% 
+  filter(((demo_race_a_p___10 == 1 | demo_race_a_p___11 == 1) & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1) & demo_ethn_v2 == 1)
+
+#Asian
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Native American Alaska Native
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
+
+#Multiracial
+
+# mutate column: true if participant marked any asian category
+# mutate column: true if participant marked native american/alaska native
+# sum white, black, asian, native american/alaska native, and additional races
+
+demographics_baseline %>% 
+  mutate_at(c(18:33), as.numeric) %>% 
+  mutate(
+    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
+                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
+         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
+                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
+         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>%
+  filter(race_total > 1)
+
+#additional races
+#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.
+
+demographics_baseline %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
 
@@
 # model with sex as covariate?
+
+```
+
+## Rerun prediction plots
+```{r}
+## Baseline - Model 1
+baseline_pred_mod1_plot <- ggplot(brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#A1C181", alpha=.2) +
+  geom_abline(a=0, b=1, color='#233D4D') +
+  #geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
+  theme_classic() + 
+  theme(text = element_text(size = 14))  
+ggsave(filename="baseline_truth_corrected_pred_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+baseline_gap_mod1_plot <- ggplot(brainage_pds_baseline_combat, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#A1C181", alpha=.2)+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
+  theme_classic() + 
+  theme(text = element_text(size = 14))  
+ggsave(filename="baseline_truth_corrected_gap_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+
+ggsave(filename = "pred_bias_baseline_mod1_combat.png",
+                          width=10, height=5, units='in', dpi=300)
+
+## Follow-Up - Model 1
+followup_pred_mod1_plot <- ggplot(brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#619B8A", alpha=.2)+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
+  theme_classic() + 
+  theme(text = element_text(size = 14))  
+ggsave(filename="followup_truth_corrected_pred_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Age Bias - truth vs corrected gap
+followup_gap_mod1_plot <- ggplot(brainage_pds_followup_combat, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#619B8A", alpha=.2)+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
+  theme_classic() + 
+  theme(text = element_text(size = 14))  
+ggsave(filename="followup_truth_corrected_gap_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Panel Plots
+ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, followup_pred_mod1_plot, followup_gap_mod1_plot, 
+          labels = c("A", "B", "C", "D"),
+          ncol = 2, nrow = 2)
+
+ggsave(filename = "pred_bias_mod1_combat.png",
+                          width=10, height=10, units='in', dpi=300)
+
+## Baseline - Model 2
+baseline_pred_mod2_plot <- ggplot(abcd_brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#E76F51", alpha=.25)+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
+  theme_classic() + 
+  theme(text = element_text(size = 12))  
+ggsave(filename="baseline_truth_corrected_pred_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Age Bias - truth vs corrected gap
+baseline_gap_mod2_plot <- ggplot(abcd_brainage_pds_baseline_combat, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#E76F51", alpha=.25)+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
+  theme_classic() + 
+  theme(text = element_text(size = 12))  
+ggsave(filename="baseline_truth_corrected_gap_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Follow-Up - Model 2
+# '#C8A2C8', '#856088','#702963', '#800080'
+followup_pred_mod2_plot <- ggplot(abcd_brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
+  geom_jitter(color="#856088", alpha=.25)+
+  geom_abline(a=0, b=1, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
+  theme_classic() + 
+  theme(text = element_text(size = 12))  
+ggsave(filename="followup_truth_corrected_pred_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+# Age Bias - truth vs corrected gap
+followup_gap_mod2_plot <- ggplot(abcd_brainage_pds_followup_combat, aes(x=truth, y=corrected_gap)) + 
+  geom_jitter(color="#856088", alpha=.25)+
+  geom_smooth(method=lm, color='#233D4D') +
+  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
+  theme_classic() + 
+  theme(text = element_text(size = 12))  
+ggsave(filename="followup_truth_corrected_gap_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Panel Plots
+ggarrange(baseline_pred_mod2_plot, baseline_gap_mod2_plot, followup_pred_mod2_plot, followup_gap_mod2_plot,
+          labels = c("A", "B", "C", "D"),
+          ncol = 2, nrow = 2)
+
+ggsave(filename = "pred_bias_mod2_combat.png",
+                          width=10, height=10, units='in', dpi=300)
 
@@
   theme(text = element_text(size = 12))  
-
 ```
@@
 
+# Model 1 = light green "#A1C181"
+# Model 2 = red "#E76F51"
+
 ggplot(both_mods_baseline, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
-     # geom_point() +
   geom_jitter(alpha=.2) +
+  scale_color_manual(values = c("#A1C181", "#E76F51")) +
   #geom_abline(a=0, b=1, color='#233D4D') +
@@
 
+# model 1: blue/teal "#619B8A"
+# model 2 : purple "#856088"
+
 ggplot(both_mods_followup, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
-     # geom_point() +
   geom_jitter(alpha=.2) +
+  scale_color_manual(values = c("#619B8A", "#856088")) +
   #geom_abline(a=0, b=1, color='#233D4D') +
@@
 ```{r}
-# Distrubutions of puberty/cognition by model
+# Distributions of puberty/cognition by model
+#make categorical variable for puberty
+#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
+categorical_pds_baseline_combat <-  brainage_pds_baseline_combat %>% 
+  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
+         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
+         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
+#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
+
+#Youth-Report Puberty
+youth_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))),  #,'#526D37'
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend = 8) +
+  theme(text = element_text(size = 12)) 
+ggsave(filename="baseline_youth_pds_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Parent-Report Puberty
+parent_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8) +
+  theme(text = element_text(size = 12)) 
+ggsave(filename="baseline_parent_pds_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Cognition
+ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette = c('#7CA352',
+ '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Standardized Cognition")  +
+  theme(text = element_text(size = 24))
+
+ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
+
+ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(cognition_cat_small)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition (.5 sd")
+
+#Cognition Quartiles
+cog_baseline_mod1_plot <- ggpar((categorical_pds_baseline_combat %>% 
+  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#7CA352',
+ '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="baseline_cognition_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_baseline_mod1_plot, parent_pds_baseline_mod1_plot, cog_baseline_mod1_plot, 
+          labels = c("A", "B", "C"),
+          ncol = 2, nrow = 2)
+
+ggsave(filename = "distributions_baseline_mod1_combat.png",
+                          width=7, height=5, units='in', dpi=300)
+
+```
+```{r}
+#Youth-Report Puberty
+#make categorical variable for puberty
+#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
+categorical_pds_followup_combat <-  brainage_pds_followup_combat %>% 
+  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))
+
+#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
+
+youth_pds_followup_mod1_plot <- ggpar((categorical_pds_followup_combat %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="followup_youth_pds_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Parent-Report Puberty
+parent_pds_followup_mod1_plot <-ggpar((categorical_pds_followup_combat %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="followup_parent_pds_mod1_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_followup_mod1_plot, parent_pds_followup_mod1_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+ggsave(filename = "distributions_followup_mod1_combat.png",
+                          width=10, height=4, units='in', dpi=300)
+
+```
+
+```{r}
+#Youth-Report Puberty
+#make categorical variable for puberty
+#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
+abcd_baseline_categorical_pds_combat <- abcd_brainage_pds_baseline_combat %>% 
+  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
+         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
+         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
+
+#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
+
+#Youth-report puberty
+youth_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds_combat %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS",  font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="baseline_youth_pds_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Parent-Report Puberty
+parent_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds_combat %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="baseline_parent_pds_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggpar((abcd_baseline_categorical_pds_combat %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette=c('#FCCA46',
+ '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores") +
+  theme(text = element_text(size = 24))
+
+ggpar((abcd_baseline_categorical_pds_combat %>% 
+  filter(!is.na(cognition_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores) .5 SD") +
+  theme(text = element_text(size = 24))
+
+#Cognition Quartiles
+cog_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds_combat %>% 
+  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
+  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#FCCA46',
+ '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="baseline_cognition_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_baseline_mod2_plot, parent_pds_baseline_mod2_plot, cog_baseline_mod2_plot, 
+          labels = c("A", "B", "C"),
+          ncol = 2, nrow = 2)
+ggsave(filename = "distributions_baseline_mod2_combat.png",
+                          width=7, height=5, units='in', dpi=300)
+
+```
+```{r}
+
+####### UPDATE COLORS
+
+#Youth-Report Puberty
+#make categorical variable for puberty
+#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
+abcd_categorical_pds_followup_combat <-  abcd_brainage_pds_followup_combat %>% 
+  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
+         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))
+
+#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)
+
+youth_pds_followup_mod2_plot <- ggpar((abcd_categorical_pds_followup_combat %>% 
+  filter(!is.na(youth_pds_cat)) %>% 
+  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#C8A2C8', '#856088','#702963', '#800080'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="followup_youth_pds_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+#Parent-Report Puberty
+parent_pds_followup_mod2_plot <-ggpar((abcd_categorical_pds_followup_combat %>% 
+  filter(!is.na(parent_pds_cat)) %>% 
+ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#C8A2C8', '#856088','#702963', '#800080'))), 
+  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend=8)  +
+  theme(text = element_text(size = 12))
+
+ggsave(filename="followup_parent_pds_mod2_combat.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggarrange(youth_pds_followup_mod2_plot, parent_pds_followup_mod2_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+ggsave(filename = "distributions_followup_mod2_combat.png",
+                          width=10, height=4, units='in', dpi=300)
 
@@
 # ggseg
+
+```
+
+```{r}
+# Model feature - follow-up
+
+# vip
+abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
+
+vip(fit_workflow_combat_followup, num_features = 20)
+
+vip(abcd_model_obj_combat_followup, num_features = 20)
+
+vi(fit_workflow_combat_followup$fit$fit$fit)
+
+# ggseg
+
+# select only name and atlas name, then join with new vip estimates to get around renaming by hand
+abcd_vi_renamed <- rio::import("~/Documents/FYP/abcd_vi_renamed.csv") %>% 
+  select(Variable, brain_labels_dk, label)
+
+abcd_vi_followup <- left_join(vi(abcd_model_obj_combat_followup), abcd_vi_renamed)
+
+abcd_vi_renamed_dk <- rio::import("~/Documents/FYP/abcd_vi_renamed_dk.csv") %>% 
+  select(Variable, label, aseg_label)
+
+abcd_vi_dk_followup <- left_join(vi(abcd_model_obj_combat_followup), abcd_vi_renamed_dk)
+
+#check labels
+brain_labels(dk)
+brain_labels(aseg)
+
+#Cortical - Volume
+dk_vi_vol_followup <- abcd_vi_dk_followup %>%
+  filter(grepl('Vol', Variable)) %>% 
+  filter(label != "") %>% 
+  select(Importance,label)
+
+dk_vi_vol_followup <- data.frame(dk_vi_vol_followup)
+
+ggplot(dk_vi_vol_followup) +
+  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
+  labs(title = "Cortical Volume")
+ggsave(filename="abcd_model_dk_vol_importance.png",
+                         width=8, height=4, units='in', dpi=300)
+
+#Cortical - Area
+dk_vi_area_followup <- abcd_vi_dk_followup %>%
+  filter(grepl('Area', Variable)) %>% 
+  filter(label != "") %>% 
+  select(Importance,label)
+
+dk_vi_area_followup <- data.frame(dk_vi_area_followup)
+
+ggplot(dk_vi_area_followup) +
+  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
+  labs(title = "Cortical Area")
+ggsave(filename="abcd_model_dk_area_importance.png",
+                         width=8, height=4, units='in', dpi=300)
+
+#Subcortical
+aseg_vi_followup <- abcd_vi_followup %>%
+  filter(label != "") %>% 
+  select(Importance,label)
+
+aseg_vi_followup <- data.frame(aseg_vi_followup)
+
+ggplot(aseg_vi_followup) +
+  geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+ggsave(filename="abcd_model_aseg_importance_coronal.png",
+                          width=7, height=4, units='in', dpi=300)
+
+ggplot(aseg_vi_followup) +
+  geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
+  theme_void() +
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+ggsave(filename="abcd_model_aseg_importance_sagittal.png",
+                          width=7, height=4, units='in', dpi=300)
 
@@
 # maybe multilevel model so that it's nested by participant? -
+
 # correlation for vlad's models
+# join baseline and follow-up corrected gaps by id
+brainage_both_waves <- left_join(brainage_pds_baseline_combat %>% select(src_subject_id, corrected_pred, corrected_gap) %>% rename(baseline_pred = corrected_pred, baseline_gap = corrected_gap), brainage_pds_followup_combat %>% select(src_subject_id, corrected_pred, corrected_gap) %>% rename(followup_pred = corrected_pred, followup_gap = corrected_gap))
+
+cor(brainage_both_waves$baseline_gap, brainage_both_waves$followup_gap, use="complete.obs")
+
+cor.test(brainage_both_waves$baseline_gap, brainage_both_waves$followup_gap, use="complete.obs")
 
 #correlation for abcd models
+# join baseline and follow-up corrected gaps by id
+brainage_abcd_both_waves <- left_join(abcd_brainage_pds_baseline_combat %>% select(src_subject_id, corrected_pred, corrected_gap) %>% rename(baseline_pred = corrected_pred, baseline_gap = corrected_gap), abcd_brainage_pds_followup_combat %>% select(src_subject_id, corrected_pred, corrected_gap) %>% rename(followup_pred = corrected_pred, followup_gap = corrected_gap))
 
+cor(brainage_abcd_both_waves$baseline_gap, brainage_abcd_both_waves$followup_gap, use="complete.obs")
+
+cor.test(brainage_abcd_both_waves$baseline_gap, brainage_abcd_both_waves$followup_gap, use="complete.obs")
+```
+
+```{r}
 # figure - probably spaghetti plot, show direction of gap change
+# is x-axis time, or age?
+# this might just be too many points to show? maybe fit a line over the top
+# vlad's model
+brainage_both_waves_long <- rbind(brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))
+
+ggplot(data = brainage_both_waves_long, aes(x = truth, y = corrected_gap, group = src_subject_id)) +
+    geom_line(size = .2, alpha=0.4)
+
+ggplot(data = brainage_both_waves_long, aes(x = wave, y = corrected_gap, group = src_subject_id)) +
+    geom_line(size = .2, alpha=0.4)
+
+# calculate difference, color by direction
+brainage_slope_count <- brainage_both_waves_long %>%
+  count(src_subject_id) %>% 
+  filter(n > 1)
+
+ggplot(data = brainage_both_waves_long %>% filter(src_subject_id %in% brainage_slope_count$src_subject_id) %>% 
+                              group_by(src_subject_id) %>%
+                mutate(slope = diff(corrected_gap)),
+       aes(x = truth, y = corrected_gap, group = src_subject_id, color = slope > 0)) +
+    geom_line(size = .2, alpha=0.4) +
+  labs(x= "Chronological Age", y = "Corrected Gap") +
+  scale_color_manual(name = "Direction of Change", labels = c("Negative", "Positive"), values = c("#E76F51",'#CCDDBB')) +
+  theme_minimal()
+
+# could also facet by slope to reduce lines
+
+# abcd model
+brainage_abcd_both_waves_long <- rbind(abcd_brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), abcd_brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))
+
+ggplot(data = brainage_abcd_both_waves_long, aes(x = truth, y = corrected_gap, group = src_subject_id)) +
+    geom_line(size = .2, alpha=0.4)
+
+ggplot(data = brainage_abcd_both_waves_long, aes(x = wave, y = corrected_gap, group = src_subject_id)) +
+    geom_line(size = .2, alpha=0.4)
+
+# calculate difference, color by direction
+brainage_abcd_slope_count <- brainage_abcd_both_waves_long %>%
+  count(src_subject_id) %>% 
+  filter(n > 1)
+
+ggplot(data = brainage_abcd_both_waves_long %>% filter(src_subject_id %in% brainage_abcd_slope_count$src_subject_id) %>% 
+                              group_by(src_subject_id) %>%
+                mutate(slope = diff(corrected_gap)),
+       aes(x = truth, y = corrected_gap, group = src_subject_id, color = slope > 0)) +
+    geom_line(size = .2, alpha=0.4) +
+  labs(x= "Chronological Age", y = "Corrected Gap") +
+  scale_color_manual(name = "Direction of Change", labels = c("Negative", "Positive"), values = c("#E76F51",'#CCDDBB')) +
+  theme_minimal()
+
+# '#568A7B' blue
 

### 2023-10-18 Revisions_Analyses&Figures.Rmd
Updated file for paper revisions - added longitudinal analyses, updated figures and tables. Added relevant demographics, figures, and tables for new follow-up model.

@@
 library(vip)
+```
+
+Load everything (corrected)
+```{r}
+load("~/Documents/FYP/Revisions/abcd_brainage_pds_baseline_combat.Rda")
+load("~/Documents/FYP/Revisions/abcd_brainage_pds_followup_combat.Rda")
+
+load("~/Documents/FYP/Revisions/brainage_pds_baseline_combat.Rda")
+load("~/Documents/FYP/Revisions/brainage_pds_followup_combat.Rda")
+
 ```
@@
 
+# fit metrics
+```{r}
+abcd_brainage_pds_followup_combat %>%
+  metrics(truth = truth, estimate = .pred)
+
+abcd_brainage_pds_followup_combat %>%
+  metrics(truth = truth, estimate = corrected_pred)
+
+cor(abcd_brainage_pds_followup_combat$truth, abcd_brainage_pds_followup_combat$corrected_gap)
+```
+
 # model features and importance
@@
 
+## testing
+load("~/Documents/FYP/UpdatedModels/fit_workflow_combat.rds")
+#model_obj <- xgb.load("~/Documents/FYP/UpdatedModels/abcd_model_obj_combat")
+#model_obj_indexed <- fit_workflow_combat$fit$fit$fit
+
+vi(fit_workflow_combat)
+
+colnames(df_train)[!(colnames(df_train) %in% vi(fit_workflow_combat)$Variable)]
+
+#xgb.importance(extract_fit_parsnip(fit_workflow_combat))
+#xgb.importance(model = model_obj)
+#xgb.importance(model = model_obj_indexed)
+
+#xgb.plot.importance(xgb.importance(model = model_obj))
+
+vip(fit_workflow_combat, num_features=20) +
+  theme_light()
+ggsave(filename="model2_vip_baseline.png",
+                          width=7, height=4, units='in', dpi=300)
+
+load("~/Documents/FYP/Revisions/fit_workflow_combat_followup.rds")
+#model_obj_followup <- xgb.load("~/Documents/FYP/Revisions/abcd_model_obj_combat_followup")
+#model_obj_indexed_followup <- fit_workflow_combat_followup$fit$fit$fit
+
+vi(fit_workflow_combat_followup)
+
+vip(fit_workflow_combat_followup, num_features=20) +
+  theme_light()
+ggsave(filename="model2_vip_followup.png",
+                          width=7, height=4, units='in', dpi=300)
 ## from model code
@@
 
-vip(abcd_model_obj_combat, num_features = 175)
-
-vip(abcd_model_obj_combat, num_features = 20)
-
-vi_updated <- vi(abcd_model_obj_combat)
+#vip(abcd_model_obj_combat, num_features = 175) 
+#vip(abcd_model_obj_combat, num_features = 20) 
+#vi_updated <- vi(abcd_model_obj_combat)
+#vi(fit_workflow_combat)
+#vi(fit_workflow_combat, method = "shap")
 
@@
 
-vi_updated2 <- vi(extract_fit_parsnip(fit_workflow_combat))
-
-#vi_model(abcd_model_obj_combat)
-
+#vi(fit_workflow_combat$fit$fit)
+#vi(fit_workflow_combat$fit$fit$fit)
+
+vi_updated <- vi(extract_fit_parsnip(fit_workflow_combat)) # this seems to work!
+
+vip(extract_fit_parsnip(fit_workflow_combat), num_features = 20)
+
+```
+
+ggseg importance plots
+```{r}
 # ggseg importance plots
-# Rename columns to fit with ggseg atlas (ugghhh)
+# Rename columns to fit with ggseg atlas
 
@@
 
-abcd_vi <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed)
+#abcd_vi <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed)
+abcd_vi <- left_join(vi_updated, abcd_vi_renamed)
 
@@
   labs(title = "Cortical Volume")
-ggsave(filename="abcd_model_dk_vol_importance.png",
+ggsave(filename="abcd_model_dk_vol_importance_combat.png",
                          width=8, height=4, units='in', dpi=300)
@@
   labs(title = "Cortical Area")
-ggsave(filename="abcd_model_dk_area_importance.png",
+ggsave(filename="abcd_model_dk_area_importance_combat.png",
                          width=8, height=4, units='in', dpi=300)
@@
 #Subcortical
-aseg_vi <- abcd_vi_renamed %>%
+aseg_vi <- abcd_vi %>%
   filter(label != "") %>% 
@@
   scale_fill_gradient(low = '#FDD872', high = '#E76F51')
-ggsave(filename="abcd_model_aseg_importance_coronal.png",
+ggsave(filename="abcd_model_aseg_importance_coronal_combat.png",
                           width=7, height=4, units='in', dpi=300)
@@
   scale_fill_gradient(low = '#FDD872', high = '#E76F51')
-ggsave(filename="abcd_model_aseg_importance_sagittal.png",
+ggsave(filename="abcd_model_aseg_importance_sagittal_combat.png",
                           width=7, height=4, units='in', dpi=300)
@@
 
-load("~/Documents/FYP/abcd_brainage_pds_followup_combat.Rda")
-
 ```
@@
 
-# maybe make a table of results
-```{r}
+# Table of results
+```{r}
+
+apa.reg.table(baseline_puberty_y_mod1, filename= "apa_baseline_puberty_y_mod1.doc", table.number = 3)
+apa.reg.table(baseline_puberty_p_mod1, filename= "apa_baseline_puberty_p_mod1.doc", table.number = 4)
+apa.reg.table(baseline_cognition_mod1, filename= "apa_baseline_cognition_mod1.doc", table.number = 5)
+
+apa.reg.table(followup_puberty_y_mod1, filename= "apa_followup_puberty_y_mod1.doc", table.number = 6)
+apa.reg.table(followup_puberty_p_mod1, filename= "apa_followup_puberty_p_mod1.doc", table.number = 7)
+
+apa.reg.table(baseline_puberty_y_mod2, filename= "apa_baseline_puberty_y_mod2.doc", table.number = 8)
+apa.reg.table(baseline_puberty_p_mod2, filename= "apa_baseline_puberty_p_mod2.doc", table.number = 9)
+apa.reg.table(baseline_cognition_mod2, filename= "apa_baseline_cognition_mod2.doc", table.number = 10)
+
+apa.reg.table(followup_puberty_y_mod2, filename= "apa_followup_puberty_y_mod2.doc", table.number = 11)
+apa.reg.table(followup_puberty_p_mod2, filename= "apa_followup_puberty_p_mod2.doc", table.number = 12)
+
+apa.reg.table(baseline_puberty_y_mod1,baseline_puberty_p_mod1, baseline_cognition_mod1,  filename= "apa_baseline_mod1.doc", table.number = 3)
 
@@
 standardize(followup_puberty_p_mod2)
+
 ```
@@
 summary(lm.beta(followup_puberty_y_mod2))
+
+apa.reg.table(lm.beta(followup_puberty_y_mod2))
+
+apa.reg.table(followup_puberty_y_mod2, filename= "apa_followup_puberty_y_mod2.doc", table.number = 1)
 ```
@@
 ```
-# New Demographics Column for Follow- Up (Model 2)
-```{r}
-nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "F")) / nrow(abcd_brainage_pds_followup_combat)
-nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "M")) / nrow(abcd_brainage_pds_followup_combat)
-
-demographics <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/pdem02.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(src_subject_id, contains("demo_race"), demo_ethn_v2, demo_ethn2_v2, demo_comb_income_v2)
-
-demographics_baseline_abcdmodel <- left_join(abcd_baseline, demographics)
+
+# recalculate demographics table
+```{r}
+demographics_baseline <- left_join(brainage_pds_baseline_combat, demographics)
+
+demographics_followup <- left_join(brainage_pds_followup_combat, demographics)
+
+demographics_baseline_abcdmodel <- left_join(abcd_brainage_pds_baseline_combat, demographics)
 
@@
 
+demographics_baseline <- left_join(demographics_baseline, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
+
+demographics_followup <- left_join(demographics_followup, demographics_long %>% filter(eventname == "2_year_follow_up_y_arm_1"))
+
 demographics_baseline_abcdmodel <- left_join(demographics_baseline_abcdmodel, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
 
-#family income
-table(demographics_baseline$demo_comb_income_v2)
-
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+```
+# Baseline (model 2)
+```{r}
 #Overall values
 #Non-hispanic white
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
@@
 #Non-hispanic Black
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
@@
 #Hispanic
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(demo_ethn_v2 == 1)
 
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(((demo_race_a_p___10 == 1 | demo_race_a_p___11 == 1) & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1) & demo_ethn_v2 == 1)
@@
 #Asian
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
@@
 #Native American Alaska Native
-demographics_baseline %>% 
+demographics_baseline_abcdmodel %>% 
   filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
@@
 
-demographics_baseline %>% 
-  mutate_at(c(18:33), as.numeric) %>% 
+demographics_baseline_abcdmodel %>% 
+  mutate_at(c(12:33), as.numeric) %>% 
   mutate(
@@
 
+demographics_baseline_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+
 demographics_baseline %>% 
@@
 
-```
-
-# sex differences
-```{r}
-# model with sex as covariate?
+demographics_followup %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+```
+
+# New Demographics Column for Follow- Up (Model 2)
+```{r}
+nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "F")) / nrow(abcd_brainage_pds_followup_combat)
+nrow(abcd_brainage_pds_followup_combat %>% filter(sex == "M")) / nrow(abcd_brainage_pds_followup_combat)
+
+demographics <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/pdem02.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, contains("demo_race"), demo_ethn_v2, demo_ethn2_v2, demo_comb_income_v2)
+
+demographics_followup_abcdmodel <- left_join(abcd_brainage_pds_followup_combat, demographics)
+
+demographics_long <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_lpds01.txt") %>%
+  filter(!collection_title=="collection_title") %>% 
+  select(src_subject_id, eventname, demo_comb_income_v2_l)
+
+demographics_followup_abcdmodel <- left_join(demographics_followup_abcdmodel, demographics_long %>% filter(eventname == "2_year_follow_up_y_arm_1"))
+
+#family income
+table(demographics_followup_abcdmodel$demo_comb_income_v2)
+# /3693 for %
+
+#Overall values
+#Non-hispanic white
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Non-hispanic Black
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Hispanic
+demographics_followup_abcdmodel %>% 
+  filter(demo_ethn_v2 == 1)
+
+demographics_followup_abcdmodel %>% 
+  filter(((demo_race_a_p___10 == 1 | demo_race_a_p___11 == 1) & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1) & demo_ethn_v2 == 1)
+
+#Asian
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)
+
+#Native American Alaska Native
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )
+
+#Multiracial
+
+# mutate column: true if participant marked any asian category
+# mutate column: true if participant marked native american/alaska native
+# sum white, black, asian, native american/alaska native, and additional races
+
+demographics_followup_abcdmodel %>% 
+  mutate_at(c(12:33), as.numeric) %>% 
+  mutate(
+    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
+                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
+         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
+                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
+         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>%
+  filter(race_total > 1)
+
+#additional races
+#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.
+
+demographics_followup_abcdmodel %>% 
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
 
@@
 
-# Fig 1 with line of identity, jitter fixed
-```{r}
-# fig 1 has different jitter for ab vs. cd. x axis scaling of ab and cd should be equal
-# add line of identity to fig 1 to vizualize bias - already done for AB, but CD has a regression line, do they want a 1:1 line on that too?
-```
-
 # Plot of model 1 & 2 prediction overlaid (baseline)
@@
 
-# change colors
-# decide if it should only show the same participants
+ggsave(filename="model1_2_overlaid_baseline.png",
+                          width=7, height=4, units='in', dpi=300)
+
 ```
@@
   theme(text = element_text(size = 14))  
-
-# change colors
-# decide if it should only show the same participants
+ggsave(filename="model1_2_overlaid_followup.png",
+                          width=7, height=4, units='in', dpi=300)
+
 ```
@@
 ```
-```{r}
-
-####### UPDATE COLORS
+
+```{r}
 
@@
 ```{r}
-# Prediction Plots
-
-```
-
-```{r}
-# Model feature - baseline
+# Model feature - follow-up
 
 # vip
-
-# ggseg
-
-```
-
-```{r}
-# Model feature - follow-up
-
-# vip
-abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
+#abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
 
@@
 
-vip(abcd_model_obj_combat_followup, num_features = 20)
-
-vi(fit_workflow_combat_followup$fit$fit$fit)
+#vip(abcd_model_obj_combat_followup, num_features = 20)
+#vi(fit_workflow_combat_followup$fit$fit$fit)
 
@@
 
-abcd_vi_followup <- left_join(vi(abcd_model_obj_combat_followup), abcd_vi_renamed)
+abcd_vi_followup <- left_join(vi(fit_workflow_combat_followup), abcd_vi_renamed)
 
@@
 
-abcd_vi_dk_followup <- left_join(vi(abcd_model_obj_combat_followup), abcd_vi_renamed_dk)
+abcd_vi_dk_followup <- left_join(vi(fit_workflow_combat_followup), abcd_vi_renamed_dk)
 
@@
   labs(title = "Cortical Volume")
-ggsave(filename="abcd_model_dk_vol_importance.png",
+ggsave(filename="abcd_model_dk_vol_importance_followup.png",
                          width=8, height=4, units='in', dpi=300)
@@
   labs(title = "Cortical Area")
-ggsave(filename="abcd_model_dk_area_importance.png",
+ggsave(filename="abcd_model_dk_area_importance_followup.png",
                          width=8, height=4, units='in', dpi=300)
@@
   scale_fill_gradient(low = '#FDD872', high = '#E76F51')
-ggsave(filename="abcd_model_aseg_importance_coronal.png",
+ggsave(filename="abcd_model_aseg_importance_coronal_followup.png",
                           width=7, height=4, units='in', dpi=300)
@@
   scale_fill_gradient(low = '#FDD872', high = '#E76F51')
-ggsave(filename="abcd_model_aseg_importance_sagittal.png",
+ggsave(filename="abcd_model_aseg_importance_sagittal_followup.png",
                           width=7, height=4, units='in', dpi=300)
@@
 
-```{r}
-# correlation between gaps from waves, or something more complicated?
-# maybe multilevel model so that it's nested by participant? -
-
-# correlation for vlad's models
+# Correlation between BrainAGE at both waves 
+```{r}
+
+# correlation for Model 1
 # join baseline and follow-up corrected gaps by id
@@
 
-```{r}
-# figure - probably spaghetti plot, show direction of gap change
+ICC for BrainAGE at both waves
+```{r}
+# Model 1
+# fit unconditional/intercept only model
+icc_model_mod1 <- lmer(corrected_gap ~ 1 + (1|src_subject_id) , data = brainage_both_waves_long)
+
+summary(icc_model_mod1)
+
+# calculate ICC from output
+# between-person variance / (between-person variance + within-person variance)
+2.415 / (2.415 + 1.034)
+
+# ABCD Model
+# fit unconditional/intercept only model
+icc_model_mod2 <- lmer(corrected_gap ~ 1 + (1|src_subject_id) , data = brainage_abcd_both_waves_long)
+
+summary(icc_model_mod2)
+
+# calculate ICC from output
+# between-person variance / (between-person variance + within-person variance)
+3.154 / (3.154+2.795)
+```
+
+Plot of longitudinal BrainAGE
+```{r}
+# Spaghetti plot, show direction of gap change
 # is x-axis time, or age?
@@
 
+ggsave(filename="model1_longitudinal.png",
+                          width=7, height=4, units='in', dpi=300)
+
 # could also facet by slope to reduce lines
@@
 
-# '#568A7B' blue
-
-# for each set of models, look at change in brainage vs chronological age?
-
-```
+ggsave(filename="model2_longitudinal.png",
+                          width=7, height=4, units='in', dpi=300)
+
+```

### 2023-11-16 Difference Scores /OldVersions/abcd_diff_model_talapas.R
Removed files related to difference scores - will become a separate repository for new project.

@@ -1,174 +1,0 @@
-
-library(dplyr)
-library(parallel)
-library(doParallel)
-library(tidymodels)
-library(xgboost)
-
-#load data
-
-
-load("/projects/dcnlab/shared/brainage/sample_train_diff.Rda")
-
-# Model prep: split, preprocessing, CV ------------------------------------
-
-# Train / test split ------------------------------------------------------
-set.seed(42)
-df_split_diff <- initial_split(
-  sample_train_diff, 
-  prop = 0.80,
-  # matching age distributions across train and test set
-  strata = "age_midpoint"
-)
-
-df_train_diff <- training(df_split_diff)
-df_validation_diff <- testing(df_split_diff)
-
-# report n training, features, testing
-#df_train_baseline %>% 
-df_train_diff%>%
-  summarise(
-    n_training = n(),
-    p_features = ncol(df_train_diff)
-  ) %>%
-  bind_cols(df_validation_diff %>%
-              summarise(n_testing = n()))
-
-# Pre-processing setup ----------------------------------------------------
-
-# define (what we want to do)
-preprocess_recipe_diff <- df_train_diff %>%
-#preprocess_recipe_baseline <- train_test %>%
-  # predict scan age by all brain features
-  recipe(age_midpoint ~ .) %>%
-  # remove near zero variance predictors
-  step_nzv(all_predictors()) %>%
-  prep() # where it all gets calculated
-
-preprocess_recipe_diff
-
-
-# Apply pre-processing ----------------------------------------------------
-
-# juice() will work with training data, `bake()` to apply this to our test data
-
-# apply on train (gives processed value)
-df_train_prep_diff <- juice(preprocess_recipe_diff)
-
-# apply on validation
-df_validation_prep_diff <- preprocess_recipe_diff %>% bake(df_validation_diff)
-#df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(validation_test)
-
-
-# Cross-validation --------------------------------------------------------
-
-# 10 fold cv repeated 10 times
-set.seed(42)
-train_cv_diff <- df_train_prep_diff %>%
-  vfold_cv(
-    v = 10, 
-    repeats = 10, 
-    strata = age_midpoint
-  )
-
-
-#Train and Fit Model (Baseline)
-
-# Extreme gradient boosting
[…]

### 2023-11-16 PaperAnalyses&Figures.Rmd
Cleaned up original script, and added code for descriptive statistics. For revision script, updated code for variable importance, edited figures to save in tiff format and arrange in proper formatting for manuscript. In model training script, updated model names for clarity after incorporation of ComBat pipeline.

@@
                "apaTables", "MetBrewer",
-               "ggseg", "vip", "tidymodels")
+               "ggseg", "vip", "tidymodels", "psych")
 if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
@@
 **Vlad's Model, Baseline Analyses**
+
 Run Models
@@
 
-summary(baseline_cognition_zscore_1_m)
-summary(baseline_cognition_zscore_2_m)
+summary(baseline_cognition_standard_1)
+summary(baseline_cognition_standard_2)
+
+summary(baseline_cognition_zscore_1)
+summary(baseline_cognition_zscore_2)
 
@@
 summary(baseline_cognition_standard_2_m)
+
+```
+
+Descriptive Stats
+```{r}
+describe(sample_1_baseline_months)
+
+describe(sample_2_baseline_months)
+
+nrow(sample_1_baseline_months %>% filter(sex == "F"))
+
+nrow(sample_2_baseline_months %>% filter(sex == "F"))
 
@@
 ```{r}
-table(brainage_pds_baseline$sex)
-
 #white
@@
 table(demographics_baseline$demo_ethn_v2)
+```
+
+```{r}
+#sex
+table(brainage_pds_baseline$sex)
 
@@
   filter(demo_ethn_v2 == 1)
+
+demographics_baseline %>% 
+  filter(((demo_race_a_p___10 == 1 | demo_race_a_p___11 == 1) & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1) & demo_ethn_v2 == 1)
 
@@
 demographics_baseline %>% 
-  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
+  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)
 
@@
 ```{r}
+# Youth-report puberty
 summary(followup_puberty_y_mean_1)
@@
 
+# Parent-report puberty
 summary(followup_puberty_p_mean_1)
@@
 summary(followup_cognition_crystal_2)
+```
+
+Descriptive Stats
+```{r}
+describe(sample_1_followup_months)
+
+describe(sample_2_followup_months)
 ```
@@
 
-summary(abcd_baseline_cognition)
+#summary(abcd_baseline_cognition)
 
@@
 ```
+
+Descriptive Stats
+```{r}
+describe(abcd_baseline_months)
+```
+
 Model Plots

### 2023-11-16 Revisions_Analyses&Figures.Rmd
Cleaned up original script, and added code for descriptive statistics. For revision script, updated code for variable importance, edited figures to save in tiff format and arrange in proper formatting for manuscript. In model training script, updated model names for clarity after incorporation of ComBat pipeline.

@@
 
-Install combat
-```{r}
-library(devtools)
-install_github("jfortin1/neuroCombatData")
-install_github("jfortin1/neuroCombat_Rpackage")
+Load Packages
+```{r}
+#library(devtools)
+#install_github("jfortin1/neuroCombatData")
+#install_github("jfortin1/neuroCombat_Rpackage")
 
@@
 
-vip(fit_workflow_combat, num_features=20) +
+vip_abcd_baseline <- vip(fit_workflow_combat, num_features=20) +
   theme_light()
+
 ggsave(filename="model2_vip_baseline.png",
@@
 
-vip(fit_workflow_combat_followup, num_features=20) +
+vip_abcd_followup <- vip(fit_workflow_combat_followup, num_features=20) +
   theme_light()
@@
 
-abcd_vi_dk <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed_dk)
+#abcd_vi_dk <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed_dk)
+
+abcd_vi_dk <- left_join(vi_updated, abcd_vi_renamed_dk)
 
@@
 
-ggplot(dk_vi_vol) +
+abcd_model_dk_vol_importance_baseline <- ggplot(dk_vi_vol) +
   geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
@@
   labs(title = "Cortical Volume")
+
 ggsave(filename="abcd_model_dk_vol_importance_combat.png",
@@
 
-ggplot(dk_vi_area) +
+abcd_model_dk_area_importance_baseline <- ggplot(dk_vi_area) +
   geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
@@
   labs(title = "Cortical Area")
+
 ggsave(filename="abcd_model_dk_area_importance_combat.png",
@@
 
-ggplot(aseg_vi) +
+abcd_model_aseg_importance_coronal_baseline <- ggplot(aseg_vi) +
   geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
   theme_void() +
-  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51')  +
+  labs(title = "Subcortical Volume")
 ggsave(filename="abcd_model_aseg_importance_coronal_combat.png",
@@
 
-ggplot(aseg_vi) +
+abcd_model_aseg_importance_sagittal_baseline <- ggplot(aseg_vi) +
   geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
@@
 
+# make one aseg figure
+abcd_model_aseg_importance_baseline <- ggarrange(abcd_model_aseg_importance_coronal_baseline, abcd_model_aseg_importance_sagittal_baseline, 
+          ncol = 1, nrow = 2)
+
+# Save as final figure
+ggarrange(vip_abcd_baseline, abcd_model_dk_vol_importance_baseline, abcd_model_dk_area_importance_baseline, abcd_model_aseg_importance_baseline, 
+          labels = c("A", "B", "C", "D"),
+          ncol = 2, nrow = 2)
+ggsave(filename = "mod2_vip_ggseg.tiff",
+                          width=12, height=12, units='in', dpi=300)
+
+```
+
+```{r}
 #troubleshooting names
@@
 
+ggsave(filename = "pred_bias_mod1_combat.tiff",
+                          width=10, height=10, units='in', dpi=300)
+
 ## Baseline - Model 2
@@
 ggsave(filename = "pred_bias_mod2_combat.png",
+                          width=10, height=10, units='in', dpi=300)
+
+ggsave(filename = "pred_bias_mod2_combat.tiff",
                           width=10, height=10, units='in', dpi=300)
@@
   theme(text = element_text(size = 12))
+
 ggsave(filename="baseline_youth_pds_mod1_combat.png",
@@
 
-```
+ggsave(filename = "distributions_baseline_mod1_combat.tiff",
+                          width=7, height=5, units='in', dpi=300)
+
+```
+
 ```{r}
@@
           ncol = 2, nrow = 1)
+
 ggsave(filename = "distributions_followup_mod1_combat.png",
+                          width=10, height=4, units='in', dpi=300)
+
+ggsave(filename = "distributions_followup_mod1_combat.tiff",
                           width=10, height=4, units='in', dpi=300)
@@
 
+ggsave(filename = "distributions_baseline_mod2_combat.tiff",
+                          width=7, height=5, units='in', dpi=300)
+
 ```
@@
 
+ggsave(filename = "distributions_followup_mod2_combat.tiff",
+                          width=10, height=4, units='in', dpi=300)
+
 ```
@@
 
-vip(fit_workflow_combat_followup, num_features = 20)
+vip_abcd_followup <- vip(fit_workflow_combat_followup, num_features = 20)
 
@@
 
-ggplot(dk_vi_vol_followup) +
+abcd_model_dk_vol_importance_followup <- ggplot(dk_vi_vol_followup) +
   geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
@@
 
-ggplot(dk_vi_area_followup) +
+abcd_model_dk_area_importance_followup <-ggplot(dk_vi_area_followup) +
   geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
@@
 
-ggplot(aseg_vi_followup) +
+abcd_model_aseg_importance_coronal_followup <- ggplot(aseg_vi_followup) +
   geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
   theme_void() +
-  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
+  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
+  labs(title = "Subcortical Volume")
 ggsave(filename="abcd_model_aseg_importance_coronal_followup.png",
@@
 
-ggplot(aseg_vi_followup) +
+abcd_model_aseg_importance_sagittal_followup <- ggplot(aseg_vi_followup) +
   geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
@@
                           width=7, height=4, units='in', dpi=300)
+
+# make one aseg figure
+abcd_model_aseg_importance_followup <- ggarrange(abcd_model_aseg_importance_coronal_followup, abcd_model_aseg_importance_sagittal_followup, 
+          ncol = 1, nrow = 2)
+
+# Save as final figure
+ggarrange(vip_abcd_followup, abcd_model_dk_vol_importance_followup, abcd_model_dk_area_importance_followup, abcd_model_aseg_importance_followup, 
+          labels = c("A", "B", "C", "D"),
+          ncol = 2, nrow = 2)
+ggsave(filename = "mod2_vip_ggseg_followup.tiff",
+                          width=12, height=12, units='in', dpi=300)
 
@@
 
-ggplot(data = brainage_both_waves_long %>% filter(src_subject_id %in% brainage_slope_count$src_subject_id) %>% 
+mod1_longitudinal <- ggplot(data = brainage_both_waves_long %>% filter(src_subject_id %in% brainage_slope_count$src_subject_id) %>% 
                               group_by(src_subject_id) %>%
@@
 
-ggplot(data = brainage_abcd_both_waves_long %>% filter(src_subject_id %in% brainage_abcd_slope_count$src_subject_id) %>% 
+mod2_longitudinal <- ggplot(data = brainage_abcd_both_waves_long %>% filter(src_subject_id %in% brainage_abcd_slope_count$src_subject_id) %>% 
                               group_by(src_subject_id) %>%
@@
 
-```
+# Save as final figure
+ggarrange(mod1_longitudinal, mod2_longitudinal, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+ggsave(filename = "longitudinal_plot.tiff",
+                          width=12, height=4, units='in', dpi=300)
+
+```

### 2023-11-16 UpdatedModels/ABCDModelTraining.Rmd
Cleaned up original script, and added code for descriptive statistics. For revision script, updated code for variable importance, edited figures to save in tiff format and arrange in proper formatting for manuscript. In model training script, updated model names for clarity after incorporation of ComBat pipeline.

@@
 ```{r}
-load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/training_sample_baseline.Rds")
+#load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/training_sample_baseline.Rds")
+load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/model2_features.Rds")
+
+model2_features_noid <- model2_features %>% 
+  select(-c(src_subject_id))
 
@@
 df_split <- initial_split(
-  training_sample_baseline, 
+  model2_features_noid, 
   prop = 0.80,
@@
 # select parsimonious params within one SE of best model
-best_xgb_params_test <- xgb_tuned_results_test %>%
+best_xgb_params <- xgb_tuned_results_test %>%
   select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)
@@
 ```{r}
-final_xgb <- finalize_workflow(
+final_xgb_combat <- finalize_workflow(
   xgb_wf,
-  best_xgb_params_test
+  best_xgb_params
 )
 
-final_xgb
+final_xgb_combat
 
-fit_workflow <- fit(final_xgb, df_train_prep)
+fit_workflow_combat <- fit(final_xgb_combat, df_train_prep)
 
@@
 
-final_xgb %>%
+final_xgb_combat %>%
   fit(data = df_train) %>%
@@
 ```{r}
-final_res <- last_fit(final_xgb, df_split)
+final_res <- last_fit(final_xgb_combat, df_split)
 
@@
 # 1. Rds
-save(fit_workflow, file = "fit_workflow.rds")
+save(fit_workflow_combat, file = "fit_workflow_combat.rds")
 # 2. rda
-save(fit_workflow, file = "fit_workflow_2.rda")
+save(fit_workflow_combat, file = "fit_workflow_2_combat.rda")
 # 3. raw xbg object, which can then be loaded with xgb.load
-model_obj <- fit_workflow$fit$fit$fit
-xgb.save(model_obj, "model_obj")
+abcd_model_obj_combat <- fit_workflow_combat$fit$fit$fit
+xgb.save(abcd_model_obj_combat, "abcd_model_obj_combat")
 ```
@@
 
-brain_age_df_test <- as.data.frame(brain_age_df_test) %>%
+# combat version
+brain_age_df_combat <-
+  abcd_model_obj_combat %>%
+  predict(newdata = as.matrix(model2_analysis %>% select(-c(src_subject_id, interview_age))))
+
+brain_age_df_combat <- as.data.frame(brain_age_df_combat) %>%
   mutate(
     # provide the chronological age at time of scan
-    truth = analysis_sample_baseline$interview_age
+    truth = model2_analysis$interview_age
   ) %>%
@@
 
-brain_age_df_workflow <- fit_workflow %>%
-  predict(new_data = analysis_sample) %>%
+brain_age_df_workflow_combat <- fit_workflow_combat %>%
+  predict(new_data = model2_analysis) %>%
   mutate(
     # provide the chronological age at time of scan
-    truth = analysis_sample$interview_age
+    truth = model2_analysis$interview_age
   ) %>%
@@
 abcd_brain_age_correction <-
-  fit_workflow %>%
+  fit_workflow_combat %>%
   predict(new_data = df_validation) %>%
@@
 #Baseline
-brain_age_corrected_df <- brain_age_df %>%
+brain_age_corrected_df_combat <- brain_age_df_workflow_combat %>%
   mutate(
@@
 ```
+
+Clean up & save
+```{r}
+# add IDs back
+brain_age_corrected_df_combat$src_subject_id <- model2_analysis$src_subject_id
+
+brain_age_corrected_df_combat <- brain_age_corrected_df_combat %>% 
+  select(6, 1:5)
+
+# save dataframe
+save(brain_age_corrected_df_combat, file = "brain_age_corrected_df_combat.Rda")
+
+```

### 2023-11-16 ABCDModel.Rmd
Updated ABCD model prediction script with new model name.

@@
 abcd_brainage_model <- readRDS(here("UpdatedModels", "abcd_fit_workflow.rds"))
+abcd_brainage_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_fit_workflow.Rds")
+
+load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_fit_workflow.Rds")
 ```
@@
 brain_age_df_baseline <-
-  abcd_brainage_model %>%
+  fit_workflow %>%
   predict(new_data = analysis_sample_baseline) %>%

### 2023-11-16 UpdatedModels/ABCDModelTraining_FollowUp.Rmd
Added model training script for follow-up data.

---
title: "Train BrainAGE with tidymodels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```

```{r}
library(tidymodels)
library(xgboost)
```

```{r}
#load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/training_sample_baseline.Rds")
load("~/Documents/FYP/Revisions/model2_features_followup.Rda")

model2_features_followup_noid <- model2_features_followup %>% 
  select(-c(src_subject_id))

#training_subset_test <- training_sample_baseline[1:100,]

# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split_followup <- initial_split(
  model2_features_followup_noid, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)
df_train_followup <- training(df_split_followup)
df_validation_followup <- testing(df_split_followup)

```

# Pre-processing setup ----------------------------------------------------

```{r}
# define (what we want to do)
preprocess_recipe_followup <- df_train_followup %>%
  # predict scan age by all brain features
  recipe(interview_age ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe_followup

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_followup_prep <- juice(preprocess_recipe_followup)

# apply on validation
df_validation_followup_prep <- preprocess_recipe %>% bake(df_validation_followup)

```

```{r}
boost_mod_followup <- boost_tree(
  mode = "regression", 
  trees = 150, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

```

```{r}
set.seed(42)

xgboost_grid_followup <- grid_latin_hypercube(
  min_n(), 
  tree_depth(), 
  loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_followup_prep),
  learn_rate(),
  size = 500 
)

```

```{r}
xgb_wf_followup <- workflow() %>%
  add_formula(interview_age ~ .) %>%
  add_model(boost_mod_followup)

xgb_wf_followup
```

```{r}
set.seed(42)

train_cv_followup <- df_train_followup_prep %>%
  vfold_cv(
    v = 10, 
    repeats = 10, 
    strata = interview_age
  )
```

```{r}
doParallel::registerDoParallel()

set.seed(42)

xgb_tuned_results_followup <- tune_grid(
  xgb_wf_followup,
  resamples = train_cv_followup,
  grid = xgboost_grid_followup,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = TRUE,
                         save_pred = TRUE)
)

xgb_tuned_results_followup
```

```{r}
xgb_tuned_results_followup %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params_followup <- xgb_tuned_results_followup %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

```

Finalize workflow and fit final model
```{r}
final_xgb_combat_followup <- finalize_workflow(
  xgb_wf_followup,
  best_xgb_params_followup
)

final_xgb_combat_followup

fit_workflow_combat_followup <- fit(final_xgb_combat_followup, df_train_followup_prep)

#xgb_final_mod_test <- boost_mod_test %>%
 # finalize_model(best_xgb_params_test) %>%
 # fit(interview_age ~ .,
#      data = df_train_prep)

```

Variable importance, ignore
```{r}
library(vip)

final_xgb_combat_followup %>%
  fit(data = df_train_followup) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

```

```{r}
final_res_followup <- last_fit(final_xgb_combat_followup, df_split_followup)

collect_metrics(final_res_followup)
```

Save models
```{r}
# three different save formats just in case
# 1. Rds
save(fit_workflow_combat_followup, file = "fit_workflow_combat_followup.rds")
# 2. rda
save(fit_workflow_combat_followup, file = "fit_workflow_2_combat_followup.rda")
# 3. raw xbg object, which can then be loaded with xgb.load
abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
xgb.save(abcd_model_obj_combat_followup, "abcd_model_obj_combat_followup")
```

Model bundling, ignore for now
```{r}
# Testing out bundling
#mod_bundle <- bundle::bundle(fit_workflow)
#save(mod_bundle, file = "xgb_workflow_bundles.rds")

#save(mod_bundle, file = "xgb_workflow.rds")

# in a new R session:
#mod_bundle <- readRDS("path/to/file.rds")
#mod_new <- unbundle(mod_bundle)

#xgb.save.raw(fit_workflow$fit$fit, raw_format = "deprecated")

#model_obj <- fit_workflow$fit$fit$fit
#xgb.save(model_obj, "model_obj")

#xgb.save.raw(model_obj, "model_obj_raw")
```

Predict
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/Revisions/model2_analysis_followup.Rds")

# combat version
brain_age_df_combat_followup <-
  abcd_model_obj_combat_followup %>%
  predict(newdata = as.matrix(model2_analysis_followup %>% select(-c(src_subject_id, interview_age))))

brain_age_df_combat_followup <- as.data.frame(brain_age_df_combat_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = model2_analysis_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#workflow_predict <- predict(fit_workflow, analysis_sample)

brain_age_df_workflow_combat_followup <- fit_workflow_combat_followup %>%
  predict(new_data = model2_analysis_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = model2_analysis_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Bias Correction
```{r}
#run model on validation set
abcd_brain_age_correction_followup <-
  fit_workflow_combat_followup %>%
  predict(new_data = df_validation_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction_followup %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod_followup <- lm(.pred ~ truth, data = abcd_brain_age_correction_followup)

# extract intercept and slope
#
abcd_bias_intercept_followup <- abcd_bias_mod_followup$coefficients[["(Intercept)"]]

abcd_bias_slope_followup <-  abcd_bias_mod_followup$coefficients[["truth"]]

# create bias correct data frame
#Baseline
brain_age_corrected_df_combat_followup <- brain_age_df_workflow_combat_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_followup) / abcd_bias_slope_followup
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

```

Clean up & save
```{r}
# add IDs back
brain_age_corrected_df_combat_followup$src_subject_id <- model2_analysis_followup$src_subject_id

brain_age_corrected_df_combat_followup <- brain_age_corrected_df_combat_followup %>% 
  select(6, 1:5)

# save dataframe
save(brain_age_corrected_df_combat_followup, file = "brain_age_corrected_df_combat_followup.Rda")

```

### 2023-11-16 Revisions_Analyses&Figures.Rmd
Updated model prediction code for follow-up, updated readme, cleaned up analysis script.

@@
 
+Optional feature name troubleshooting for ggseg
 ```{r}
@@
 ### Rerun old figures
+# Distributions - Baseline (Model 1)
 ```{r}
@@
 
+# Distributions - Follow-Up (Model 1)
 ```{r}
@@
 
+# Distributions - Baseline (Model 2)
 ```{r}
@@
 
-```{r}
-
+# Distributions - Follow-Up (Model 2)
+```{r}
 #Youth-Report Puberty
@@
 
-```{r}
-# Model feature - follow-up
+# Model features - follow-up
+```{r}
 
@@
 
-figure out how many baseline training v analysis are in follow-up
+Check how many baseline training v analysis are in follow-up
 ```{r}
@@
 # Spaghetti plot, show direction of gap change
-# is x-axis time, or age?
-# this might just be too many points to show? maybe fit a line over the top
-# vlad's model
+
+# Existing model
 brainage_both_waves_long <- rbind(brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))
@@
 
-# could also facet by slope to reduce lines
-
-# abcd model
+# ABCD model
 brainage_abcd_both_waves_long <- rbind(abcd_brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), abcd_brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))

### 2023-11-16 UpdatedModels/ABCDModelTraining_FollowUp.Rmd
Updated model prediction code for follow-up, updated readme, cleaned up analysis script.

@@
 
+brain_age_df_workflow_combat_followup <- fit_workflow_combat_followup %>%
+  predict(new_data = model2_analysis_followup) %>%
+  mutate(
+    # provide the chronological age at time of scan
+    truth = model2_analysis_followup$interview_age
+  ) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
+```
+
+Alternate/Backup Prediction - Use if prior version doesn't work
+```{r}
+# The following version converts the workflow object to a raw xgb object.
+Over the course of the project, we've discovered that certain features in the xgboost package haven't always worked with updates within tidymodels.
+In case of deprecation of certain functions/ (or issues with version compatibility), this code should let you access the actual xgboost object, which can then be used to make model predictions.
+I don't anticipate this will be needed, but if the code on lines 221-233 doesn't work, this might!
+
+abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
+
 # combat version
@@
 
-#workflow_predict <- predict(fit_workflow, analysis_sample)
-
-brain_age_df_workflow_combat_followup <- fit_workflow_combat_followup %>%
-  predict(new_data = model2_analysis_followup) %>%
-  mutate(
-    # provide the chronological age at time of scan
-    truth = model2_analysis_followup$interview_age
-  ) %>%
-  # compute the brain age gap by subtracting chronological age from prediction
-  mutate(gap = .pred - truth)
 ```

### 2023-11-16 FreeSurferSetupBrainAGE.Rmd
Added script for converting names from freesurfer output, updated readme.

---
title: "NavNetworks BrainAGE"
author: "Lucy Whitmore"
date: "4/3/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "nlme", "lme4",
               "ggpubr", "lubridate", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Import & combine data
```{r}
left_area <- rio::import("left_area.txt")  %>% 
  rename(subid = lh.aparc.area)

right_area <- rio::import("right_area.txt")   %>% 
  rename(subid = rh.aparc.area)

right_volume <- rio::import("right_volume.txt")   %>% 
  rename(subid = rh.aparc.volume)

left_volume<- rio::import("left_volume.txt")   %>% 
  rename(subid = lh.aparc.volume)

aseg_volume <- rio::import("aseg_volume.txt")  %>% 
  rename(subid = "Measure:volume")

```

rename columns
```{r}
#aseg_volume just needs _Vol added
#regex mutate column names
#add a column for aseg vs aparc and grep based on that

# replace cortex with cort
# ventricle with vent

left_volume <- left_volume %>% rename_with(~str_replace(., "_volume", "_GrayVol"))

right_volume <- right_volume %>% rename_with(~str_replace(., "_volume", "_GrayVol"))

names(aseg_volume)[-1] <- names(aseg_volume)[-1] %>% 
  paste0("_Vol")

names(aseg_volume)[-1] <- paste0("FS_", names(aseg_volume)[-1])
```

```{r}
brain_area <- left_join(left_area, right_area, by = "subid")

brain_volume <- left_join(left_volume, right_volume, by = "subid")

brain_aparc <- left_join(brain_area, brain_volume, by = "subid")

brain <- left_join(brain_aparc, aseg_volume, by ="subid")
```

```{r}

brain_names <- brain

brain_names <- brain_names %>% 
  rename_with(~str_replace_all(., c("lh_" = "FS_L_", "rh_" = "FS_R_", "FS_Left-" = "FS_L_", "FS_Right-" = "FS_R_", "FS_lh" = "FS_L_", "FS_rh" = "FS_R_", "Cortex" = "Cort",
                    "Ventricle" = "Vent", "-" = "_")))

colnames(brain_names) <- gsub("(?<=^|_)([a-z])", "\\U\\1", colnames(brain_names), perl=T)

```

add extra columns to brain, pivot and cbind to pivoted training set
```{r}
#brain_copy <- brain
training_copy <- training_sample_baseline
namevector <- names(brain[177:211])
training_copy[,namevector] <- NA

row_num <- seq(1,211)
#feature_df <- as.data.frame(row_num, names(brain_copy), names(training_sample_baseline))

feature_df <- as.data.frame(cbind(row_num, names(brain), names(training_copy))) %>% 
  select(V2, V3) %>% 
  rename(nav_names = V2,
         model_names = V3)

#make values from 140 to end in nav NAs
feature_df$model_names[177:211] <- NA
```

regex test
- clean1 and 2 used because subsequent gsub statements will override the others, nesting didn't work. 
```{r}

#change lh_ to FS_L_, rh_ to FS_R_
#group$group.no.e <- gsub("e", "", group$group)
feature_df$nav_names_clean <- gsub("lh_", "FS_L_", feature_df$nav_names)
feature_df$nav_names_clean <- gsub("rh_", "FS_R_", feature_df$nav_names_clean)

#replace Left- with FS_L_
feature_df$nav_names_clean <- gsub("FS_Left-", "FS_L_", feature_df$nav_names_clean)
feature_df$nav_names_clean <- gsub("FS_Right-", "FS_R_", feature_df$nav_names_clean)

# replace cortex with cort
feature_df$nav_names_clean <- gsub("Cortex", "Cort", feature_df$nav_names_clean)
# ventricle with vent
feature_df$nav_names_clean <- gsub("Ventricle", "Vent", feature_df$nav_names_clean)
# - with _
feature_df$nav_names_clean <- gsub("-", "_", feature_df$nav_names_clean)

#capitalize first letter after _
feature_df$nav_names_clean <- gsub("(?<=^|_)([a-z])", "\\U\\1", feature_df$nav_names_clean, perl=TRUE)

```

find missing variables -old
```{r}
# Note: this is just one way to do this, and compares the new data to the training sample, which is not hosted publicaly. Comparing to feature_list.txt might be a workaround.

# get columns in analysis sample that aren't in brain
hcp_diff <- as.data.frame(setdiff(names(analysis_sample_baseline), names(brain)))

# get columns in analysis sample that aren't in brain
nav_diff <- as.data.frame(setdiff(names(brain), names(analysis_sample_baseline)))

max_length <- max(length(hcp_diff), length(nav_diff))

#set length of each vector equal to max length
length(hcp_diff) <- max_length                      
length(nav_diff) <- max_length

#cbind the two vectors together
cbind(hcp_diff, nav_diff)

diff_df <- as.data.frame(c(setdiff(names(analysis_sample_baseline), names(brain)), setdiff(names(brain), names(analysis_sample_baseline))))
```

rename remaining columns
```{r}
brain_names <- brain_names %>% 
rename(FS_LCort_GM_Vol = FS_L_CortVol_Vol,
FS_RCort_GM_Vol = FS_R_CortVol_Vol,
FS_TotCort_GM_Vol = FS_CortVol_Vol,   
FS_L_LatVent_Vol = FS_L_Lateral_Vent_Vol,
FS_L_InfLatVent_Vol = FS_L_Inf_Lat_Vent_Vol, 
FS_L_Cerebellum_WM_Vol = FS_L_Cerebellum_White_Matter_Vol,
FS_L_ThalamusProper_Vol = FS_L_Thalamus_Vol,
FS_3rdVent_Vol = FS_3rd_Vent_Vol, 
FS_4thVent_Vol = FS_4th_Vent_Vol,
FS_BrainStem_Vol = FS_Brain_Stem_Vol,
FS_L_Hippo_Vol = FS_L_Hippocampus_Vol,
FS_L_AccumbensArea_Vol = FS_L_Accumbens_Area_Vol,
FS_L_VentDC_Vol = FS_L_VentralDC_Vol,
FS_R_LatVent_Vol = FS_R_Lateral_Vent_Vol,
FS_R_InfLatVent_Vol = FS_R_Inf_Lat_Vent_Vol,
FS_R_Cerebellum_WM_Vol = FS_R_Cerebellum_White_Matter_Vol,
FS_R_ThalamusProper_Vol = FS_R_Thalamus_Vol,
FS_R_Hippo_Vol = FS_R_Hippocampus_Vol,
FS_R_AccumbensArea_Vol = FS_R_Accumbens_Area_Vol,
FS_R_VentDC_Vol = FS_R_VentralDC_Vol,
FS_CC_MidPosterior_Vol = FS_CC_Mid_Posterior_Vol,
FS_CC_MidAnterior_Vol = FS_CC_Mid_Anterior_Vol,
FS_InterCranial_Vol = ETIV.x.x,
FS_SupraTentorial_Vol = FS_SupraTentorialVol_Vol,
FS_SubCort_GM_Vol = FS_SubCortGrayVol_Vol)
```

remaining data features without corresponding model features:

FS_L_WhiteSurfArea_Area
BrainSegVolNotVent.x.x
FS_R_WhiteSurfArea_Area
BrainSegVolNotVent.y.x
ETIV.y.x
BrainSegVolNotVent.x.y
ETIV.x.y
BrainSegVolNotVent.y.y
ETIV.y.y
FS_L_vessel_Vol
FS_L_choroid_plexus_Vol
FS_R_vessel_Vol
FS_R_choroid_plexus_Vol
FS_5th_Vent_Vol
FS_WM_hypointensities_Vol
FS_L_WM_hypointensities_Vol
FS_R_WM_hypointensities_Vol
FS_Non_WM_hypointensities_Vol
FS_L_non_WM_hypointensities_Vol
FS_R_non_WM_hypointensities_Vol
FS_Optic_Chiasm_Vol
FS_BrainSegVol_Vol
FS_BrainSegVolNotVent_Vol
FS_LhCerebralWhiteMatterVol_Vol
FS_RhCerebralWhiteMatterVol_Vol
FS_CerebralWhiteMatterVol_Vol
FS_TotalGrayVol_Vol
FS_SupraTentorialVolNotVent_Vol
FS_MaskVol_Vol
FS_BrainSegVol_to_eTIV_Vol
FS_MaskVol_to_eTIV_Vol
FS_LhSurfaceHoles_Vol
FS_RhSurfaceHoles_Vol
FS_SurfaceHoles_Vol
FS_EstimatedTotalIntraCranialVol_Vol

cleaning
```{r}
#drop columns that aren't in model
common_cols <- intersect(colnames(training_sample_baseline), colnames(brain_names))

#test to make sure we don't have extras
col_test <- training_sample_baseline %>%
      select(!common_cols)

brain_subset <- brain_names %>% 
  select(Subid, common_cols)

```

### 2024-01-29 DrobininModelPredictions.Rmd
Added script for Drobinin model predictions

--- 2022-12-05 DataSetupOld.Rmd
@@
 ---
-title: "DataSetup"
+title: "Drobonin Model Prediction"
+author: "Lucy Whitmore"
+date: "10/6/2023"
 output: html_document
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+User notes: this script should be able to make predictions based on the Drobonin et al., 2022 BrainAGE model.
+The script is currently set up to predict for 2 waves of data, baseline and follow-up (ABCD 2 year followup).
 
-packages <-  c("tidyverse",
-               "reshape2",
-               "nlme", "lme4",
-               "data.table", "psych",
-               "parallel","lubridate",
-               "ggpubr", "broom", 
-               "apaTables", "MetBrewer")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-lapply(packages, library, character.only = TRUE)
+In order to run predictions, you'll need 2 things: the xgboost model, available at https://github.com/GitDro/DevelopmentalBrainAge/blob/master/model/xgboost_9to19_brain_age_mod.rds, and a data frame.
+The dataframe should have only interview age (in years) and columns for brain measures.
+IDs can be reattached after predictions are made, the rows won't change in order.
+
+```{r}
+library(tidymodels)
+library(xgboost) 
 ```
+Code for renaming columns
+```{r} 
+# The following code should rename ABCD smri variables to match the Drobonin model. Note, not all the features from the model exist in ABCD, so we also have to make empty columns for those.
 
-Load Structural Data
-```{r}
-#Load list of included subjects
-struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(1:12) %>% 
-  mutate(interview_age = as.numeric(interview_age),
-        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
-        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
-  filter(imgincl_t1w_include == 1)
-
-#Load structural data
-#Filter for volume & area measurements
-smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
-  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
-  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))
-
-smri_baseline <- smri_1%>% 
-  filter(eventname=="baseline_year_1_arm_1")
-
-smri_followup <- smri_1%>% 
-  filter(eventname=="2_year_follow_up_y_arm_1")
-
-#T1 intensity, might not need
-#smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
- # filter(!collection_title=="collection_title") %>% 
- # filter(src_subject_id%in%struc_include$src_subject_id)
-
-#Rename data to fit naming conventions
-#????
-ogfeatures <- rio::import("~/Documents/FYP/feature_list.txt")
-
-```
-
-Load Other Data
-```{r}
-#Puberty
-youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-#parent sum scores abcd_mhp01
-#youth sum scores abcd_mhy01
-
-#Cognition
-cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)
-
-#Height/weight metrics
-physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
-  filter(!collection_title=="collection_title")  %>% 
-  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)
-
-#Sexual attraction and behavior
-gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
-  filter(!collection_title=="collection_title")
-
-#Grade level, parent report
-grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
-```
-
-Template for renaming sMRI variables
-```{r, eval=F}
 renamed_smri_baseline <- smri_baseline %>% 
@@
 
-#renamed_smri <- subset(renamed_smri, select = -c(269,279))
-
 #Convert to numeric
@@
 
-smri_abcd_model <- renamed_smri_baseline[, 1:180] 
-```
-
-Rename sMRI for follow-up
-```{r}
-
-renamed_smri_followup <- smri_followup %>% 
-  rename(FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
-#FS_BrainSeg_Vol
-#FS_BrainSeg_Vol_No_Vent
-#FS_BrainSeg_Vol_No_Vent_Surf
-FS_LCort_GM_Vol = smri_vol_cdk_totallh,
-FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
-FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
-FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
-#FS_Total_GM_Vol #maybe just calculate?
-FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
-#FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
-#FS_SupraTentorial_No_Vent_Voxel_Count
-#FS_Mask_Vol
-#FS_BrainSegVol_eTIV_Ratio
-#FS_MaskVol_eTIV_Ratio
-FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
-FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
-FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
-FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
-FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
-FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
-FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
-FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
-FS_3rdVent_Vol = smri_vol_scs_3rdventricle,
-FS_4thVent_Vol = smri_vol_scs_4thventricle,
-FS_BrainStem_Vol = smri_vol_scs_bstem,
-FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
-FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
-FS_CSF_Vol = smri_vol_scs_csf,
-FS_L_AccumbensArea_Vol = smri_vol_scs_aal,
-FS_L_VentDC_Vol = smri_vol_scs_vedclh,
-#FS_L_Vessel_Vol
-#FS_L_ChoroidPlexus_Vol
-FS_R_LatVent_Vol = 	smri_vol_scs_ltventriclerh,
-FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
-FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
-FS_R_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexrh,
-FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
-FS_R_Caudate_Vol = smri_vol_scs_caudaterh,
-FS_R_Putamen_Vol = smri_vol_scs_putamenrh,
-FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh,
-FS_R_Hippo_Vol = smri_vol_scs_hpusrh,
-FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh,
-FS_R_AccumbensArea_Vol = 	smri_vol_scs_aar,
-FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
-#FS_R_Vessel_Vol
-#FS_R_ChoroidPlexus_Vol
-#FS_OpticChiasm_Vol
-FS_CC_Posterior_Vol = smri_vol_scs_ccps,
-FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps,
-FS_CC_Central_Vol = smri_vol_scs_ccct,
-FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat,
-FS_CC_Anterior_Vol = smri_vol_scs_ccat,
-FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh,
-FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh,
-FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh,
-FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh,
-FS_L_Entorhinal_Area = smri_area_cdk_ehinallh,
-FS_L_Fusiform_Area = smri_area_cdk_fusiformlh,
-FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh,
-FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh,
-FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh,
-FS_L_Lateraloccipital_Area = smri_area_cdk_locclh,
-FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh,
-FS_L_Lingual_Area = smri_area_cdk_linguallh,
-FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh,
-FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh,
-FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh,
-FS_L_Paracentral_Area = smri_area_cdk_paracnlh,
-FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh,
-FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh,
-FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh,
-FS_L_Pericalcarine_Area = smri_area_cdk_pericclh,
-FS_L_Postcentral_Area = smri_area_cdk_postcnlh,
-FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh,
-FS_L_Precentral_Area = smri_area_cdk_precnlh,
-FS_L_Precuneus_Area = smri_area_cdk_pclh,
-FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh,
-FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh,
-FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh,
-FS_L_Superiorparietal_Area = smri_area_cdk_supllh,
-FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh,
-FS_L_Supramarginal_Area = smri_area_cdk_smlh,
-FS_L_Frontalpole_Area = smri_area_cdk_frpolelh,
-FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh,
-FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh,
-FS_L_Insula_Area = smri_area_cdk_insulalh,
-FS_R_Bankssts_Area = smri_area_cdk_banksstsrh,
-FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh,
-FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh,
-FS_R_Cuneus_Area = smri_area_cdk_cuneusrh,
-FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh,
-FS_R_Fusiform_Area = smri_area_cdk_fusiformrh,
-FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh,
-FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh,
-FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh,
-FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh,
-FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh,
-FS_R_Lingual_Area = smri_area_cdk_lingualrh,
-FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh,
-FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh,
-FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh,
-FS_R_Paracentral_Area = smri_area_cdk_paracnrh,
-FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh,
-FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh,
-FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh,
-FS_R_Pericalcarine_Area = smri_area_cdk_periccrh,
-FS_R_Postcentral_Area = smri_area_cdk_postcnrh,
-FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh,
-FS_R_Precentral_Area = smri_area_cdk_precnrh,
-FS_R_Precuneus_Area = smri_area_cdk_pcrh,
-FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh,
-FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh,
-FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh,
-FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh,
-FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh,
-FS_R_Supramarginal_Area = smri_area_cdk_smrh,
-FS_R_Frontalpole_Area = smri_area_cdk_frpolerh,
-FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh,
-FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh,
-FS_R_Insula_Area = smri_area_cdk_insularh,
-#abcd technically doesn't say gray matter volume for the following varibles
-FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh,
-FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh,
-FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh,
-FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh,
-FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh,
-FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh,
-FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh,
-FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh,
-FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh,
-FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh,
-FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh,
-FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh,
-FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh,
-FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh,
-FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh,
-FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh,
-FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh,
-FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh,
-FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh,
-FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh,
-FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh,
-FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh,
-FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh,
-FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh,
-FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh,
-FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh,
-FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh,
-FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh,
-FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh,
-FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh,
-FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh,
-FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh,
-FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh,
-FS_L_Insula_GrayVol = smri_vol_cdk_insulalh,
-FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh,
-FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh,
-FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh,
-FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh,
-FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh,
-FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh,
-FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh,
-FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh,
-FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh,
-FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh,
-FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh,
-FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh,
-FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh,
-FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh,
-FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh,
-FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh,
-FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh,
-FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh,
-FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh,
-FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh,
-FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh,
-FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh,
-FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh,
-FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh,
-FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh,
-FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh,
-FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh,
-FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh,
-FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh,
-FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh,
-FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh,
-FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh,
-FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh,
-FS_R_Insula_GrayVol = smri_vol_cdk_insularh
-)
-
-#Create empty columns to match feature list
-renamed_smri_baseline[,"FS_BrainSeg_Vol"] <- NA
-renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent"] <- NA
-renamed_smri_baseline[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
-renamed_smri_baseline[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
-renamed_smri_baseline[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
-renamed_smri_baseline[,"FS_Mask_Vol"] <- NA
-renamed_smri_baseline[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
-renamed_smri_baseline[,"FS_MaskVol_eTIV_Ratio"] <- NA
-renamed_smri_baseline[,"FS_L_Vessel_Vol"] <- NA
-renamed_smri_baseline[,"FS_L_ChoroidPlexus_Vol"] <- NA
-renamed_smri_baseline[,"FS_R_Vessel_Vol"] <- NA
-renamed_smri_baseline[,"FS_R_ChoroidPlexus_Vol"] <- NA
-renamed_smri_baseline[,"FS_OpticChiasm_Vol"] <- NA
-renamed_smri_baseline[,"FS_Total_GM_Vol"] <- NA
-
-#renamed_smri <- subset(renamed_smri, select = -c(269,279))
-
-#Convert to numeric
-renamed_smri_baseline[,6:ncol(renamed_smri_baseline)] <- sapply(renamed_smri_baseline[,6:ncol(renamed_smri_baseline)],as.numeric)
-renamed_smri_baseline <- renamed_smri_baseline %>% 
-  mutate(interview_age = as.numeric(interview_age)/12)
-
-save(renamed_smri_baseline, file="renamed_smri_baseline.Rda")
-
-smri_abcd_model <- renamed_smri_baseline[, 1:180]
-
-```
-
-Split Samples
-```{r}
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1.Rda")
-load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2.Rda")
-load("renamed_smri_baseline.Rda")
-
-training_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_1$src_subject_id) %>% 
-  select(-c(subjectkey, src_subject_id, sex, eventname))
-
-analysis_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_2$src_subject_id)
-
-save(training_sample_baseline, file="training_sample_baseline.Rds")
-save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")
-
-```
-
-Make dataframe for ML testing
-```{r, eval = F}
-train_test <- training_sample_baseline %>% 
-  #select(1:15) %>% 
-  slice_head(n=20)
-
-#validation_test <- df_validation_baseline %>% 
-  #select(1:15) %>% 
-#  slice_head(n=20)
-
-save(train_test, file="train_test.Rds")
-#save(validation_test, file="validation_test.Rds")
-
-```
-
-Feature list template
-```{r, eval=F}
-#Template
-FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
-FS_BrainSeg_Vol
-FS_BrainSeg_Vol_No_Vent
-FS_BrainSeg_Vol_No_Vent_Surf
-FS_LCort_GM_Vol = smri_vol_cdk_totallh,
-FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
-FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
-FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
-FS_Total_GM_Vol #maybe just calculate?
-FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
-FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
-FS_SupraTentorial_No_Vent_Voxel_Count
-FS_Mask_Vol
-FS_BrainSegVol_eTIV_Ratio
-FS_MaskVol_eTIV_Ratio
-FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
-FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
-FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
-FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
-FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
-FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
-FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
-FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
-FS_3rdVent_Vol = 	smri_vol_scs_3rdventricle,
-FS_4thVent_Vol = smri_vol_scs_4thventricle,
-FS_BrainStem_Vol = smri_vol_scs_bstem,
-FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
-FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
-FS_CSF_Vol = 	smri_vol_scs_csf,
-FS_L_AccumbensArea_Vol = 	smri_vol_scs_aal,
-FS_L_VentDC_Vol = smri_vol_scs_vedclh,
-FS_L_Vessel_Vol
-FS_L_ChoroidPlexus_Vol
-FS_R_LatVent_Vol = smri_vol_scs_ltventriclerh,
-FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
-FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,
-
-# no cerebellum cort vol?
-FS_R_ThalamusProper_Vol = smri_vol_scs_tprh,
-FS_R_Caudate_Vol = smri_vol_scs_caudaterh,
-FS_R_Putamen_Vol = smri_vol_scs_putamenrh,
-FS_R_Pallidum_Vol = smri_vol_scs_pallidumrh,
-FS_R_Hippo_Vol = 	smri_vol_scs_hpusrh,
-FS_R_Amygdala_Vol = smri_vol_scs_amygdalarh,
-FS_R_AccumbensArea_Vol = smri_vol_scs_aar,
-FS_R_VentDC_Vol = smri_vol_scs_vedcrh,
-FS_R_Vessel_Vol
-FS_R_ChoroidPlexus_Vol
-FS_OpticChiasm_Vol
-FS_CC_Posterior_Vol = smri_vol_scs_ccps,
-FS_CC_MidPosterior_Vol = smri_vol_scs_ccmidps,
-FS_CC_Central_Vol = smri_vol_scs_ccct,
-FS_CC_MidAnterior_Vol = smri_vol_scs_ccmidat,
-FS_CC_Anterior_Vol = smri_vol_scs_ccat,
-FS_L_Bankssts_Area = 	smri_area_cdk_banksstslh,
-FS_L_Caudalanteriorcingulate_Area = smri_area_cdk_cdacatelh,
-FS_L_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrlh,
-FS_L_Cuneus_Area = 	smri_area_cdk_cuneuslh,
-FS_L_Entorhinal_Area = smri_area_cdk_ehinallh,
-FS_L_Fusiform_Area = smri_area_cdk_fusiformlh,
-FS_L_Inferiorparietal_Area = smri_area_cdk_ifpllh,
-FS_L_Inferiortemporal_Area = smri_area_cdk_iftmlh,
-FS_L_Isthmuscingulate_Area = smri_area_cdk_ihcatelh,
-FS_L_Lateraloccipital_Area = smri_area_cdk_locclh,
-FS_L_Lateralorbitofrontal_Area = smri_area_cdk_lobfrlh,
-FS_L_Lingual_Area = smri_area_cdk_linguallh,
-FS_L_Medialorbitofrontal_Area = smri_area_cdk_mobfrlh,
-FS_L_Middletemporal_Area = smri_area_cdk_mdtmlh,
-FS_L_Parahippocampal_Area = smri_area_cdk_parahpallh,
-FS_L_Paracentral_Area = smri_area_cdk_paracnlh,
-FS_L_Parsopercularis_Area = smri_area_cdk_parsopclh,
-FS_L_Parsorbitalis_Area = smri_area_cdk_parsobislh,
-FS_L_Parstriangularis_Area = smri_area_cdk_parstgrislh,
-FS_L_Pericalcarine_Area = smri_area_cdk_pericclh,
-FS_L_Postcentral_Area = smri_area_cdk_postcnlh,
-FS_L_Posteriorcingulate_Area = smri_area_cdk_ptcatelh,
-FS_L_Precentral_Area = smri_area_cdk_precnlh,
-FS_L_Precuneus_Area = smri_area_cdk_pclh,
-FS_L_Rostralanteriorcingulate_Area = smri_area_cdk_rracatelh,
-FS_L_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrlh,
-FS_L_Superiorfrontal_Area = smri_area_cdk_sufrlh,
-FS_L_Superiorparietal_Area = smri_area_cdk_supllh,
-FS_L_Superiortemporal_Area = smri_area_cdk_sutmlh,
-FS_L_Supramarginal_Area = smri_area_cdk_smlh,
-FS_L_Frontalpole_Area = smri_area_cdk_frpolelh,
-FS_L_Temporalpole_Area = smri_area_cdk_tmpolelh,
-FS_L_Transversetemporal_Area = smri_area_cdk_trvtmlh,
-FS_L_Insula_Area = smri_area_cdk_insulalh,
-FS_R_Bankssts_Area = smri_area_cdk_banksstsrh,
-FS_R_Caudalanteriorcingulate_Area = smri_area_cdk_cdacaterh,
-FS_R_Caudalmiddlefrontal_Area = smri_area_cdk_cdmdfrrh,
-FS_R_Cuneus_Area = smri_area_cdk_cuneusrh,
-FS_R_Entorhinal_Area = smri_area_cdk_ehinalrh,
-FS_R_Fusiform_Area = smri_area_cdk_fusiformrh,
-FS_R_Inferiorparietal_Area = smri_area_cdk_ifplrh,
-FS_R_Inferiortemporal_Area = 	smri_area_cdk_iftmrh,
-FS_R_Isthmuscingulate_Area = smri_area_cdk_ihcaterh,
-FS_R_Lateraloccipital_Area = smri_area_cdk_loccrh,
-FS_R_Lateralorbitofrontal_Area = smri_area_cdk_lobfrrh,
-FS_R_Lingual_Area = smri_area_cdk_lingualrh,
-FS_R_Medialorbitofrontal_Area = smri_area_cdk_mobfrrh,
-FS_R_Middletemporal_Area = smri_area_cdk_mdtmrh,
-FS_R_Parahippocampal_Area = smri_area_cdk_parahpalrh,
-FS_R_Paracentral_Area = smri_area_cdk_paracnrh,
-FS_R_Parsopercularis_Area = smri_area_cdk_parsopcrh,
-FS_R_Parsorbitalis_Area = smri_area_cdk_parsobisrh,
-FS_R_Parstriangularis_Area = smri_area_cdk_parstgrisrh,
-FS_R_Pericalcarine_Area = smri_area_cdk_periccrh,
-FS_R_Postcentral_Area = smri_area_cdk_postcnrh,
-FS_R_Posteriorcingulate_Area = smri_area_cdk_ptcaterh,
-FS_R_Precentral_Area = smri_area_cdk_precnrh,
-FS_R_Precuneus_Area = smri_area_cdk_pcrh,
-FS_R_Rostralanteriorcingulate_Area = smri_area_cdk_rracaterh,
-FS_R_Rostralmiddlefrontal_Area = smri_area_cdk_rrmdfrrh,
-FS_R_Superiorfrontal_Area = smri_area_cdk_sufrrh,
-FS_R_Superiorparietal_Area = 	smri_area_cdk_suplrh,
-FS_R_Superiortemporal_Area = smri_area_cdk_sutmrh,
-FS_R_Supramarginal_Area = smri_area_cdk_smrh,
-FS_R_Frontalpole_Area = smri_area_cdk_frpolerh,
-FS_R_Temporalpole_Area = smri_area_cdk_tmpolerh,
-FS_R_Transversetemporal_Area = smri_area_cdk_trvtmrh,
-FS_R_Insula_Area = 	smri_area_cdk_insularh,
-#abcd technically doesn't say gray matter volume for the following varibles
-FS_L_Bankssts_GrayVol = smri_vol_cdk_banksstslh,
-FS_L_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacatelh,
-FS_L_Caudalmiddlefrontal_GrayVol = smri_vol_cdk_cdmdfrlh,
-FS_L_Cuneus_GrayVol = smri_vol_cdk_cuneuslh,
-FS_L_Entorhinal_GrayVol = smri_vol_cdk_ehinallh,
-FS_L_Fusiform_GrayVol = smri_vol_cdk_fusiformlh,
-FS_L_Inferiorparietal_GrayVol = smri_vol_cdk_ifpllh,
-FS_L_Inferiortemporal_GrayVol = smri_vol_cdk_iftmlh,
-FS_L_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcatelh,
-FS_L_Lateraloccipital_GrayVol = smri_vol_cdk_locclh,
-FS_L_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrlh,
-FS_L_Lingual_GrayVol = smri_vol_cdk_linguallh,
-FS_L_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrlh,
-FS_L_Middletemporal_GrayVol = smri_vol_cdk_mdtmlh,
-FS_L_Parahippocampal_GrayVol = smri_vol_cdk_parahpallh,
-FS_L_Paracentral_GrayVol = smri_vol_cdk_paracnlh,
-FS_L_Parsopercularis_GrayVol = smri_vol_cdk_parsopclh,
-FS_L_Parsorbitalis_GrayVol = smri_vol_cdk_parsobislh,
-FS_L_Parstriangularis_GrayVol = smri_vol_cdk_parstgrislh,
-FS_L_Pericalcarine_GrayVol = smri_vol_cdk_pericclh,
-FS_L_Postcentral_GrayVol = smri_vol_cdk_postcnlh,
-FS_L_Posteriorcingulate_GrayVol = smri_vol_cdk_ptcatelh,
-FS_L_Precentral_GrayVol = smri_vol_cdk_precnlh,
-FS_L_Precuneus_GrayVol = smri_vol_cdk_pclh,
-FS_L_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracatelh,
-FS_L_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrlh,
-FS_L_Superiorfrontal_GrayVol = smri_vol_cdk_sufrlh,
-FS_L_Superiorparietal_GrayVol = smri_vol_cdk_supllh,
-FS_L_Superiortemporal_GrayVol = smri_vol_cdk_sutmlh,
-FS_L_Supramarginal_GrayVol = smri_vol_cdk_smlh,
-FS_L_Frontalpole_GrayVol = smri_vol_cdk_frpolelh,
-FS_L_Temporalpole_GrayVol = smri_vol_cdk_tmpolelh,
-FS_L_Transversetemporal_GrayVol = smri_vol_cdk_trvtmlh,
-FS_L_Insula_GrayVol = smri_vol_cdk_insulalh,
-FS_R_Bankssts_GrayVol = smri_vol_cdk_banksstsrh,
-FS_R_Caudalanteriorcingulate_GrayVol = smri_vol_cdk_cdacaterh,
-FS_R_Caudalmiddlefrontal_GrayVol = 	smri_vol_cdk_cdmdfrrh,
-FS_R_Cuneus_GrayVol = smri_vol_cdk_cuneusrh,
-FS_R_Entorhinal_GrayVol = smri_vol_cdk_ehinalrh,
-FS_R_Fusiform_GrayVol = smri_vol_cdk_fusiformrh,
-FS_R_Inferiorparietal_GrayVol = smri_vol_cdk_ifplrh,
-FS_R_Inferiortemporal_GrayVol = smri_vol_cdk_iftmrh,
-FS_R_Isthmuscingulate_GrayVol = smri_vol_cdk_ihcaterh,
-FS_R_Lateraloccipital_GrayVol = smri_vol_cdk_loccrh,
-FS_R_Lateralorbitofrontal_GrayVol = smri_vol_cdk_lobfrrh,
-FS_R_Lingual_GrayVol = smri_vol_cdk_lingualrh,
-FS_R_Medialorbitofrontal_GrayVol = smri_vol_cdk_mobfrrh,
-FS_R_Middletemporal_GrayVol = smri_vol_cdk_mdtmrh,
-FS_R_Parahippocampal_GrayVol = smri_vol_cdk_parahpalrh,
-FS_R_Paracentral_GrayVol = smri_vol_cdk_paracnrh,
-FS_R_Parsopercularis_GrayVol = smri_vol_cdk_parsopcrh,
-FS_R_Parsorbitalis_GrayVol = smri_vol_cdk_parsobisrh,
-FS_R_Parstriangularis_GrayVol = smri_vol_cdk_parstgrisrh,
-FS_R_Pericalcarine_GrayVol = smri_vol_cdk_periccrh,
-FS_R_Postcentral_GrayVol = smri_vol_cdk_postcnrh,
-FS_R_Posteriorcingulate_GrayVol = 	smri_vol_cdk_ptcaterh,
-FS_R_Precentral_GrayVol = smri_vol_cdk_precnrh,
-FS_R_Precuneus_GrayVol = smri_vol_cdk_pcrh,
-FS_R_Rostralanteriorcingulate_GrayVol = smri_vol_cdk_rracaterh,
-FS_R_Rostralmiddlefrontal_GrayVol = smri_vol_cdk_rrmdfrrh,
-FS_R_Superiorfrontal_GrayVol = smri_vol_cdk_sufrrh,
-FS_R_Superiorparietal_GrayVol = smri_vol_cdk_suplrh,
-FS_R_Superiortemporal_GrayVol = smri_vol_cdk_sutmrh,
-FS_R_Supramarginal_GrayVol = smri_vol_cdk_smrh,
-FS_R_Frontalpole_GrayVol = smri_vol_cdk_frpolerh,
-FS_R_Temporalpole_GrayVol = smri_vol_cdk_tmpolerh,
-FS_R_Transversetemporal_GrayVol = smri_vol_cdk_trvtmrh,
-FS_R_Insula_GrayVol = smri_vol_cdk_insularh
-
 ```
@@
 ```{r}
-#Puberty
-#Raw scores, youth
-youth_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
-  filter(!collection_title=="collection_title") %>% 
-  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
-  select(5:38) %>% 
-  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
-  mutate_at(c(5:28), as.numeric)
+# Import Drobonin BrainAGE Model
+mod_1 <- readRDS(
+  file = here::here(
+    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))
 
-#Raw scores, parent
-parent_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
-  filter(!collection_title=="collection_title")  %>% 
-  filter(eventname== "2_year_follow_up_y_arm_1") %>% 
-  select(5:27) %>% 
-  select(-c(interview_date, pds_select_language___1)) %>% 
-  mutate_at(c(5:21), as.numeric)
+# Note: model1_features_baseline refers to a dataframe with columns only for interview age and brain features, named in accordance with https://github.com/GitDro/DevelopmentalBrainAge/blob/master/feature_list.txt
 
-#Cognition Scores (Age-Corrected)
-cognition_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
-  filter(!collection_title=="collection_title")  %>% 
-  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
-  select(src_subject_id,interview_age, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected) %>% 
-  mutate_at(c(2:5), as.numeric)
+# check for any columns that aren't in model
+names.use <- names(model1_features_baseline)[!(names(model1_features_baseline) %in% mod_1$fit$feature_names)]
 
-#Existing Models
-corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")
+# baseline
+# remove those columns
+model1_features_baseline_clean <- model1_features_baseline %>% 
+  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))
+
+# reorder column names to match order of feature names
+newdata_baseline = as.matrix(model1_features_baseline_clean)
+newdata_baseline = newdata_baseline[ , mod_1$fit$feature_names]
+
+### Follow-Up
+# remove those columns
+model1_features_followup_clean <- model1_features_followup %>% 
+  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))
+
+# reorder column names to match order of feature names
+newdata_followup = as.matrix(model1_features_followup_clean)
+newdata_followup = newdata_followup[ , mod_1$fit$feature_names]
 ```
 
-## Format BrainAge
+run predictions - BASELINE
 ```{r}
-corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
-```
-## Format Puberty Data 
-```{r}
-#two or more missing values = exclude
-#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al.
-(2021).
-Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4).
-Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)
 
-youth_pds_followup[youth_pds_followup == 777] <- NA
-youth_pds_followup[youth_pds_followup == 999] <- NA
+# predict brainage - baseline
 
-#Youth-report, removes everyone w/o complete data
-youth_pds_followup <- youth_pds_followup %>% 
-  mutate(youth_sum = case_when(sex == "M" ~ #male
-                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
-                         sex == "F" ~ #female
-                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
-  mutate(youth_mean = youth_sum/5)
+# Predict
+# This prediction method gives the output as a vector, we'll fix that later on
+brainage_mod1_baseline <-
+  mod_1$fit %>%
+  predict(newdata = newdata_baseline)
 
-youth_pds_use_followup <- youth_pds_followup %>% 
-  select(src_subject_id, sex, youth_sum, youth_mean)
+# Calculate gap
+# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
+brain_age_df_mod1_baseline <- as.data.frame(brainage_mod1_baseline) %>%
+  rename(.pred = brainage_mod1_baseline) %>% 
+  mutate(
+    # provide the chronological age at time of scan
+    truth = model1_features_baseline$interview_age) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
 
-#Parent-report, removes everyone w/o complete data
-parent_pds_followup[parent_pds_followup == 777] <- NA
-parent_pds_followup[parent_pds_followup == 999] <- NA
-parent_pds_followup <- parent_pds_followup %>% 
-    mutate(parent_sum = case_when(sex == "M" ~ #male
-                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
-                         sex == "F" ~ #female
-                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
-  mutate(parent_mean = parent_sum/5)
+# bias correction
+# Slope and intercept are hardcoded
+bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
+bias_slope <- 0.55 # xgb_bias_mod$coefficients[["truth"]]
 
-parent_pds_use_followup <- parent_pds_followup %>% 
-  select(src_subject_id, sex, parent_sum, parent_mean)
+# create bias corrected data frame
+brain_age_df_mod1_baseline_corrected <- brain_age_df_mod1_baseline %>%
+  mutate(
+    # corrected brain age prediction
+    corrected_pred =  (.pred - bias_intercept) / bias_slope
+  ) %>%
+  # corrected corrected brain age gap
+  mutate(corrected_gap = corrected_pred - truth)
+
+# reattach IDs
+brain_age_df_mod1_baseline_corrected$src_subject_id <- harmonized_mri_baseline$src_subject_id
 ```
 
-## Make big dataframe
+run predictions - FOLLOW-UP
 ```{r}
-brainage_pds_followup <- corrected_followup_brainage[, c(6,1:5)] %>% 
-  left_join(youth_pds_use_followup) %>% 
-  left_join(parent_pds_use_followup) %>% 
-  left_join(cognition_followup)
+# predict brainage - followup
+brainage_mod1_followup <-
+  mod_1$fit %>%
+  predict(newdata = newdata_followup)
 
-#Split sample
-smp_size_followup <- floor(0.5 * nrow(brainage_pds_followup))
+# Calculate gap
+# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
+brain_age_df_mod1_followup <- as.data.frame(brainage_mod1_followup) %>%
+  rename(.pred = brainage_mod1_followup) %>% 
+  mutate(
+    # provide the chronological age at time of scan
+    truth = model1_features_followup$interview_age) %>%
+  # compute the brain age gap by subtracting chronological age from prediction
+  mutate(gap = .pred - truth)
 
-## set the seed to make your partition reproducible
-set.seed(123)
-sample_ind_followup <- sample(seq_len(nrow(brainage_pds_followup)), size = smp_size_followup)
+# create bias corrected data frame
+brain_age_df_mod1_followup_corrected <- brain_age_df_mod1_followup %>%
+  mutate(
+    # corrected brain age prediction
+    corrected_pred =  (.pred - bias_intercept) / bias_slope
+  ) %>%
+  # corrected corrected brain age gap
+  mutate(corrected_gap = corrected_pred - truth)
 
-sample_1_followup <- brainage_pds_followup[sample_ind_followup, ]
-sample_2_followup <- brainage_pds_followup[-sample_ind_followup, ]
-
-#t test for age differences
-t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)
-
-#Save samples
-save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
-save(sample_1_followup, file="Sample1FollowUp.Rda")
-save(sample_2_followup, file = "Sample2FollowUp.Rda")
-
+# reattach IDs
+brain_age_df_mod1_followup_corrected$src_subject_id <- harmonized_mri_followup$src_subject_id
 ```
-
-Attach baseline pds/cognition data to sample 2 / abcd model brainage
-
-## Format BrainAge
-```{r}
-abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id
-
-abcd_brainage_baseline <- abcd_brainage_baseline[, c(6,1:5)]
-```
-
-```{r}
-abcd_baseline <- sample_2 %>% 
-  select(1, 7:16)
-
-abcd_baseline <- left_join(abcd_baseline, abcd_brainage_baseline)
-
-str(abcd_baseline)
-
-save(abcd_baseline, file="abcd_baseline.Rda")
-```
-
-#################
-Make Dataframe of sMRI features for combined wave model
-################
-```{r}
-
-```
